### 2016-10-05 paper/draft.tex
first commit

\documentclass[11pt]{article}
\usepackage{coling2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{scrextend}
\makeatletter
\renewcommand*{\p@section}{\S\,}
\renewcommand*{\p@subsection}{\S\,}
\makeatother

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Syntactic realization with data-driven neural tree grammars}

% \author{Brian McMahan \\
%   Computer Science, Rutgers University \\
%   {\tt brian.mcmahan@rutgers.edu} \\\And
%   Matthew Stone \\
%   Computer Science, Rutgers University \\
%   {\tt matthew.stone@rutgers.edu} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}

A key component in surface realization in natural language generation
is to choose concrete syntactic relationships to express a target
meaning.
%
We develop a new method for syntactic choice based on learning a
stochastic tree grammar in a neural architecture.
%
This framework can exploit state-of-the-art methods for modeling word
sequences and generalizing across vocabulary.
%
Our methods also embed elementary tree structures using convolutional
coding to generalize over tree structures.
%
We evaluate the models on the task of linearizing unannotated
dependency trees, achieving improvements over previous methods.

\end{abstract}

\section{Introduction}

Where natural language understanding systems face problems of
ambiguity, natural language generation (NLG) systems face problems of
choice. A wide coverage NLG system must be able to formulate messages
using specialized linguistic elements in the exceptional circumstances
where they are appropriate; however, it can only achieve fluency by
expressing frequent meanings in routine ways. Empirical methods have
thus long been recognized as crucial to NLG; see
e.g. \newcite{langkilde1998generation}.

With traditional stochastic modeling techniques, NLG researchers have
had to predict choices using factored models with handcrafted
representations and strong independence assumptions, in order to avoid
combinatorial explosions and address the sparsity of training data.
By contrast, in this paper, we leverage recent advances in deep
learning to develop new models for syntactic choice that free
engineers from many of these decisions, but still generalize more
effectively and match human choices more closely than traditional
techniques.

We adopt the characterization of syntactic choice from
\newcite{bangalore2000exploiting} which uses a stochastic tree model and a language model to produce a linearized string from an unordered,
unlabeled dependency graph.
%
The first step to producing a linearized string is
to assign each node an appropriate supertag---a fragment of a parse tree with
a leaf left open for a lexical item---using a stochastic tree model.
%A language model is then used to select among the many possible derivation t
The resulting assignments are consistent with many possible derivation trees
because of adjunct ambiguitity.
%
To finish the linearization, a language model is used to select among them.
%
While improving the language model would improve the linearized string,
this work focuses on more accurately predicting the correct supertags from
unlabeled dependency trees.

In this work, we make two novel contributions to improve upon the stochastic tree model:
a supertag embedding technique which uses convolutional neural networks and
a recurrent tree network which can model hierarchical relationships.
%
By embedding supertags and using a recurrent tree network for the relationships between them,
we enable more information to be used at each node.
%
We evaluate our contributions in two ways: first, by varying the technique used to embed
supertags, and then by comparing a feed-forward model against our
recurrent tree model.

Our presentation begins in \ref{sec:tree} with an introduction to tree
grammars and a deterministic metholodogy for inducing the elementary trees of the grammar.
%
Next, \ref{sec:neural} presents the techniques we have developed to represent
a tree grammar using a neural architecture.
%
Then, in \ref{sec:models}, we describe the specific model we have
implemented, and in \ref{sec:algo}, we describe the algorithms used to describe the classification results.
%
The experiments in \ref{sec:expt} demonstrate the improvement of the
model over baseline results based on previous work on stochastic
surface realization.
%
We conclude with a brief discussion of the future potential for neural
architectures to predict grammatical choices in NLG.

\section{Tree Grammars}
\label{sec:tree}

Broadly, tree grammars are a family of tree rewriting formalisms that
modify structure to produce strings.  The atomic syntactic units are
called elementary trees; elementary trees combine using tree-rewrite
rules to form derived phrase structure trees describing complex
sentences.  Inducing a tree grammar involves fixing a formal inventory
of structures and operations for elementary trees and then inferring
instances of those structures to match corpus data.

\subsection{The Grammar Formalism}

The canonical tree grammar is perhaps lexicalized tree-adjoining
grammar (LTAG) \cite{joshi1991tree}.
%
The elementary trees of LTAG consist of two disjoint sets with distinct operations
%%%
The elementary trees of LTAG are disjoint sets based on the operations
they can perform: initial trees can perform substitution operations
and auxiliary trees can perform adjunction operations.
%
The substitution operation replaces a non-terminal leaf of a target
tree with an identically-labeled root node of an initial tree.
%
The adjunction operation modifies the internal structure of a target
tree at by expanding a node identically-labeled with the root and a
distinguished foot note in the auxiliary tree.
%
The lexicalization of the the grammar requires each elementary tree to
have at least one lexical item as a leaf.

LTAG incurs computational costs because it is mildly context-sensitive
in generative power.
%
Several variants reduce the complexity of the formalism by limiting
the range of adjunction operations.
%
For example, the Tree Insertion Grammar allows for adjunction as long as it is either a left or right auxiliary tree \cite{Schabes1995}.
%
Tree Substitution Grammars, meanwhile, allow for no adjunction and
only substitutions \cite{cohn2009inducing}.

We adopt one particular restricted adjunction, called
sister-adjunction or \emph{insertion}, which allows trees to attach to
an interior node and add itself as a first or last child
\cite{chiang2000statistical}.
%
Chiang's sister-adjunction allows for the flat structures in
the Penn Treebank while limiting the formalism to context-free power.

\subsection{Inducing a tree grammar}

In lexicalized tree grammars, the lexicon and the grammatical rules
are one and the same.
%
The set of possible grammatical moves which can be made are simultaneously the set of possible words which can be used next.
%
This means inducing a tree grammar from a dataset is a matter of inferring the set of constructions in the data.

%%%%\tabularnewline

We follow previous work in using bracketed phrase structure corpora
and deterministic rules to induce the grammar
\cite{chiang2000statistical,bangalore2001impact}.
%
Broadly, the methodology is to split the observed trees into the
constituents which make it up, according to the grammar formalism.
%
We use head rules
\cite{chiang2000statistical,collins1997three,magerman1995statistical}
to associate internal nodes in a bracketed tree with the lexical item
that owns it.
%
We use additional rules to classify some children as complements,
corresponding to substitution sites and root notes of complement
trees; and other children as adjuncts, corresponding to insertion
trees that combine with the parent node, either to the right or to the
left of the head; this allows us to segment the tree into units of
substitution and insertion.%
\footnote{One particular downside of deterministically constructing
  the grammar this way is that it can produce an excess of superfluous
  elementary trees.
%
We minimize this by collapsing repeated projections in the treebank.
%
Other work has provided Bayesian models for reducing grammar
complexity by forcing it to follow Dirichlet or Pitman-Yor processes
\cite{Cohn2010}---an interesting direction for future work.}

\begin{figure*}[tH!]
\centering
\includegraphics[width=\textwidth]{spineembed.pdf}
\caption{Embedding supertags using convolutional neural networks. In (A), a tree is encoded by its features and then embedded.
 In (B), convolutional layers are used to encode the supertag into a vector.}
 \label{fig:spineembedding}
\end{figure*}

\section{Neural Representations}
\label{sec:neural}

%There is a tension between vectorized representations and dynamic, nonlinear data structures.
%
%The tension is emphasized by requirement of specialized hardware that both dramatically increases the data throughput and reduces the space of representable models\footnote{Specifically, to run on Graphical Processing Units (GPUs), model structure must be known at compile time or incur heavy computational costs.}.
%
This work makes two contributions to improve stochastic tree modeling with neural networks.
%This work makes two contributions to reduce that tension and progress methodologies for representing nonlinear data structures in linear spaces.
%
First, supertags are represented as vectors through convolutional coding,
opening up the recent repertoire of vector space techniques.
%
Second, hierarchical dependence between nodes are modeled using a recurrent tree
network.
%
%recurrent neural network variant that isn't restricted to conditioning hidden states on their temporal predecessors.
%In brief, this technique is made possible through recent advanced in StackLSTMs \cite{dyer2015transition} and the use of topological indexing of recurrent states.% hidden states of

\subsection{Embedding Supertags}

%The first step to embedding supertags with a convolutional neural network is to
%embed each of its nodes and combine them to form a tensor.
%
%
%To embed a supertag, we first embed all of its nodes to form a tensor and then use a
%To embed a supertag and summarize it into a vector, each of its nodes are
%first embedded, combined to form a tensor, and then
%We embed supertags by embed each of its nodes, combine
To embed a supertag, each of its nodes are embedded, grouped to form a tensor, and then summarized into a single vector using a series of convolutional neural networks.
%
Since the weights of a convolutional neural network are optimized during training, the data should determine which parts of the supertag are important.
%
%The purpose of convolutional coding is to allow the data to determine which facets of a supertag are relevant and useful.
%
The more usual alternative approach is to treat each supertag as the atomic
unit---essentially a statistical singleton---and map it either to a feature or
an embedding vector.
%
We evaluate our contribution and compare these two approaches to get a sense for whether convolutional coding robustly generalizes across supertags.

The goal of embedding supertags is to summarize the tree structure into a single vector.
%
Inside the tree structure, there are two separate pieces of information which need to be summarized:
the syntactic category of each node and the role each node plays in the tree grammar.
%
The root receives a special role---either substitution or insertion to represent
the supertag's tree operation.
%g
The remaining nodes either have the substitution point role or the spine role---they are along the spine from root to the lexical attachment point.

The nodes of a supertag can be independently embedded and combined to form a
tensor of embeddings.
%
Specifically, symbols representing the syntactic category and node roles are treated as
distinct vocabulary tokens, mapped to integers, and used to retrieve a vector
representation that is learned during training.
%
The vectors are grouped into a tensor by placing the root node into the
first cell of the first row and left-aligning the decendents in the subsequent rows.
%
The two tensors are combined by concatenating along the embedding dimension.
%
This embed-and-group method is shown in on the left in Figure \ref{fig:spineembedding}.
%the root node is in first cell in the first row and the subsequent nodes are

Using a series of convolutional neural networks which learn their weights during training, the tensor of embeddings can be reduced to a single vector.
%
To reduce the tensor to a vector, the convolutions are designed with increasingly larger filter sizes.
%The convolutions are applied with increasingly larger filter sizes, leading to a reduction in dimensions.
%
Additionally, the dimensions are reduced alternatingly to also facilitate the capture of features.
%
The entire process is summarized in Eq. \ref{eq:embed} with $\Lambda$ representing the supertags, $G$ representing embedding matrices, and
$C$ representing the convolutional neural network layers.
%
Specifically, $G_s$ is the syntactic category embedding matrix and
$G_r$ is the node role embedding matrix.
%
Each convolutional layer $C$ is shown with its corresponding height and width as $C^{i,j}$.
%
The encoding first constructs the tensor, $T_\Lambda$, through the embed-and-group method.
%
Then, the embedding matrix $G_\Lambda$ is summarized from $T_\Lambda$ using the
series of convolutional layers.

\begin{align}
& T_\Lambda = [G_s(\Lambda_{syntactic~category});~G_r(\Lambda_{role})] \nonumber \\
& G_\Lambda = C^{4,5}(C^{3,1}(C^{1,3}(C^{2,1}(C^{1,2}(T_\Lambda))))) \label{eq:embed}
\end{align}

The final product, a vector per supertag, is aggregated with the other vectors
and turned into an embedding matrix.
%
This is visualized in on the right Figure \ref{fig:spineembedding}.
%
During training and test time, supertags are input as indices and their feature
representations retrieved as an embedding.
%
Importantly, the convolutional layers are connected to the computational graph during training, so the parameters are optimized with respect to the task.

%%%%
\subsection{Recurrent Tree Networks}
\label{subsec:rtn}

\begin{figure*}[tH!]
\centering
\includegraphics[width=\textwidth]{rtn.pdf}
\caption{A recurrent tree network. (A) The dependency structure as a tree.  (B) the dependency structure as a sequence.}
 \label{fig:rtn}
\end{figure*}

To model the relationship between the supertags in the derivation tree, we use
a recurrent tree network.
%
While recurrent neural networks normally use the previous hidden state in the sequential order of the inputs,
%require  the hidden state to be from the
%temporally previous hidden state,
recurrent tree networks use the hidden state from the parent.
%
Utilizing the parent's hidden state rather than the sequentially previous hidden
state, the recurrent connection can travel down the branches of a tree.
%
An example of a recurrent tree network is shown in \ref{fig:rtn}.

In our recurrent tree network, child nodes gain access to a parent's hidden state
through an internal \emph{tree state}.
%
During a tree recurrence, the nodes in the dependency graph are enumerated in a
top-down traversal.
%
At each step in the recurrence, the resulting recurrent state is stored in the
tree state at the step index.
%
Descendents access the recurrent state using a topological index that is passed in as data.

The formulation is summarized in Equation \ref{eq:rnnsetup}.
%
The input to each time step in the current tree is the data, $x_t$, and a topological
index, $p_t$.
%
The recurrent tree uses $p_t$ to retrieve the parent's hidden state, $s_p$, from the
tree state, $S_{tree}$, and applies the recurrence function, $g(\dot)$.
%
The resulting recurrent state is the hidden state for child node, $s_c$.
%
The recurrent state $s_c$ is stored in the tree state, $S_{tree}$, at index $t$.

\begin{align}
       s_c~&=~RTN(x_t,p_t) \nonumber \\
           &=~g(x_t,S_{tree}[p_t]) \nonumber \\
           &=~g(x_t, s_p) \nonumber \\
S_{tree}[t]&=~s_c \label{eq:rnnsetup}
\end{align}

The use of topological indices allows for many recurrent tree networks to be
run in parallel on a GPU, increasing the efficiency of the implementation.
%
The primary concern for running parallel GPU computations is homogeneity because
the same operation will be applied to the entire data structure.
%
Normally, tree operations require flow control, making homogeneity impossible.
%
However, using topological indices and a tree state eliminates the need for flow control
by creating locally linear, homogenous computations.

\section{Models}
\label{sec:models}

In this work, we present two stochastic tree models which represent different modeling decisions.
%
The first, Fergus-N\footnote{Fergus-N is shorthand for Fergus-Neuralized}, is a
reimplementation of \newcite{bangalore2000exploiting}'s original FERGUS model but
implemented in a feed-forward neural network.
%
The second model, Fergus-R\footnote{Fergus-R is shorthand for Fergus-Recurrent}, utilizes a recurrent tree network.
%
Fergus-N serves as a baseline model to which we compare the recurrent tree network.

For both stochastic tree models, a recurrent neural network language model is used to complete the linearization task.
%Both models produce partial results which a recurrent neural network language model is used to complete the linearization task.
%
%It is important to note that it is the same language model
%By using the same language model to finishing linearizing the output of both models, any observed differences in performance can be localized to the performance of the stochastic tree models.
%
The same language model is to eliminate the confound of
language model performance and measure performance differences in the
stochastic tree modeling.
%

\subsection{Model 1: Fergus-N}

Fergus-N is a stochastic tree model which uses local parent-child information as inputs to a feed-forward network.
%
Each parent-child pair are treated as independent of all other pairs.
%
The probability of the parent's supertag is computed using an embedding of the pair's lexical material and an embedding of the child's supertag.
%
During training, the negative log probability of the observed parent supertags is minimized for each parent-child pair.
%
This objective function is formally defined in Eq. \ref{eq:fergusn_obj} where $tag_p$ is the parent supertag, $tag_c$ is the child supertag, $lex_p$ is the parent's lexical material, and $lex_c$ is the child's lexical material.
%
Note that the probability of supertags for the leaves of the tree are computed
with respect to their parent's lexical material.

\begin{align}
&min_{\theta} -[\sum_p\sum_{p\to c} log[P_\theta(tag_{p} | lex_{p}, lex_{c}, tag_{c})] + \sum_c log[P_\theta(tag_{c} |lex_{p}, lex_{c})]] \label{eq:fergusn_obj}
\end{align}

The model is implemented as a feed-forward neural network that computes a probability distribution over parent supertags conditioned on lexical material and the child's supertag.
%
Equation \ref{eq:fergusn} details the model formulation.
%
The lexical material, $lex_p$ and $lex_c$, are embedded using the word embedding matrix, $G_w$, concatenated, and mapped to a new vector, $\omega_{lex}$, with a fully connected layer, $FC_1$.
%
The child supertag, $tag_c$, is embedded with $G_\Lambda$ and concatenated the lexical vector, $\omega_{lex}$, forming an intermediate vector representation of the node, $\omega_{node}$.
%
The node vector is repeated for each of the parent's possible supertags, $tagset_p$, and then concatenated with their embeddings to construct the set of treelet vectors, $\Omega_{treelet}$.
%
The vector states for the leaf nodes are similarly constructed, but instead combine the lexical vector, $\omega_{lex}$ with the embeddings of the child's possible supertags, $tagset_c$.
%
The final operation induces a probability distribution over the treelet and leaf vectors using a score computed by the vectorized function, $\Psi_{predict}$, as the scalar in a softmax distribution.

\begin{align}
&\omega_{lex} = FC_1([G_w(lex_p); G_w(lex_c)]) \label{eq:fergusn} \\
&\omega_{node}=concat([G_\Lambda(tag_c);~\omega_{lex}]) \nonumber \\
&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_\Lambda(tagset_p)]) \nonumber \\
&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_\Lambda(tagset_c)]) \nonumber \\
&P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
\frac{exp(\Psi(\omega_{treelet_i})))}
{\sum_{j \in |tagset_p|} exp(\Psi(\omega_{treelet_j})))} \nonumber \\
&P_\theta(tag_{c,i} |lex_{p}, lex_{c}) = 
\frac{exp(\Psi(\omega_{leaf_i})))}
{\sum_{j \in |tagset_c|} exp(\Psi(\omega_{leaf_j})))} \nonumber 
\end{align}

A decoding step uses the probability distributions to compute a high probability assignment for all supertags simueltaneously.
%
There are two primary difficulties that arise in this computation.
%
First, the conditional relationship of parents on children implies that the probability for the root supertag depends on the supertags of the entire tree.
%
Second, while it is easy to maintain local consistency---matching the syntactic
category of the child to an appropriate node on the parent---two children may choose substitution supertags which assume attachment to the same position on the parent.

To efficiently decode the supertag classifications, we implement an A* algorithm to incrementally select consistent supertag assignments.
%
At each step, the algorithm uses a priority queue to select subtrees based on their inside-outside scores.
%
The inside score is computed as the sum of the log probabilities of the supertags in the subtree.
%
The outside score is the sum of the best supertag for nodes outside the subtree, similar to \newcite{lewis2014improved}.
%
Once selected, the subtree is attached to the possible supertags of its parent that are both locally consistent and consistent among its already attached children.
%
These resulting subtrees are placed into the priority queue and the algorithm iterates to progress the search.
%
The search concludes either when a single complete tree has been found\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}.

\subsection{Model 2: Fergus-R}

Fergus-R is a stochastic tree model implemented in a top-down recurrent tree network and augmented with soft attention.
%
For each node in the input dependency tree, soft attention---a method which learns a vectorized function to weight a group of vectors and sum into a single vector---is used to summarize its children.
%

The advantage of using a recurrent tree is that nodes are able to be informed of their ancestors.
%
Each node is further informed using soft attention over its children ---a method which learns a vectorized function to weight a group of vectors and sum into a single vector.
%
The soft attention vector and the node's embedded lexical material serve as the input to the recurrent tree.
%
The output of the recurrent tree represents the vectorized state of each node and is combined with each node's possible supertags to form prediction states.
%
%The recurrent tree output and the embedded supertag of the node's parent are used to compute a probability distribution over supertags.
%
%Using only summarized lexical information from its ancestors and children, the model computes a probability distribution over supertags for each node's recurrent tree output.
%
%so that the soft attention and recurrence relationships are only communicating lexical information.
Importantly, removing the conditional dependence on descendents' supertags
results in the simplified objective function in Eq. \ref{eq:rtnobj} where $lex_C$ is the children's lexical information, $lex_p$ is the parent's lexical information, $tag_p$ is the supertag for the parent node, and $RTN$ is the recurrent tree network.

\begin{align}
&min_{\theta} -[\sum_{(p,C)} P_\theta(tag_{p}|RTN,~lex_p,~lex_{C})] \label{eq:rtnobj}
\end{align}

The Fergus-R model uses only lexical information as input to calculate the probability distribution over each node's supertags. 
%
The specific formulation is detailed in Eq. \ref{eq:fergusr}.
%
First, a parent node's children, $lex_C$, are embedded using the word embedding matrix, $G_w$, and then summarized with an attention function, $\Psi_{attn}$, to form the child context vector, $\omega_{C}$. 
%
The child context is concatenated with the embedded lexical information of the parent node, $lex_p$, and mapped to a new vector space with a fully connected layer, $FC_1$, to form the lexical context vector, $\omega_{lex}$.
%
The context vector and a topological vector for indexing the internal tree state (see \ref{subsec:rtn}) are passed to the recurrent tree network, $RTN$, to compute the full state vector for the parent node, $\omega_{node}$.
%
Similar to Fergus-N, the state vector is repeated and concatenated with the vectors of the parent node's possible supertags, ${tagset}_p$, and mapped to a new vector space with a fully connected layer, $FC_2$.
%
A vector in this vector space is labeled $\omega_{elementary}$ because the combination of supertag and lexical item constitutes an elementary tree.
%
The last step is to compute the probability of each supertag using the vectorized function, $\Psi_{predict}$.

\begin{align}
&\omega_{C} = \Psi_{attn}(G_w(lex_C)) \label{eq:fergusr} \\
&\omega_{lex} = FC_1(concat(\omega_{C},~G_w(lex_p))) \nonumber \\
&\omega_{node} = RTN(\omega_{lex},~topology) \nonumber \\
&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_\Lambda(tagset_p))) \nonumber \\
&P_\theta(tag_{p,i}~|~RTN,~lex_p,~lex_{C}) = 
\frac{exp(\Psi_{predict}(\omega_{elementary_i})))}
{\sum_{j \in |\Omega|} exp(\Psi_{predict}(\omega_{elementary_j))}} \nonumber
\end{align}

Although the same A* algorithm from Fergus-N is used, the decoding for Fergus-R is far simpler.
%
As supertags are incrementally selected in the algorithm, the inside score of the subsequent subtree is computed.  
%
Where Fergus-N had to compute a incremental dynamic program to evaluate the inside score, Fergus-R decomposes into a sum of conditionally independent distributions. 
%
The resulting setup is a chart parsing problem where the inside score of combining two consistent (non-conflicting) edges is just the sum of their inside scores.

\subsection{Linearization}

The final step to linearizing the output of Fergus-N and Fergus-R---a dependency tree annotated with supertags and partial attachment information---is a search over possible orderings with a language model. 
%
There are many possible orderings due to adjunction operation
What remains to be determined is the order in which adjuncts are attached. 
%
Following \newcite{bangalore2000exploiting}, a language model is used to select between the alternate orderings. 
%
The language model used is a two-layer LSTM trained using the Keras library on
the surface form of the Penn Treebank.
%
The surface form was minimally cleaned\footnote{With respect to the surface form, the only
cleaning operations were to merge proper noun phrases into single tokens.  Punctuation and other
common cleaning operations were not performed.} to simulate realistic scenarios.

The difficulty of selecting orderings with a language model is that the possible linearizations can grow exponentially.
%
In particular, our implementations result in a large amount of insertion trees\footnote{Many of the validation examples had more than $2^{40}$ possible linearizations.}.
%
We approach this problem using a prefix tree which stores the possible linearizations as back-pointers to their last step and the word for the current step. 
%
The prefix tree is greedily searched with 32 beams.

\section{Experiments}
\label{sec:expt}

In this work, we make two contributions to stochastic tree modeling and evaluate them on a supertag prediction and linearization task. 
%
The evaluation uses the Wall Street Journal sections of the Penn Treebank which have been previously used for statistical tree grammars \cite{chiang2000statistical}\footnote{A possible additional data source, the data from the 2011 Shared Task on Surface Realization, was not available}.
%
Our data pipeline transforms the treebank into a collection of elementary trees and derivation tree---the structured relationships required to compose them into the observed data (see \ref{}).
%
Abstracting syntactic information from the derivation tree leads to the unlabeled dependency trees our models assume as input.

We evaluate the convolutional coding of the supertags by comparing it against standard embedding techniques.
%
Specifically, we construct a matrix that has as many rows as supertags and is initialized with a Glorot uniform distribution.
%
Fergus-N and Fergus-R are retrained with the exact same parameters, except that instead of the supertag indices referencing a matrix constructed by a series of convolutions, it references our constructed embedding matrix (which gets optimized during learning). 
%
The impact of the embedding will be different for each model---Fergus-N uses embedded supertags in constructing the node state and Fergus-R does not---but any differences between the models and their alternate embedding counterparts can be perceived as evidence for either embedding technique.

To evaluate whether a recurrent tree network can capture hierarchical and long-distance relationships, we compare it against a feed-forward model. 
%
Specifically, for each model, we measure how often it predicts the correct supertag or if it predicts a different supertag but with the correct root syntactic category, root node role, or lexical leaf node\footnote{The leaf to which a lexical item would attach} syntactic category.
%
These four quantities represent different aspects that are important to get right in a stochastic tree model.

%Fergus-R's predictions never observe its neighboring nodes' supertags, so any predictive improvements are based solely on lexical and ancestral context.

We also compare the differences in supertag embeddings and stochastic tree models on the linearization task itself.
%
The linearization tasks licenses more freedom in supertag classifications because it may have the same root and leaf syntactic categories but differ in minor constructions, such as number of arguments. 
%
The freedom allows for models to be more generalized and less fit to the specific supertags, but at the same time, it also mutes the distinctions between classification decisions.
%
The metrics used in the linearization task were taken from previous work, who argue that it correlates with generation performance better than n-gram counting methods.

\subsection{Results}

\begin{figure}
\centering
\begin{tabular}{|l|p{2cm}|p{1.5cm}|}
\hline
Model & Test Set Size & Accuracy \\ \hline
Fergus-Penn & 100 & 0.749 \\
\hline \hline
Fergus-N $Top_{40}$  & 2187 & 0.699 \\
\hline
\end{tabular}
\label{table:results}
\caption{ }
\end{figure}

\section{Related Work}
\label{sec:relatedwork}

\section{Conclusion}
\label{sec:conclusion}

\bibliography{coling2016}
\bibliographystyle{acl}

\appendix

Model specifications
- layers, learning, etc

Head to supertag mapper

The data used in our experiments was taken from the Wall Street Jounral sections of the Treebank corpus.

For the unlabeled dependency trees, we used the derivation structures which resulted from the deterministic grammar induction (see \ref{}).

For the Fergus-N and Fergus-R learning problems, we processed the data into their relevant tuples.

Supertags were preprocessed into two data structures: one for the convolutional coding and one to map lexical items to supertags. 
%
The supertag data structure used for convolutional coding is a three-dimensional tensor where the first dimension is the supertag index, the second dimension is depth from root, and the third is relative sibling index. 
%
The tokens for syntactic category and node role are mapped to indices, stored in the data structure, and the data structure is stored in GPU memory.

During

in model intro:
    - learning parameters in appendix
    - used keras + theano

in conv coding maybe:
    - gpu data structure

in experiments
    - where the data comes from

\end{document}

### 2016-10-05 fergus/algorithms/__init__.py

@@ -1,0 +1,4 @@
+from __future__ import absolute_import, print_function
+from . import tagger
+from .. import models
+#from . import linearizer

### 2016-10-05 fergus/algorithms/linearizer/__init__.py

@@ -1,0 +1,49 @@
+from __future__ import absolute_import
+from . import partialtree, astar, dp, states, decode
+
+from sqlitedict import SqliteDict as SD
+import numpy as np
+import time
+from . import utils
+import sys
+
+def run(tree_list, *args, **kwargs):
+    partial = partialtree.PartialTree.from_list(tree_list)
+    partial.prep()
+    results = astar.run(partial, *args, **kwargs)
+    return results
+
+def run_dp(tree_list, *args, **kwargs):
+    partial = partialtree.PartialTree.from_list(tree_list)
+    partial.prep()
+    print(partial.measure_difficulty())
+    try:
+        results = dp.run(partial, *args, **kwargs)
+    except utils.MemoryException as me:
+        print(me.msg)
+        results = []
+    return results, partial.measure_difficulty()
+
+def run_tests():
+    utils.level = 1
+    db_path = '/home/cogniton/research/code/gist/gist/alternate_models/fergus616_dev_R.db'
+    with SD(db_path, tablename='indata') as db:
+        data = list(db.items())
+
+    n = len(data)
+    r_times = []
+    s_times = []
+    start = time.time()
+    last = start
+    import pprint
+    for idx in np.random.choice(np.arange(len(data)), 10, False):
+        pprint.PrettyPrinter(indent=2).pprint(data[idx][1][0])
+        run(data[idx][1][0])
+        now = time.time()
+        r_times.append(now-start)
+        s_times.append(now - last)
+        last = now
+        print("[tic. r:{:0.2f} toc. s:{:0.2f}] \r".format(now-start, now-last))
+        print(data[idx][1][1])
+        sys.stdout.flush()
+    print("finished.  {:0.2f} seconds".format(time.time()-start))

### 2016-10-05 fergus/algorithms/linearizer/__main__.py

@@ -1,0 +1,186 @@
+from __future__ import print_function, absolute_import, division
+from ..linearizer import run, run_dp
+from sqlitedict import SqliteDict as SD
+import numpy as np
+import time
+from . import utils
+import sys
+try:
+    input = raw_input
+except:
+    pass
+
+def test():
+    test_datum = (u'(ROOT(S(NP)(VP(VBD ratified)(NP))))',
+                     [(u'(NP(NNS workers))',
+                       [(u'(NP*(NNS communications))', []),
+                        (u'(NP*(DT the))', []),
+                        (u'(*NP(PP(IN of)(NP)))', [(u'(NP(NNP america))', [])])]),
+                      (u'(NP(NP)(CC and)(NP))',
+                       [(u'(NP(NN contract))',
+                         [(u'(NP*(JJ regional))', []),
+                          (u'(NP*(JJ new))', []),
+                          (u'(NP*(DT a))', [])]),
+                        (u'(NP(DT all))',
+                         [(u'(NP*(CC but))', []),
+                          (u'(NP*(CD one))', []),
+                          (u'(*NP(PP(IN of)(NP)))',
+                           [(u'(NP(NNS agreements))',
+                             [(u'(NP*(JJ local))', []),
+                              (u'(NP*(DT the))', []),
+                              (u'(*NP(PP(IN with)(NP)))',
+                               [(u'(NP(NNS bell_atlantic_corp))', [])])])])])]),
+                      (u'(*S(. .))', [])])
+    run(test_datum)
+
+
+def crap():
+    datum = ( u'(ROOT(S(NP)(VP(VBD determined)(SBAR))))',
+  [ ( u'(NP(NN panel))',
+      [ (u'(NP*(JJ international))', []),
+        (u'(NP*(DT an))', []),
+        ( u'(*NP(VP(VBN set)))',
+          [ (u'(VP*(IN up))', []),
+            ( u'(*VP(PP(IN under)(NP)))',
+              [ ( u'(NP(NNS rules))',
+                  [ (u'(NP*(DT the))', []),
+                    ( u'(*NP(PP(IN of)(NP)))',
+                      [ ( u'(NP(NN agreement))',
+                          [ (u'(NP*(JJ general))', []),
+                            (u'(NP*(DT the))', []),
+                            ( u'(*NP(PP(IN on)(NP)))',
+                              [ ( u'(NP(NN trade))',
+                                  [ (u'(NP*(CC and))', []),
[…]

### 2016-10-05 fergus/algorithms/linearizer/astar.py

@@ -1,0 +1,284 @@
+'''
+A* core 
+
+a linearizer takes as input a partial tree
+the partial tree is a linked-daughter-list tree with elementary spines assigned to each node
+each node can then bind in a limited number of ways given its elementary spine
+
+a* chart parsing
+- enumerate position ordering for children
+- substitution nodes don't move (bindings could, in theory swap)
+- an insertion which has multiple binding spots could change the ordering number of other nodes
+- so, for each insertion, for each of their positions, enumerate the gorn orderings
+- this is basically finding the permutations based on gorn addressing
+- these are assigned indices in order
+- children are then put into the *queue with uniqueness enforced on position index
+- pubsub configuration used; 
+- sub to completed events to find children, open events to find parents
+- pub to completed events when finished, open events when still need more children
+- algo is:
+    + pop from queue
+    + publish appropriately
+    + subscribe appropriate
+
+'''
+from __future__ import absolute_import, print_function, division
+import heapq
+from collections import defaultdict, deque
+from .utils import logprint
+from . import utils
+
+import sys
+sys.setrecursionlimit(100000)
+#sys.settrace()
+
+class StarCache(object):
+    def __init__(self):
+        self._cache = dict()
+        self._mapping = dict()
+        self._flip = dict()
+        self._id = 0
+    
+    def __getitem__(self, state_id):
+        return self._cache[state_id][1]
+            
+    def __call__(self, state):
+        _mapping = self._mapping
+        _flip = self._flip
+        _cache = self._cache
+        
+        state_key = state.key
+        score = state.score
+        
+        if state_key in _mapping:
+            # want the min score
+            state_id = _mapping[state_key] 
+            if _cache[state_id][0] >= score:
+                _cache[state_id] = (score, state)
+            return state_id
+        else:
+            state_id = self._id
+            _cache[state_id] = (score, state)
+            _mapping[state_key] = state_id
+            _flip[state_id] = state_key
+            self._id += 1
+            return state_id
+            
+    def solution(self, states=None, state_ids=None, all_solutions=False):
+        states = states or [self[state_id] for state_id in state_ids]
+        states_ = filter(lambda state: state.completed and state.op_pos=='ROOT', 
+                        states)
+        if len(states_) == 0:
+            states_ = filter(lambda state: state.op_pos=='ROOT', states)
+            if len(states_) == 0:
+                return False
[…]

### 2016-10-05 fergus/algorithms/linearizer/decode.py

@@ -1,0 +1,112 @@
+from __future__ import absolute_import, print_function, division
+from . import utils
+from fergus.models import language_model
+from sqlitedict import SqliteDict as SD
+from tqdm import tqdm
+import numpy as np
+import editdistance
+import os
+import sys
+
+######## different ways to decode
+
+def w2v(words, igor):
+    vec = np.zeros((igor.batch_size, igor.max_sequence_len), dtype=np.int32)
+    for i, word in enumerate(words):
+        vec[0, i] = igor.vocabs.words[word]
+    return vec
+
+def step(model, sentence, choices, verbose=0):
+    vec = w2v(sentence)
+    idx = len(sentence) - 1# with just <start>, len is 1. we want post over start, so subtract 1
+    probs = model.model.predict(vec)[0,idx,:]
+    best = 0.
+    best_word = ""
+    best_idx = -1
+    decisions = []
+    for i, word in enumerate(choices):
+        word_val = probs[model.igor.vocabs.words[word]]
+        decisions.append((word, word_val))
+        if word_val > best:
+            best = word_val
+            best_word = word
+            best_idx = i
+    #if verbose:
+    #    print("selecting {}".format(best_word))
+    #    print("-------")
+    return sentence + [best_word], best_idx, decisions
+
+def mle(memos, verbose=0):
+    bwd, fwd = memos
+    cur_idx = -1
+    sentence = ['<START>']
+    decisions = []
+    while cur_idx in fwd:
+        next_idxs = fwd[cur_idx]
+        choices = [bwd[poss_idx][1] for poss_idx in next_idxs]
+        sentence, selection_idx, decision = step(model, sentence, choices, verbose)
+        decisions.append(decision)
+        cur_idx = list(next_idxs)[selection_idx] 
+    if verbose: print(' '.join(sentence))
+    return cur_idx, decisions
+
+
+def vectorize(sentences, igor):
+    out = np.zeros((igor.batch_size, igor.max_sequence_len), dtype=np.int32)
+    tp1_idx = np.zeros((igor.batch_size), dtype=np.int32)
+    for i, sentence in enumerate(sentences):
+        for j, word in enumerate(sentence):
+            out[i,j] = igor.vocabs.words[word]
+        tp1_idx[i] = j ## will be len(sentence) - 1
+    return out, tp1_idx
+
+def beam(memos, model, verbose=0):
+    igor = model.igor
[…]

### 2016-10-05 fergus/algorithms/linearizer/dp.py

@@ -1,0 +1,138 @@
+'''run the dynamic program for linearization
+
+process:
+    input the partial tree
+    instantiate all nodes as states
+    iterate over states until no states left (last one should be complete):
+        if state is complete, try to advance dot of parent
+            if successful, state is done
+        else, try to find daughter to advance dot
+            if successful, daughter is done
+        
+'''
+from collections import defaultdict, deque
+from copy import deepcopy
+import sys
+import psutil
+from . import utils
+
+
+sys.setrecursionlimit(40000)
+
+def run(partial, verbose=1, *args, **kwargs):
+    '''
+    run the dynamic program. 
+    iterate over states, combining children with parents
+    finish when last parent is complete
+    '''
+    state_map = {state._id:state for state in partial.enumerate_states(version='dp')}
+    stk = deque(state_map.values())
+    top_listeners = defaultdict(lambda: None)
+    bottom_listeners = defaultdict(lambda: None)
+    for state in stk:
+        if state.completed:
+            bottom_listeners[state._id] = state
+        else:
+            top_listeners[state.next_dot] = state
+    counter = 0
+    utils.logprint(2, "Starting the dynamic program")
+    while stk:
+        #if psutil.virtual_memory().percent > 90.:
+        #    raise utils.MemoryException("search explosion")
+        counter += 1
+        utils.logprint(3, "iteration<{}> -- stack<{}>".format(counter, len(stk)))
+        if counter > 10**4:
+            print('crap')
+            import pdb
+            pdb.set_trace()
+        # get next thing off stack
+        state = stk.popleft()
+        
+        # this skips states that were advanced by a parent being popped
+        if state.advanced:
+            continue
+        
+        # it is a child that could progress a dot
+        if state.completed and state._id == partial._id:
+            utils.logprint(1, "Decoding....\r")
+            #sys.stdout.flush()
+            results = utils.prefix_decoding(state, state_map)
+            utils.logprint(1, "{} solutions".format(len(results[0])))
+            assert len(results[0]) > 0
+            return results
+        
+        if state.completed:
+            utils.logprint(3, 'firing completed state<{}>'.format(len(state.position_indices)))
+            # try to get a parent. if it's listening for this state, I think it's guaranteed to have it as its next dot
+            parent = top_listeners[state._id] # top are those from above listening for state
+            # the None check is for parents who aren't ready for this state
+            if parent is not None and parent.next_dot == state._id:
[…]

### 2016-10-05 fergus/algorithms/linearizer/partialtree.py

@@ -1,0 +1,254 @@
+"""A dynamic programming algorithm to linearize unordered tree grammar trees
+
+Plan:
+    1. each tree orders its children by their gorn prefix
+    2. split the children into two groups: left and right
+    3. iterate through each prefix group in the left-right groups
+    4. going from left to right ala prefix, assign position numbers to children
+        - don't worry about situations which offset child indices for no reason
+            - this occurs when an insertion can go higher up in the tree and be to the left of others
+            - but it also has an option to go to the right of the others. 
+            - if we naively enumerate all positions, the possibly-empty position will still be marked
+        - assign each children the fully combination number (so if 3 children in group, t+1, t+2, t+3)
+    5. once all nodes have their indices, spawn them with it
+    6. nodes with no daughters are marked as complete
+    7. each other node specifies an numbered iteration over its daughters and looks for hte first
+    8. upon finding a daughter, a node accepts it if its position number is not taken
+        - this means we will have some duplicate states, but completed daughters could mark their completeness
+          to make this not as much of an issue. 
+
+guarantees:
+    - because a node doesn't accept any position or child more than once, most states won't repeat
+
+PartialTree's responsibility:
+    - enumerate the prefixes and tell all children their possible position ids
+    - be able to produce its id, a separate instantiation for each of its position ids, and 
+      the list of its daughter ids
+
+State's use of PartialTree:
+    - state will ask for ptree to make all of its instantiations
+    - it will make a single state for each of them
+    - this state will be responsible for tracking the children it's accepted so far
+    - the state will accept a child, include its position number in the already done
+
+the only way the numbering system proposed could result in a bad state is if:
+    - selecting one child meant that another child's spot was impossible
+    - the other spot would also not have been taken yet
+    - could we have two substutions at the same site?
+    - no, because we don't do the multiple ordering spaces, 
+        + nevermind. yes. we need to keep track of the sub valencies so far
+        + in addition to the ordering valency
+        + because, for example, we don' tknow what comes first, an insertion or a substitution
+        + so we allow for both instantiations by giving them two possible child numbers
[…]

### 2016-10-05 fergus/algorithms/linearizer/states.py

@@ -1,0 +1,207 @@
+from copy import deepcopy 
+from . import utils
+
+class StarState(object):
+    def __init__(self, node, position_indices): 
+        self.node_id = node._id
+        #self.sub_gorn = None if node.tree.is_insertion else node.pos2gorn[position_index]
+        self.is_insertion = node.tree.is_insertion
+        self.sub_points = node.sub_points
+        self.pos2gorn = node.pos2gorn
+
+        self.dot_ids = [d._id for d in node.daughters]
+        self.dot_map = {d._id:d.head for d in node.daughters}
+        self.n_daughters = len(self.dot_map)
+        self.next_dot = (self.dot_ids[0] if (len(self.dot_ids)>0) else None)
+
+        self.finished = {(0,node.head_index,):node.head}
+
+        self.positions_filled = tuple()
+
+        self.n_finished = 0
+        self.position_indices = position_indices
+        if self.completed:
+            self.seq = tuple(x[1] for x in sorted(self.finished.items()))
+        else:
+            self.seq = tuple(sorted(self.finished.items()))
+        # self.key = hash((self.node_id, self.position_index, 
+        #                  self.n_finished, 
+        #                  self.sub_points,
+        #                  self.positions_filled,
+        #                  self.seq))
+        self.key = hash((self.node_id, self.seq))
+
+        self.score = -1
+
+
+    def step(self, child_state):        
+        cls = self.__class__
+        for pos_idx in child_state.position_indices:
+            if pos_idx in self.positions_filled:
+                continue
+                #return None
+
+            if not child_state.is_insertion:
+                gorn = child_state.pos2gorn[pos_idx]
+                if gorn not in self.sub_points:
+                    #return None
+                    continue
+                else:
+                    new_subs = tuple([pt for pt in self.sub_points if pt != gorn]) 
+            else:
+                new_subs = self.sub_points
+
+            try:
+                assert child_state.node_id == self.next_dot
+            except:
+                import pdb
+                pdb.set_trace()
+
+            new_state = cls.__new__(cls)
+            new_state.__dict__.update(self.__dict__)
+
+            new_state.sub_points = new_subs
+
+            new_state.finished = copy(self.finished)
+            new_state.finished.update({(0,pos_idx,)+k[1:]:v 
+                                       for k,v in child_state.finished.items()})
+
+            try:
+                assert len(new_state.finished) == len(self.finished) + len(child_state.finished)
+            except:
+                import pdb
+                pdb.set_trace()
[…]

### 2016-10-05 fergus/algorithms/linearizer/utils.py

@@ -1,0 +1,221 @@
+from __future__ import print_function 
+import psutil
+from collections import defaultdict
+import numpy as np
+from copy import copy
+import ikelos as ike
+
+level = 0
+## for now, i'm lazy. everything debugs. 
+logger = ike.utils.loggers.duallog('linearizer')
+tolog = {i:logger.debug for i in range(1, 6)}
+tolog[0] = logger.info
+
+try:
+    range = xrange
+except:
+    pass
+
+def logprint(lvl, *msg): 
+    # return
+    #print(lvl, level)
+    if level >= lvl:
+        tolog[lvl](' '.join(map(str, msg)))
+        #print(*msg)
+        
+
+class MemoryException(Exception):
+    def __init__(self, msg):
+        self.msg = msg
+        self.usage = psutil.virtual_memory().percent
+    def __str__(self):
+        return "OutOfMemory<{}%%>: {}".format(self.usage, self.msg)
+
+    
+
+def decode_one(memos, seq_idx, verbose=0):
+    bwd_memo, fwd_memo = memos
+    seq_idx, item = bwd_memo[seq_idx]
+    out = [item]
+    while seq_idx >= 0:
+        if verbose: print(out)
+        seq_idx, item = bwd_memo[seq_idx]        
+        out.append(item)
+        if verbose: 
+            print(seq_idx)
+            print(out)
+            print("--")
+    return out[::-1]
+    
+def frontier(memos, seq_idx):
+    bwd_memo, fwd_memo = memos
+    return fwd_memo[seq_idx]
+    
+def prefix_decoding(root_state, state_map):    
+    idx_iter = (i for i in range(10**10))
+    bwd_memo = defaultdict(lambda: next(idx_iter))
+    fwd_memo = defaultdict(lambda: set())
+    memos = (bwd_memo, fwd_memo)
+    final_sequences = _prefix_decoding(root_state, state_map, memos)
+    return final_sequences, memos
+    
+def prefix_step(memos, in_idx, item):
+    bwd_memo, fwd_memo = memos
+    out_idx = bwd_memo[in_idx, item]
+    bwd_memo[out_idx] = (in_idx, item)
+    fwd_memo[in_idx].add(out_idx)
+    return out_idx
+    
+def _prefix_decoding(state, state_map, memos, in_seqs=None):
+    bwd_memo, fwd_memo = memos
+    #print(in_seqs)
[…]

### 2016-10-05 fergus/algorithms/tagger/__init__.py

@@ -1,0 +1,20 @@
+'''
+organization of algorithm:
+    - tagger.Tagger serves as the entry point for the algorithm
+    - tagger.Tagger runs the GPU model
+    - tagger.Tagger can either have multiple methods or be subclasses
+    - tagger_node.* handle the specific node propertis for Fergus-0 and Fergus-R
+    - astar_core.run handles the astar algorithm proper
+    - astar_state.* should implement specific state information for Fergus-0 and Fergus-R
+    -
+'''
+from __future__ import absolute_import
+from .tagger import Tagger
+
+
+def fergus_zero():
+    pass
+
+def fergus_recur():
+    pass
+

### 2016-10-05 fergus/algorithms/tagger/astar_core.py

@@ -1,0 +1,247 @@
+'''
+run a* chart decoder for supertag classification
+
+procedure:
+    - pop highest priority
+    - if complete, find parent set
+    - else, find next daughter set
+    - combine with the set
+    - add results back to the priority queue
+
+priority:
+    - sum of inside-outside values
+    - similar to mike lewis's priority 
+        + if not exactly it
+'''
+from __future__ import absolute_import, print_function, division
+import pickle
+import baal.induce.enrich as ten
+from gist import PATH#, fergus
+#import fergus.algorithms.tree_chooser as tc
+#import fergus.algorithms.tree_chooser_r as tcr
+#from fergus.algorithms import linearizer, driver
+from copy import copy, deepcopy
+import numpy as np
+import heapq
+from collections import defaultdict
+
+
+
+class StarCache(object):
+    def __init__(self, tg_hook):
+        self._cache = dict()
+        self._mapping = dict()
+        self._flip = dict()
+        self._id = 0
+    
+    def __getitem__(self, state_id):
+        return self._cache[state_id][1]
+            
+    def __call__(self, state):
+        _mapping = self._mapping
+        _flip = self._flip
+        _cache = self._cache
+        
+        state_key = state.key
+        score = state.score
+        
+        if state_key in _mapping:
+            # want the min score
+            state_id = _mapping[state_key] 
+            if _cache[state_id][0] >= score:
+                _cache[state_id] = (score, state)
+            return state_id
+        else:
+            state_id = self._id
+            _cache[state_id] = (score, state)
+            _mapping[state_key] = state_id
+            _flip[state_id] = state_key
+            self._id += 1
+            return state_id
+            
+    def solution(self, states=None, state_ids=None, all_solutions=False):
+        states = states or [self[state_id] for state_id in state_ids]
+        states_ = filter(lambda state: state.completed and state.op_pos=='ROOT', 
+                        states)
+        if len(states_) == 0:
+            states_ = filter(lambda state: state.op_pos=='ROOT', states)
+            if len(states_) == 0:
+                return False
+                import pdb
+                pdb.set_trace()
+                print("something wrong with states.. it's 0-length")
+        if all_solutions:
+            return sorted(states_, key=lambda state: state.score)
+        else:
+            return min(states_, key=lambda state: state.score)
+    
+    def __len__(self):
+        return len(self._cache)
+
+class StarQueue(object):
+    def __init__(self, cache):
[…]

### 2016-10-05 fergus/algorithms/tagger/astar_state.py

@@ -1,0 +1,166 @@
+from copy import copy, deepcopy
+
+class BaseState(object):
+    def __init__(self, node, spine_id):
+        self.node_id = node.id
+        self.spine_id = spine_id
+        try:
+            assert spine_id > 1
+        except:
+            import pdb
+            pdb.set_trace()
+        #tree_str = node.ref.make_tree(spine_id, 'foo')
+        self.op_pos, self.op_type = node.ref.igor.spine_lookup[spine_id]
+        #tree_op = node.hypothetical_tree(spine_id).E.tree_operation
+        #self.op_pos = tree_op.target['pos_symbol']
+        #self.op_type = 'ins' if tree_op.is_insertion else 'sub'
+        
+        self.daughter_ids = tuple(d.id for d in node.daughters)
+        self.finished_dots = tuple()
+        self.n_finished = 0
+        self.n_daughters = len(self.daughter_ids)
+        if len(self.daughter_ids) > 0:
+            self.next_dot = self.daughter_ids[0]        
+        else:
+            self.next_dot = None
+        self.sub_valency, self.ins_valency = node.get_valency_options(spine_id, True)
+
+        self.score = node.initial_score(spine_id)
+        self.daughter_scores = node.daughter_scores
+        self.rep_string = None
+        #self.rep_string = self.make_str()
+        self.key = hash((self.node_id, self.op_type, self.op_pos, self.ins_valency, 
+                         self.n_finished, self.next_dot, self.sub_valency))
+    @classmethod
+    def make_all(cls, node, tg):
+        #spineset = node.spineset#tg.igor.get_spineset(node.head_id)
+        return [cls(node, spine) for spine in node.spineset if spine != tg.igor.vocabs.spines.unk_id]
+
+    ### IMPORTANT. Implement this.
+    def step(self, state_id, state, node):
+        ''' returns deep copy of self that's taken one step daughter id'''
+        ## early failure
+        state_op_type = state.op_type
+        state_op_pos = state.op_pos
+        sub_valency = self.sub_valency
+        
+        if not state.completed:
+            return None
+        if state_op_type == 'sub' and state_op_pos not in sub_valency:
+            return None
+        if state_op_type == 'ins' and state_op_pos not in self.ins_valency:
+            return None       
+            
+        daughter_id = state.node_id
+        spine_id = state.spine_id
+        assert daughter_id == self.next_dot
+        cls = self.__class__
+        new_state = cls.__new__(cls)
+        new_state.__dict__.update(self.__dict__)
[…]

### 2016-10-05 fergus/algorithms/tagger/tagger.py

@@ -1,0 +1,210 @@
+from __future__ import absolute_import, print_function, division
+from .utils import decode_node, decode_spine, encode_spine, cartesian_iterator
+from ...models import fergus_neuralized, fergus_recurrent
+from baal.structures import DerivationTree
+from . import astar_core as astar
+from baal.induce import enrich as ten
+from copy import deepcopy
+from . import tagger_node
+from . import astar_state
+import numpy as np
+import sys
+import time
+
+def singleton(cls):
+    cached = {}
+    def getinstance(*args, **kwargs):
+        if cls not in cached:
+            cached[cls] = cls(*args, **kwargs)
+        return cached[cls]
+    return getinstance
+
+class Tagger(object):
+    def __init__(self, path, model):
+        self.model = model
+        self.igor = model.igor
+        self.root = model.Node.from_path(path, self)
+        self.root.annotate()
+        self.nodes = self.root.flat
+        self.runtime = -1
+        
+    @classmethod
+    def fergusn(cls, path):
+        tg = cls(path, MNeuralized())
+        return tg
+
+    @classmethod
+    def fergusr(cls, path):
+        tg = cls(path, MRecurrent())
+        return tg
+
+    def stats(self, raw=False):
+        stats = self.root.full_comparison_stats()
+        if not raw:
+            stats = {k:float(sum(v))/len(v) for k,v in stats.items()}
+        return stats
+    
+    @property
+    def experiment_stats(self):
+        stats = {}
+        stats['before_astar'] = self.classify_only_stats
+        stats['after_astar'] = self.stats(True)        
+        stats['length'] = len(self.root.flat)
+        stats['time'] = self.runtime
+        return stats
+
+    def before_stats(self):
+        for node in self.root.flat.values():
+            assert node.best_index is not None
+            node.set_supertag(node.best_index)
+        self.classify_only_stats = self.stats(True)
+        
+    
+            
+    def run(self, verbose=0):
+        start_time = time.time()
+        self.model.compute_likelihoods(self.root)
+        topside_score = self.root.compute_downside()
+        self.root.cache_upside(topside_score)
+        self.before_stats()
+        astar.run(self)
+        self.model.Node.reset()
+        self.runtime = time.time() - start_time
+
+    def get_partial(self):
[…]

### 2016-10-05 fergus/algorithms/tagger/tagger_node.py

@@ -1,0 +1,276 @@
+from baal.induce import enrich as ten
+from .utils import decode_node, decode_spine, encode_spine, cartesian_iterator
+import numpy as np
+
+try:
+    range = xrange
+except:
+    pass
+
+class Node(object):
+    id_gen = (x for x in range(100000000))
+    def __init__(self, node, ref, parent=None):
+        self.head = node.head
+        if any([n in self.head for n in '1234567890']):
+            self.head = "<NUMBER>"  
+        self.id = next(Node.id_gen)
+        self.ref = ref
+        self.parent = parent
+        self.head_id = self.ref.igor.vocabs.words[self.head]
+        self.head_treelets = []
+        self.child_treelet = -1
+        self.true_bracket = node.bracketed
+
+        keyify = lambda spines: tuple(map(tuple, spines))
+        encoded_spine = encode_spine(node.get_spine(), self.ref.igor)
+        self.spine_id = self.ref.igor.vocabs.spines[keyify(encoded_spine)]
+        self.selected_tag = None
+
+        self.spineset = [spine_id for spine_id in list(self.ref.igor.get_spineset(self.head_id)) 
+                                  if spine_id != self.ref.igor.vocabs.spines.unk_id]
+        self.spine_lookup = {s:i for i, s in enumerate(self.spineset)}
+
+
+    @classmethod
+    def from_derivation(cls, tree, ref, parent=None):
+        node = cls(tree, ref, parent)
+        node.daughters = [cls.from_derivation(child, ref, node) for child in tree.children]
+        node.subtree_ids = set([node.id] + [d.id for d in node.daughters])
+        node.daughter_lookup = {d.id:i for i, d in enumerate(node.daughters)}
+        return node
+
+    @classmethod
+    def from_path(cls, path, ref):
+        cls.reset()
+        tree = ten.rollout(path)
+        return cls.from_derivation(tree, ref)
+    
+    @staticmethod
+    def reset():
+        Node.id_gen = (x for x in range(1000000))
+    
+    @property
+    def count(self):
+        return 1 + sum([d.count for d in self.daughters])
+    
+    @property
+    def treelets(self):
+        tlets = [(self, d) for d in self.daughters]
+        for d in self.daughters:
+            tlets.extend(d.treelets)
+        return tlets
+    
+    @property
+    def leaves(self):
+        leaves = []
+        if len(self.daughters) == 0:
+            leaves.append(self)
+        else:
+            for d in self.daughters:
[…]

### 2016-10-05 fergus/algorithms/tagger/utils.py

@@ -1,0 +1,324 @@
+from __future__ import print_function, division
+from baal.structures import DerivationTree
+from copy import deepcopy
+from collections import OrderedDict, defaultdict
+import itertools
+import heapq
+from copy import copy, deepcopy
+
+
+def _limit(i, n=500):
+    for j in range(100):
+        if j**i > n:
+            return j-1
+
+
+def cartesian_iterator(n_nums, lazy=True):
+    if lazy:
+        n_nums = [len(n) for n in n_nums]
+    max_n = _limit(len(n_nums))
+    duosorted = sorted([(min(n,max_n),i) for i,n in enumerate(n_nums)], reverse=True)
+    
+    unsort = {orig_idx:new_idx for new_idx, (_, orig_idx) in zip(xrange(len(n_nums)), duosorted)}
+    def f_unsort(indices):
+        return [indices[unsort[i]] for i in range(len(n_nums))]
+
+    range_iters = tuple([range(l) for l, _ in duosorted])
+    all_idx = sorted(map(f_unsort, list(itertools.product(*range_iters))), key=lambda x: (sum(x), x))
+    biggest = max(map(sum, all_idx))
+    idx_by_sum = {k:[x for x in all_idx if sum(x)==k] for k in range(1, biggest)}
+    for i in range(1,max_n):
+        yield idx_by_sum[i]
+
+def decode_node(igor, node):
+    p,t = node
+    p = igor.vocabs.pos.lookup(p)
+    t = igor.vocabs.type.lookup(t)
+    return "{}/{}".format(p,t)
+
+def decode_spine(igor, spine_idx=None, spine=None):
+    if spine_idx is not None:
+        spine = igor.vocabs.spines.lookup(spine_idx)
+    elif spine is None:
+        raise Exception("crap")
+
+    return "|"+"|-->|".join(["+".join(map(lambda x: decode_node(igor,x), context)) 
+                                    for context in spine])+"|"
+
+def encode_spine(spine, igor):
+    encoded_spine = []
+    for context in spine:
+        encoded_nodeset = []
+        for pos, spine_info in context:
+            encoded_node = (igor.vocabs.pos.add(pos), 
+                            igor.vocabs.type.add(spine_info.upper()))
+            encoded_nodeset.append(encoded_node)
+        encoded_spine.append(encoded_nodeset)
+    return encoded_spine
+
+def make(node):
+    treestr, selfhlf, targethlf, targetgorn = node
[…]

### 2016-10-05 fergus/configs/__init__.py

@@ -1,0 +1,23 @@
+import yaml
+import os
+import sys
+PATH = os.path.dirname(os.path.realpath(__file__))
+
+
+def get_config(filename):
+    filename = os.path.join(PATH, filename)
+    config = {}
+
+    with open(filename) as fp:
+        for key, value in yaml.load(fp.read()).items():
+            if "dir_rel" in key:
+                config[key.replace("_rel","")] = os.path.join(PATH, value)
+            elif "dir_abs" in key:
+                config[key.replace("_abs","")] = value
+            else:
+                config[key] = value
+    return config
+        
+def global_config(filename='global_config.conf'):
+    return get_config(filename)
+    

### 2016-10-05 fergus/data_pipeline/1_parse/final_fixer.py

@@ -1,0 +1,23 @@
+import glob
+
+for file in glob.glob("preprocessed/*.mrg"):
+    trees = []
+    mode = "pass"
+    bucket = ""
+    print("starting ", file)
+    with open(file) as fp: 
+
+        for line in fp.readlines():
+            if "Original tree" in line:
+                mode = "pass"
+            elif "Tree transformed" in line:
+                mode = "accum"
+            elif "-------------" in line:
+                trees.append(bucket)
+                bucket = ""
+            elif mode == "accum":
+                bucket += line
+    newf = file.replace("preprocessed", "final")
+    with open(newf, "w") as fp:
+        for t in trees:
+            fp.write(t)

### 2016-10-05 fergus/data_pipeline/1_parse/merge_sections.py

@@ -1,0 +1,10 @@
+import glob
+
+for folder in glob.glob("*/"):
+    if not folder.isdigit(): continue
+    combined = ""
+    for file in glob.glob(folder+"/*.mrg"):
+        with open(file) as fp:
+            combined += "\n" + fp.read()
+    with open("merged/"+folder.replace("/",".mrg"), "w") as fp:
+        fp.write(combined)

### 2016-10-05 fergus/data_pipeline/1_parse/run_stanford_coordination.py

@@ -1,0 +1,11 @@
+import subprocess
+
+command = "~/research/code/stanford-parser/bcm_transformer.sh merged/{0} > sfprocced/{0}"
+
+import glob
+
+for file in glob.glob("merged/*.mrg"):
+    new_c = command.format(file.replace("merged/",""))
+    print("running.. {}".format(new_c))
+    subprocess.call(new_c, shell=True)
+

### 2016-10-05 fergus/data_pipeline/2_grammar_induction/wsj2paths.py

@@ -1,0 +1,104 @@
+"""Process the WSJ treebank to paths and possible surface forms
+
+Usage:
+    wsj2paths.py convert <config_file> [-s]
+    wsj2paths.py (-h | --help)
+
+    convert <config_file>         all necessary configuration should be in here
+                                  an example of the config is below. 
+    -s --surface                  save the surface form from the paths
+    -h --help                     show this
+
+
+
+Example (Cohn, Chiang, etc) Config:
+---
+wsj_root_dir: /research/data/LDC/treebank3/treebank_3/parsed/mrg/wsj/
+splits: 
+  - 02-21
+  - 22-22
+  - 23-23
+save_name: chiang_wsj
+---
+
+Example (Mikolov, etc) Config:
+---
+wsj_root_dir: /research/data/LDC/treebank3/treebank_3/parsed/mrg/wsj/
+splits: 
+  - 00-20
+  - 21-22
+  - 23-24
+save_name: mikolov_wsj
+---           
+"""
+from __future__ import print_function
+from collections import defaultdict, Counter
+from os.path import join
+from docopt import docopt
+from baal.nlp.induce.tree_enrichment import get_trees, string2cuts
+from baal.science.gist import common
+from baal.nlp.structures import DerivationTree, ElementaryTree
+import baal
+import pickle
+import yaml
+from tqdm import tqdm, trange
+import sys
+import glob
+
+from ..common import config
+
+def parse_splits(config):
+    splits = config.splits
+    _F = lambda i: "{:02}".format(i)
+    F = lambda l,u: [_F(i) for i in range(l,u)]
+    splits = [x.split("-") for x in splits]
+    splits = [F(int(l), int(u)+1) for l,u in splits]
+    return {k:v for k,v in zip(["train", "dev", "test"],splits)}
+
+def safe_single_parse(parse):
+    try:
+        _, sf = common.parse2predicates(parse)
+        tree, subtrees = baal.nlp.induce.tree_enrichment.parse2derivation(parse)
+        assert sf == str(tree).replace("_", " ")
+        return True, (tree.pre_order_features(), sf)
+    except Exception as e:
+        return False, (parse, e)
+
+def run():
+    splits = parse_splits(config)
+    base_dir = os.path.join(config.wsj_root_dir, "final/[0-9]*.mrg")
+    files = {'weird': "weird_stuff_{}.txt".format(config['save_name']),
+             'good_save': config['save_name']+'_{}.pkl',
[…]

### 2016-10-05 fergus/data_pipeline/3_generate_datasets/fergus_n.py

@@ -1,0 +1,23 @@
+import yaml
+from tqdm import tqdm
+
+
+def make():
+    config = Blob("../pipeline_settings.conf")
+    data_files = ["{}_{}.pkl".format(config.data_prefix, split) 
+                    for split in ("train", "dev", "test")]
+
+    for data_file in tqdm(data_files, desc='data file', position=0):
+        with open(join(config.processed_data_dir, data_file)) as fp:
+            data = pickle.load(fp)
+        dataset = []
+        data_name = data_file.replace(".pkl","")
+        for datum in tqdm(data, desc="{} data".format(data_name), position=1):
+            full_tree = rollout(datum)
+            tree_features = extract_features(full_tree)
+            ## note: this needs to be deprecated by fixing this in the induction
+            #roll_features = [(tempfix(a), tempfix(b), c, d, e) 
+            #                    for a,b,c,d,e in full_tree.modo_roll_features()]
+            dataset.extend(tree_features)
+        with open(join(data_fp, data_file.replace(".pkl", "_fergus_N_supertags.pkl")), "w") as fp:
+            pickle.dump(dataset, fp)

### 2016-10-05 fergus/data_pipeline/common.py

@@ -1,0 +1,17 @@
+import yaml
+
+class Blob(object):
+    def __init__(self, conf=None, conf_file=None):
+        if conf is None and conf_file is not None:
+            with open(conf_file) as fp:
+                conf = yaml.load(fp)
+        elif conf is None:
+            raise Exception("Incorrect usage of config container")
+
+        self.__dict__.update(conf.items())
+
+
+config = Blob("pipeline_settings.conf")
+
+
+

### 2016-10-05 fergus/models/__init__.py

@@ -1,0 +1,3 @@
+from __future__ import absolute_import
+from . import fergus_neuralized
+from . import fergus_recurrent

### 2016-10-05 fergus/models/common/__init__.py

@@ -1,0 +1,78 @@
+from ikelos.layers import DynamicEmbedding, LambdaMask
+import keras.backend as K
+from keras.layers import Convolution2D, Lambda, Flatten, DataLayer, Embedding
+from keras.regularizers import l2
+from keras.engine import merge
+import os
+concat = lambda layers: merge(layers, mode='concat')
+
+def make_convolutional_embedding(igor):
+    p_emb = igor.p_emb_dropout
+    p_W = igor.p_W_dropout
+    p_U = igor.p_U_dropout
+    w_decay = igor.weight_decay
+    p_mlp = igor.p_mlp_dropout
+    p_type = p_emb * igor.total_num_type / igor.total_num_pos # weight by the ratio to even things out
+    Z_spinepos_layer = DataLayer(igor.X_spinepos, input_dtype='int32', name='spine_pos_lexicon')
+    Z_spinetype_layer = DataLayer(igor.X_spinetype, input_dtype='int32', name='spine_type_lexicon')
+
+    Z_spinepos = Z_spinepos_layer.tensor
+    Z_spinetype = Z_spinetype_layer.tensor
+    igor.preloaded_data = [Z_spinepos, Z_spinetype]
+
+    F_embedpos = Embedding(igor.total_num_pos, igor.pos_embedding_size, mask_zero=True,
+                           W_regularizer=l2(w_decay), dropout=p_emb, name='embed_pos')
+    F_embedtype = Embedding(igor.total_num_type, igor.type_embedding_size, mask_zero=True,
+                           W_regularizer=l2(w_decay), dropout=p_type, name='embed_type')
+
+    F_stripmask = LambdaMask(lambda x, mask: None, name='strip_mask')
+    spine_pos = F_stripmask(F_embedpos(Z_spinepos))
+    spine_type = F_stripmask(F_embedtype(Z_spinetype))
+
+    L = igor.spine_lexicon_size
+    Sp = igor.max_spine_length
+    C = igor.max_context_size
+    poty = igor.pos_embedding_size + igor.type_embedding_size
+
+    spine_cat = concat([spine_pos, spine_type])
+    spine_cat = Lambda(lambda xin: K.permute_dimensions(xin, (0,3,1,2)), output_shape=lambda *args:(L, poty, Sp, C))(spine_cat)
+    spine_conv1 = Convolution2D(igor.spine_convsame_filters, 1, 2, border_mode='valid')(spine_cat) 
[…]

### 2016-10-05 fergus/models/fergus_neuralized/__init__.py

@@ -1,0 +1,33 @@
+from __future__ import absolute_import
+import os
+from . import model
+from ...configs import global_config, get_config
+
+PATH = os.path.dirname(os.path.realpath(__file__))
+
+def TrainingModel():
+    print("Retrieving training model")
+    config = global_config()
+    config_file = config['fergus_n_train_config']
+    config_file = os.path.join(config['config_dir'], config_file)
+    return model.FergusNModel.from_yaml(config_file)
+    
+def DevelopmentModel():
+    print("Retrieving development model")
+    config = global_config()
+    config_file = config['fergus_n_dev_config']
+    config_file = os.path.join(config['config_dir'], config_file)
+    return model.FergusNModel.from_yaml(config_file)
+
+def TestingModel():
+    print("Retrieving testing model")
+    config = global_config()
+    config_file = config['fergus_n_test_config']
+    config_file = os.path.join(config['config_dir'], config_file)
+    return model.FergusNModel.from_yaml(config_file)
+
+def globally_set_model():
+    config = global_config()
+    return {'dev': DevelopmentModel,
+            'test': TestingModel,
+            'train': TrainingModel}[config['tag_mode']]()

### 2016-10-05 fergus/models/fergus_neuralized/igor.py

@@ -1,0 +1,393 @@
+from __future__ import absolute_import, print_function, division
+
+from collections import Counter, defaultdict
+from tqdm import tqdm
+import numpy as np
+import itertools
+import os
+from os.path import join
+
+from baal.structures import DerivationTree
+from keras.layers import Embedding, LSTM, Input
+from keras.utils.np_utils import to_categorical
+from ikelos.data import VocabManager, Vocabulary, DataServer
+
+try:
+    import cPickle as pickle
+except:
+    import pickle
+
+from . import utils
+from ...configs import global_config
+
+class Igor(DataServer):
+    """Handle the data for FERGUS 
+
+    Learning Problem 1.
+        - Each datapoint has: mother_head, daughter_heads, mother_supertag (e-tree)
+        - Given the dataset, we can collect all supertags for a given head
+        - Now, each datapoint has: mother_head, possible_supertags, daughter_heads, target_supertag
+        - So, form the dataset as:
+            X[bth_batch, ith_supertag][:spine_length]=spine_context_matrix
+            Y[bth_batch] = to_categorical(truth_index, max_possible_supertags)
+
+        Note: 
+            + at run time, we can pre-compute all of the spine
+    """ 
+    
+    def verify_directories(self):
+        d1 = os.path.join(self.model_location, self.saving_prefix)
+        d = os.path.dirname(d1)
+        if not os.path.exists(d):
+            os.makedirs(d)
+        
+    def format_single(self, daughter_head, parent_head, daughter_spine, parent_spine):
+        '''Current formulation:
+            quadruplet of (daughter head, parent head, daughter spine, parent spine)
+                       are observed. there are two ways we can treat the classification
+                       so, I'm doing them both. 
+            1. classify to correct daughter spine
+            2. classify to correct parent spine
+
+            in each case, we need the total spine set for the hypotheses
+            we also need the index of the correct one. for simplicity, just going to pass in 2 extra vectors.
+        '''
+        N_t = self.max_num_spines
+        X_dspineset = np.zeros((N_t,))
+        X_pspineset = np.zeros((N_t,))
+        X_dspine = np.array([daughter_spine])
+        X_pspine = np.array([parent_spine])
+        X_dhead = np.array([daughter_head])
+        X_phead = np.array([parent_head])
+        Y_dspine = np.array([-1]) 
+        Y_pspine = np.array([-1])
+
[…]

### 2016-10-05 fergus/models/fergus_neuralized/model.py

@@ -1,0 +1,302 @@
+from __future__ import absolute_import, print_function, division
+
+import baal
+import theano
+import time
+import json
+import numpy as np
+from os.path import join, exists
+
+
+#### need ikelos library; see braingineer/ikelos
+from ikelos.data import Vocabulary, VocabManager
+from ikelos.layers import ProbabilityTensor, SoftAttention, Summarize, Fix, LambdaMask, \
+                          set_name, compose, LastDimDistribute, DynamicEmbedding
+
+#### need my version of keras; see braingineer/keras
+import keras.backend as K
+from keras.layers import Embedding, LSTM, Input, Lambda, \
+                         Distribute, RepeatVector, \
+                         Reshape, Flatten, Dense, Dropout, \
+                         InputLayer, Convolution2D, MaxPooling2D
+from keras.engine import Model, merge
+from keras.optimizers import Adam
+from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, LearningRateScheduler
+from keras.regularizers import l2
+from keras.utils.visualize_util import plot as kplot
+
+### project imports
+from .igor import Igor
+from ..common import make_convolutional_embedding, make_logger
+
+
+try:
+    import cPickle as pickle
+except:
+    import pickle
+import sys;
+sys.setrecursionlimit(40000)
+
+concat = lambda layers: merge(layers, mode='concat')
+merge_batch = lambda layers: merge(layers, mode='concat', concat_axis=0)
+
+
+def traverse_nodes(layer, from_name="top", down=True):
+    #print("- At {} from {}".format(layer.name, from_name))
+    nodes = layer.inbound_nodes
+    #print("\t+{} inbound nodes".format(len(nodes)))
+    mask_cache = {}
+    for node in layer.inbound_nodes:
+        #print("\t + {} layers feeding into node".format(len(node.inbound_layers)))
+        #print("\t + {} shapes feeding into node".format("; ".join(map(str, node.input_shapes))))
+        #print("\t + ({}) masks feeding into node".format("; ".join(["yes" if mask is not None else "no" 
+        #                                                          for mask in node.input_masks])))
+        for i, mask in enumerate(node.input_masks):
+            if mask is not None:
+                mask_cache['{}.in.mask.{}'.format(layer.name, i)] = mask
+        for layer_down in node.inbound_layers:
+            mask_cache.update(traverse_nodes(layer_down, layer.name))
+    return mask_cache
+
+class FergusNModel(object):
+    def __init__(self, igor):
[…]

### 2016-10-05 fergus/models/fergus_neuralized/utils.py

@@ -1,0 +1,183 @@
+from __future__ import print_function
+from os.path import join
+import random
+import baal
+from baal.utils import loggers
+from baal.structures import DerivationTree
+from ikelos.data import Vocabulary, VocabManager
+from keras.layers import Embedding, LSTM, Input
+from tqdm import tqdm
+from collections import Counter
+try:
+    import cPickle as pickle
+except:
+    import pickle
+
+def make(node):
+    treestr, selfhlf, targethlf, targetgorn = node
+    tree = DerivationTree.from_bracketed(treestr)
+    tree.set_path_features(self_hlf=selfhlf)
+    if targethlf:
+        tree.set_path_features(target_hlf=targethlf, target_gorn=targetgorn)
+    return tree
+
+def rollout(datum):    
+    out = None
+    for node in datum:
+        dnode  = make(node)
+
+        if any([char in dnode.E.head for char in '1234567890']):
+            dnode.E.head = "<NUMBER>"
+        if out is None:
+            out = dnode
+        else:
+            out = out.operate(dnode)
+            
+    return out
+
+def tempfix(s):
+    if any([n in s for n in '1234567890']):
+        #print(s, "replaced by <NUMBER>")
+        #raw_input()
+        return "<NUMBER>"
+    return s
+
+def create_dataset():
+    """
+    Target data:
+        [(head, spines paths w/ context, children_set), ...]
+        s.t. 
+        spine_path w/ context will represent the supertag
+        a supertag := an elementary tree
+        supertag_set := a list of the trees for head word
+        children_set := the list of children for that head word
+    """
+    gist_fp = join(baal.PATH, "science/gist")
+    data_fp = join(gist_fp, "data")
+    data_files = ["chiang_wsj_train.pkl", "chiang_wsj_dev.pkl", "chiang_wsj_test.pkl"]
+    for data_file in tqdm(data_files, desc='data file', position=0):
+        with open(join(data_fp, data_file)) as fp:
+            data = pickle.load(fp)
+        dataset = []
+        for datum in tqdm(data, desc="{} data".format(data_file.replace(".pkl","")),
+                          position=1):
+            #assume path)
+            full_tree = rollout(datum)
+            roll_features = [(tempfix(a), tempfix(b), c, d, e) for a,b,c,d,e in full_tree.modo_roll_features()]
+            dataset.extend(roll_features)
[…]

### 2016-10-05 fergus/models/fergus_recurrent/__init__.py

@@ -1,0 +1,34 @@
+from __future__ import absolute_import
+import os
+from . import model
+from ...configs import global_config, get_config
+
+PATH = os.path.dirname(os.path.realpath(__file__))
+
+def TrainingModel():
+    print("Retrieving training model")
+    config = global_config()
+    config_file = config['fergus_r_train_config']
+    config_file = os.path.join(config['config_dir'], config_file)
+    return model.FergusRModel.from_yaml(config_file)
+    
+    
+def DevelopmentModel():
+    print("Retrieving development model")
+    config = global_config()
+    config_file = config['fergus_r_dev_config']
+    config_file = os.path.join(config['config_dir'], config_file)
+    return model.FergusRModel.from_yaml(config_file)
+
+def TestingModel():
+    print("Retrieving testing model")
+    config = global_config()
+    config_file = config['fergus_r_test_config']
+    config_file = os.path.join(config['config_dir'], config_file)
+    return model.FergusRModel.from_yaml(config_file)
+    
+def globally_set_model():
+    config = global_config()
+    return {'dev': DevelopmentModel,
+            'test': TestingModel,
+            'train': TrainingModel}[config['tag_mode']]()

### 2016-10-05 fergus/models/fergus_recurrent/igor.py

@@ -1,0 +1,425 @@
+from __future__ import absolute_import, print_function, division
+
+from collections import Counter, defaultdict
+from tqdm import tqdm
+import numpy as np
+import itertools
+import os
+from os.path import join
+
+
+from baal.structures import DerivationTree
+from keras.layers import Embedding, LSTM, Input
+from keras.utils.np_utils import to_categorical
+from ikelos.data import VocabManager, Vocabulary, DataServer
+
+try:
+    import cPickle as pickle
+except:
+    import pickle
+
+from . import utils
+from ...configs import global_config
+
+class Igor(DataServer):
+    """Handle the data for FERGUS-R
+
+    """ 
+        
+    def verify_directories(self):
+        d1 = os.path.join(self.model_location, self.saving_prefix)
+        d = os.path.dirname(d1)
+        if not os.path.exists(d):
+            os.makedirs(d)
+        
+
+    def prep_date(self, tree_list):
+        stk = [tree_list]
+        root_head, root_spine = tree_list[0]
+        out = [(-1, self.vocabs.words[root_head], self.vocabs.spines[root_spine])]
+        while stk:
+            top, rest = stk
+            for d in rest:
+                out.append((p_idx, head_idx, spine_idx))
+                stk.append(d)
+
+
+
+    def format_single(self, data):
+        '''Need:
+            - format each item in the tree set
+            - not necessary a sequence, but we are treating the formation as it
+
+        '''
+        N_spine = self.max_num_spines
+        N_seq = self.max_sequence
+        N_dctx = self.max_daughter_size
+
+        X_dcontext = np.zeros((N_seq, N_dctx))
+        X_spineset = np.zeros((N_seq, N_spine))
+        X_head = np.zeros((N_seq))
+        X_topo = np.zeros((N_seq))
+        X_gpspine = np.zeros((N_seq))
+        Y_spine = np.zeros((N_seq, N_spine))
+
+        y_idx = []
+        gpspines = []
+        for n_i, (p_idx, head_idx, spine_idx, dot_idxs) in enumerate(data):
+            X_head[n_i] = head_idx
+            X_topo[n_i] = p_idx
+            X_dcontext[n_i][:len(dot_idxs)] = dot_idxs
+            if n_i == 0:
+                X_gpspine[n_i] = 2
+            else:
+                X_gpspine[n_i] = gpspines[p_idx]
+            gpspines.append(spine_idx)
[…]

### 2016-10-05 fergus/models/fergus_recurrent/model.py

@@ -1,0 +1,249 @@
+from __future__ import absolute_import, print_function, division
+
+from os.path import join, exists
+from tqdm import tqdm
+import time
+import json
+import numpy as np
+import sys
+
+import theano
+
+import keras.backend as K
+from keras.engine import Model, merge
+from keras.layers import Embedding, LSTM, Input, Lambda, \
+                         Distribute, RepeatVector, \
+                         Reshape, Flatten, Dense, Dropout, \
+                         InputLayer, DataLayer, Convolution2D, MaxPooling2D
+from keras.optimizers import Adam
+from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, LearningRateScheduler
+from keras.regularizers import l2
+from keras.utils.visualize_util import plot as kplot
+
+from ikelos.layers import ProbabilityTensor, SoftAttention, Summarize, Fix, LambdaMask, set_name, compose, \
+                          BranchLSTM, LastDimDistribute, DynamicEmbedding
+from ikelos.data import Vocabulary, VocabManager
+
+
+from .igor import Igor
+from ..common import make_convolutional_embedding, make_logger
+from ...configs import global_config
+
+try:
+    import cPickle as pickle
+except:
+    import pickle
+
+sys.setrecursionlimit(40000)
+
+concat = lambda layers: merge(layers, mode='concat')
+merge_batch = lambda layers: merge(layers, mode='concat', concat_axis=0)
+
+class FergusRModel(object):
+    def __init__(self, igor):
+        config = global_config()
+        self.logger = igor.logger = make_logger(igor, 'fergusr_'+config['tag_mode'])
+        self.igor = igor
+        self.igor.verify_directories()
+        self.model = None
+        
+    @classmethod
+    def from_yaml(cls, yamlfile, kwargs=None):
+        igor = Igor.from_file(yamlfile)
+        igor.prep()
+        model = cls(igor)
+        model.make(kwargs)
+        return model
+
+    def load_checkpoint_weights(self):
+        weight_file = join(self.igor.model_location, 
+                           self.igor.saving_prefix,
+                           self.igor.checkpoint_weights)
+        if exists(weight_file):
+            self.logger.info("+ Loading checkpoint weights")
+            self.model.load_weights(weight_file, by_name=True)
+        else:
+            self.logger.warning("- Checkpoint weights do not exist; {}".format(weight_file))
+
+
+    def plot(self):
+        filename = join(self.igor.model_location, 
[…]

### 2016-10-05 fergus/models/fergus_recurrent/utils.py

@@ -1,0 +1,207 @@
+from __future__ import print_function
+from os.path import join
+import random
+import gist
+from baal.utils import loggers
+from baal.structures import DerivationTree
+from ikelos.data import Vocabulary, VocabManager
+from keras.layers import Embedding, LSTM, Input
+from tqdm import tqdm
+from collections import Counter
+try:
+    import cPickle as pickle
+except:
+    import pickle
+
+def make(node):
+    treestr, selfhlf, targethlf, targetgorn = node
+    tree = DerivationTree.from_bracketed(treestr)
+    tree.set_path_features(self_hlf=selfhlf)
+    if targethlf:
+        tree.set_path_features(target_hlf=targethlf, target_gorn=targetgorn)
+    return tree
+
+def rollout(datum):    
+    out = None
+    for node in datum:
+        dnode  = make(node)
+
+        if any([char in dnode.E.head for char in '1234567890']):
+            dnode.E.head = "<NUMBER>"
+        if out is None:
+            out = dnode
+        else:
+            out = out.operate(dnode)
+            
+    return out
+
+def remove_numbers(s):
+    if any([n in s for n in '1234567890']):
+        #print(s, "replaced by <NUMBER>")
+        #raw_input()
+        return "<NUMBER>"
+    return s
+
+def make_filters(igor):
+    """Filter the data with the data_frequency_cutoff specified in the conf
+
+    - Count on train.  Filter on rest.
+    - Since we are filtering based on classification target, remove unseen classes
+      in test and dev as well.  
+    - Report the number of removed supertags, and the reduction in sizes of train, test, and dev sets
+    """
+    counts = Counter([bracket for sentence in igor.train_data for _,_,_,_,bracket in sentence])
+    igor.good_filter = set([ex for ex,count in counts.items() if count>igor.data_frequency_cutoff])
+    igor.bad_filter = set([ex for ex,count in counts.items() if count<=igor.data_frequency_cutoff])
+    igor.total_good = sum([counts[ex] for ex in igor.good_filter])
+    igor.total_bad = sum([counts[ex] for ex in igor.bad_filter])
+
+
+
+def load_dataset(filepath):
+    with open(filepath) as fp:
+        dataset = pickle.load(fp)
+    return dataset
+
+def process_data(igor):
+    igor.head2spine = {}
+    make_filters(igor)
+
+    igor.logger.info("Cutoff Frequency={}.  {} supertags retained; {} supertags cut.".format(igor.data_frequency_cutoff,
[…]

### 2016-10-05 runners/run_tagger.py

@@ -1,0 +1,86 @@
+from __future__ import print_function, division, absolute_import
+import pickle
+import sys
+import os
+here = os.path.dirname(os.path.realpath(__file__))
+sys.path.append(os.path.join(here, '..'))
+import fergus
+
+### additional packages needed
+from tqdm import tqdm
+from sqlitedict import SqliteDict
+import time
+### code in this repo
+from fergus.algorithms.tagger.utils import rollout
+from fergus.configs import global_config
+from fergus.algorithms.tagger import Tagger
+
+def run(model, config):
+
+    print("loading data")
+    filename_key = "{}_filename".format(config['tag_mode'])
+    filepath = os.path.join(config['data_dir'], config[filename_key])
+    
+    with open(filepath, 'rb') as fp:
+        data = pickle.load(fp)
+    print("data loaded")
+    
+    if config['fast_mode']: 
+        import numpy as np
+        data = [data[i] for i in np.random.random_integers(0, len(data), config['fast_mode'])]
+
+    all_stats = []
+    with SqliteDict(config['data_store_db'], tablename='indata') as datadb:
+        capacitor = {}
+        dbkeys = set(map(int,datadb.keys()))
+        if len(dbkeys)>0:
+            print("{} already done; overwrite={}".format(len(dbkeys), config['overwrite_store']))
+        skipped = 0
+        for i, datum in tqdm(enumerate(data), total=len(data)):
+            if i in dbkeys and not config['overwrite_store']: continue
+
+            if len(datum) > 100: 
+                skipped += 1
+                continue
+
+            tg = model(datum)
+            if len(tg.nodes) == 1:
+                print("Degenerate example (#{}). Skipping.".format(i))
+                continue
+            start = time.time()
+            try:
+                tg.run()
+                capacitor[i] = (tg.root.to_list(), str(rollout(datum)), tg.experiment_stats)
+            except Exception as e:
+                print("EXCEPTION: {}".format(e))
+                print("sys.exc_info(): {}".format(sys.exc_info()))
+            
+            if config['verbose']:
+                for k, v in sorted(tg.stats().items()):
+                    print("{:<25} =  {:>0.3f}".format(k, v))
+                    
+            if (i+1) % 20 == 0:
+                datadb.update(capacitor)
+                datadb.commit()
+                capacitor = {}
+                
+        if len(capacitor) > 0:
+            datadb.update(capacitor)
[…]

### 2016-10-05 runners/run_trainer.py

@@ -1,0 +1,84 @@
+"""An implemenation of FERGUS's supertagger
+
+Usage:
+    run_trainer.py fergusr
+    run_trainer.py fergusn
+    run_trainer.py (-h | --help)
+
+    fergusr                     train the fergus-r model
+    fergusn                     train the fergus-n model
+
+
+Notes: 
+as per Bangalore & Rambow, 2000 
+"Exploiting a Probabilistic Hierarchical Model for Generation"
+
+
+
+"""
+import os
+import sys
+here = os.path.dirname(os.path.realpath(__file__))
+sys.path.append(os.path.join(here, '..'))
+import fergus
+
+from ikelos.data import Vocabulary
+from fergus.models import fergus_recurrent, fergus_neuralized
+from fergus.configs import global_config, get_config
+from os.path import join
+from docopt import docopt
+import yaml
+import json
+
+import theano
+from theano.compile.nanguardmode import NanGuardMode
+
+
+if __name__ == "__main__":
+
+    args = docopt(__doc__, version='FERGUS trainer; October 2016')
+
+    try:
+        if args['fergusr']:
+            model = fergus_recurrent.TrainingModel()
+        elif args['fergusn']:
+            model = fergus_neuralized.TrainingModel()
+        else:
+            print("No other options currently")
+            sys.exit(0)
+        model.train()
+
+    except KeyboardInterrupt as e:
+        print("Exceptional Keyboard interruption... saving model now")
+        model.model.save("model_safetynet.h5")
+    except Exception as e:
+        print("Unknown exception: {}".format(e))
+        print("Some more info: {}".format(sys.exc_info()))
+        print("Saving model just in case")
+        model.model.save("model_extrasafetynet.h5")
+          
+        # elif args['debug']:
+        #     kwargs = {"mode":NanGuardMode(nan_is_error=True, 
+        #                                   inf_is_error=True, 
+        #                                   big_is_error=True)}
+        #     #stage(igor)
+        #     #model = TrainingModel(igor)
+        #     #model.make(kwargs)
+        #     #model.debug()
+        #     #model.train()
+
+        #     model = TrainingModel.from_yaml(args['<config_file>'], kwargs)
+        #     model.train()
+
+        # elif args['profile']:
+        #     kwargs = {"profile": True}
+        #     stage(igor)
+        #     model = TrainingModel(igor)
+        #     model.make(kwargs)
+        #     model.profile(num_iterations=1)
+
[…]

### 2016-10-05 paper/draft.tex
syncing

@@
 %
-The evaluation uses the Wall Street Journal sections of the Penn Treebank which have been previously used for statistical tree grammars \cite{chiang2000statistical}\footnote{A possible additional data source, the data from the 2011 Shared Task on Surface Realization, was not available}.
+The evaluation uses the Wall Street Journal sections of the Penn Treebank which have been previously used for evaluating statistical tree grammars \cite{chiang2000statistical}\footnote{A possible additional data source, the data from the 2011 Shared Task on Surface Realization, was not available}.
 %
@@
 %
-Abstracting syntactic information from the derivation tree leads to the unlabeled dependency trees our models assume as input.
+Abstracting syntactic information from the derivation tree structure leads to the unlabeled dependency trees our models assume as input.
 
@@
 
+Table \ref{table:accresults} shows the results of supertag classification using the stochastic tree models is shown. 
+%
+The scores are grouped in first by the model and then by the embedding method. 
+%
+The accuracy of each model is additionally broken down by how often it got the various components correct. 
+%
+The root category and role refer to the syntactic label and grammatical role of a supertag. 
+%
+The leaf category refers to the syntactic label of the leaf to which a lexical item would attach---essentially, this is the part of speech of the lexical item.
+%
+All differences between model are significant using a Paired-Sample t-test ($p<10^{5}$)
+
+The Fergus-N model out-performed the Fergus-R model on supertag classification with convolutional supertag embeddings, but the Fergus-R model performed much closer to that of Fergus-N on the individual components.
+
 \begin{figure}
 \centering
-\begin{tabular}{|l|p{2cm}|p{1.5cm}|}
+\begin{tabular}{|l|p{2cm}|p{2cm}|p{2.3cm}|p{2cm}|p{2.3cm}|}
 \hline
-Model & Test Set Size & Accuracy \\ \hline
-Fergus-Penn & 100 & 0.749 \\
-\hline \hline
-Fergus-N $Top_{40}$  & 2187 & 0.699 \\
+Model & Embedding & Supertag & Root Category & Root Role & Leaf Category  \\ \hline
+Fergus-N &  Convolution &  57.4\% & 81.8\% & 81.0\% & 86.8\% \\ \cline{2-6}
+         &  Token       &  XX & XX & XX & \\ \cline{2-6}
+\hline
+Fergus-R &  Convolution &  36.4\% & 76.9\% & 78.3\% & 80.8\% \\ \cline{2-6}
+         &  Token       &  XX & XX & XX & \\ \cline{2-6}
 \hline
 \end{tabular}
-\label{table:results}
-\caption{ }
+\label{table:accresults}
+\caption{Classification accuracy by supertag and its individual properties}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tabular}{|l|p{2cm}|p{1.5cm}|r|}
+\hline
+Model & Embedding & Beam Type & Accuracy \\ \hline
+FERGUS & n.a. & n.a. & 74.9 \% \\
+\hline \hline
+Fergus-N & Convolution & $Top_{all}$  & 68.9\% \\ \cline{3-4}
+         &             & $Best_{all}$ & 76.4\% \\ \cline{2-4}
+         & Token       & $Top_{all}$  & XX\% \\ \cline{3-4}
+         &             & $Best_{all}$ & XX\% \\ \cline{3-4}
+\hline
+Fergus-R & Convolution & $Top_{all}$  & 71.9\% \\ \cline{3-4}
+         &             & $Best_{all}$ & 78.9\% \\ \cline{2-4}
+         & Token       & $Top_{all}$  & XX \% \\ \cline{3-4}
+         &             & $Best_{all}$ & XX \% \\ \cline{3-4}
+\hline
+\end{tabular}
+\label{table:linresults}
+\caption{Shown above as accuraccy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
 \end{figure}
@@
 
+There are several lines of related work which explore stochastic tree models, ranging exploration with recurrent neural networks to agglomerative pairwise associations. 
+%
+Most similar to ours, is the top down tree structure of \newcite{zhang2016top}.
+%
+In this model, first, two long short-term memory (LSTM) networks are used to make branch exploration decisions, and then an additional two LSTM networks are used to continue along these branches. 
+%
+The work of \newcite{Tai2015} similarly use an LSTM, but instead work from the leaves upward to merge the hidden states of children inside the recurrent step. 
+%
+The upward merging of children to form representations for parents has also been studied as pairwise agglomerations using recursive neural networks \cite{Socher2010}.
+
+A growing body of work investigates modeling long distance relationships using efficient stacks in modern neural architecture. 
+%
+\newcite{dyer2015transition} utilize an indexable LSTM as a neural stack for transition-based parsing.  
+%
+Similarly, \newcite{bowman2016fast} utilize both a stack and buffer for sentence comprehension by utilizing an indexable structure and performing computations with indices to these structures. 
+%
+Finally, recent work models long distance relationships using the dynamic composability in ApolloCaffe, a fork of the Caffe library \cite{jia2014caffe}, to build different network structures based on parse trees \cite{Andreas2016LearningTC}.
+
 \section{Conclusion}
 \label{sec:conclusion}
+
+While many neural models of grammar have focused on parsing, there has been relatively less attention on generation.
+%
+This is in large part due to the different kinds of decisions that have to be made in order for generation to work.
+
+The models outlined in this paper push forward methodologies for modeling tree grammars in neural networks.
+
+A lot of work has to be done to take advantage of deep learning and translate that to tasks beyond parsing, especially language production tasks.
+
+This is a first step in a larger research program of stochastic generation models with the kind of generalization that deep learning techniques like word embeddings and recurrent neural networks give you.
 

### 2016-10-05 fergus/algorithms/tagger/tagger.py

@@ -173,8 +173,6 @@ class MRecurrent(object):
         pmap = {-1:0}
         nmap = {}
         out = []
-        import pdb
-        pdb.set_trace()
         # child, parent id, head id, daughter ids
         for i, (child, pid, hid, dids) in enumerate(root.sequence):
             out.append((pmap[pid], hid, dids))

### 2016-10-05 fergus/models/common/__init__.py

@@ -44,11 +44,22 @@ def make_convolutional_embedding(igor):
 
     condensed_spine = Flatten()(spine_conv5)
 
+    igor.logger.info(" + Making convolutional embeddings")
     igor.F_embedspine = DynamicEmbedding(condensed_spine,
                                          W_regularizer=l2(w_decay), # Output: Spine lexicon x Feature
                                          dropout=p_emb,
                                          name='embed_spine') 
 
+
+def make_token_embedding(igor):
+        igor.logger.info(" + Making token embeddings")
+        igor.F_embedspine = Embedding(igor.spine_lexicon_size, 
+                                      igor.spine_convsame_filters, 
+                                      mask_zero=True, 
+                                      W_regularizer=l2(igor.weight_decay), 
+                                      name='token_embedding')
+        igor.preloaded_data = []
+    
 
 def make_logger(igor, loggername):
     import logging

### 2016-10-05 fergus/models/fergus_neuralized/igor.py

@@ -321,6 +321,7 @@ class Igor(DataServer):
             os.makedirs(save_dir)
         
         config = global_config()
+        self.embedding_type = config['embedding_mode']
         data_dir = config['data_dir']
         self.vocman_file = os.path.join(data_dir, self.vocman_file)
         self.vocabs = VocabManager.from_file(self.vocman_file)

### 2016-10-05 fergus/models/fergus_neuralized/model.py

@@ -27,7 +27,7 @@ from keras.utils.visualize_util import plot as kplot
 
 ### project imports
 from .igor import Igor
-from ..common import make_convolutional_embedding, make_logger
+from ..common import make_convolutional_embedding, make_logger, make_token_embedding
 
 
 try:
@@ -99,7 +99,14 @@ class FergusNModel(object):
         Arg:
             theano_kwargs as dict for debugging theano or submitting something custom
         '''
-        make_convolutional_embedding(self.igor)
+        
+        if self.igor.embedding_type == "convolutional":
+            make_convolutional_embedding(self.igor)
+        elif self.igor.embedding_type == "token":
+            make_token_embedding(self.igor)
+        else:
+            raise Exception("Incorrect embedding type")
+        
         B = self.igor.batch_size
         spine_input_shape = (B, self.igor.max_num_supertags) 
         child_input_shape = (B, 1)

### 2016-10-05 fergus/models/fergus_recurrent/igor.py

@@ -347,6 +347,7 @@ class Igor(DataServer):
             os.makedirs(save_dir)
             
         config = global_config()
+        self.embedding_type = config['embedding_mode']
         data_dir = config['data_dir']
         self.vocman_file = os.path.join(data_dir, self.vocman_file)
         self.vocabs = VocabManager.from_file(self.vocman_file)

### 2016-10-05 fergus/models/fergus_recurrent/model.py

@@ -26,7 +26,7 @@ from ikelos.data import Vocabulary, VocabManager
 
 
 from .igor import Igor
-from ..common import make_convolutional_embedding, make_logger
+from ..common import make_convolutional_embedding, make_logger, make_token_embedding
 from ...configs import global_config
 
 try:
@@ -89,8 +89,12 @@ class FergusRModel(object):
             Output:
                 - Distribution over supertags for node_t
         """
-        
-        make_convolutional_embedding(self.igor)
+        if self.igor.embedding_type == "convolutional":
+            make_convolutional_embedding(self.igor)
+        elif self.igor.embedding_type == "token":
+            make_token_embedding(self.igor)
+        else:
+            raise Exception("Incorrect embedding type")
         
         
         spine_input_shape = (self.igor.batch_size,

### 2016-10-05 runners/run_trainer.py

@@ -1,14 +1,14 @@
 """An implemenation of FERGUS's supertagger
 
 Usage:
-    run_trainer.py fergusr
-    run_trainer.py fergusn
+    run_trainer.py fergusr 
+    run_trainer.py fergusn 
     run_trainer.py (-h | --help)
 
-    fergusr                     train the fergus-r model
-    fergusn                     train the fergus-n model
-
-
+    fergusr                   train the fergus-r model
+    fergusn                   train the fergus-n model
+    
+    
 Notes: 
 as per Bangalore & Rambow, 2000 
 "Exploiting a Probabilistic Hierarchical Model for Generation"

### 2016-10-07 fergus/__init__.py
cleaning

@@ -1,0 +1,4 @@
+from __future__ import absolute_import
+
+import os
+ROOTPATH = os.path.dirname(os.path.realpath(__file__))

### 2016-10-07 fergus/algorithms/tagger/__init__.py

@@ -3,18 +3,11 @@ organization of algorithm:
     - tagger.Tagger serves as the entry point for the algorithm
     - tagger.Tagger runs the GPU model
     - tagger.Tagger can either have multiple methods or be subclasses
-    - tagger_node.* handle the specific node propertis for Fergus-0 and Fergus-R
+    - tagger_node.* handle the specific node propertis for Fergus-N and Fergus-R
     - astar_core.run handles the astar algorithm proper
-    - astar_state.* should implement specific state information for Fergus-0 and Fergus-R
+    - astar_state.* should implement specific state information for Fergus-N and Fergus-R
     -
 '''
 from __future__ import absolute_import
 from .tagger import Tagger
-
-
-def fergus_zero():
-    pass
-
-def fergus_recur():
-    pass
 

### 2016-10-07 fergus/algorithms/tagger/astar_state.py

@@ -156,11 +156,11 @@ class BaseState(object):
 
 
 
-class RState(BaseState):    
+class FergusRSearchState(BaseState):    
     def update_score(self, *args):
         return -self.daughter_scores[self.n_finished]
 
-class ZeroState(BaseState):
+class FergusNSearchState(BaseState):
     def update_score(self, spine_id, node):     
         return node.parent.score_update(self.n_finished, node.spine_lookup[spine_id], self.spine_id)
 

### 2016-10-07 fergus/algorithms/tagger/tagger.py

@@ -29,13 +29,13 @@ class Tagger(object):
         self.runtime = -1
         
     @classmethod
-    def fergusn(cls, path):
-        tg = cls(path, MNeuralized())
+    def fergusn(cls, path, init_config):
+        tg = cls(path, NeuralizedModel(init_config))
         return tg
 
     @classmethod
     def fergusr(cls, path):
-        tg = cls(path, MRecurrent())
+        tg = cls(path, RecurrentModel(init_config))
         return tg
 
     def stats(self, raw=False):
@@ -79,12 +79,12 @@ class Tagger(object):
 
 
 @singleton
-class MNeuralized(object):
-    Node = tagger_node.ZeroNode
-    State = astar_state.ZeroState
+class NeuralizedModel(object):
+    Node = tagger_node.FergusNNode
+    State = astar_state.FergusNSearchState
 
-    def __init__(self):
-        self.model = fergus_neuralized.globally_set_model()
+    def __init__(self, init_config):
+        self.model = fergus_neuralized.get_model(init_config)
         self.model.igor.compute_affordances()
         self.igor = self.model.igor
 
@@ -160,12 +160,12 @@ class MNeuralized(object):
                 leaf.distribution = -1 * np.log(prior_values[li])
 
 @singleton
-class MRecurrent(object):
-    Node = tagger_node.RNode
-    State = astar_state.RState
+class RecurrentModel(object):
+    Node = tagger_node.FergusRNode
+    State = astar_state.FergusRSearchState
 
-    def __init__(self):
-        self.model = fergus_recurrent.globally_set_model()
+    def __init__(self, init_config):
+        self.model = fergus_recurrent.get_model(init_config)
         self.model.igor.compute_affordances()
         self.igor = self.model.igor
 

### 2016-10-07 fergus/algorithms/tagger/tagger_node.py

@@ -162,9 +162,9 @@ class Node(object):
 
 
 
-class ZeroNode(Node):
+class FergusNNode(Node):
     def __init__(self, node, ref, parent=None):
-        super(ZeroNode, self).__init__(node, ref, parent)
+        super(FergusNNode, self).__init__(node, ref, parent)
         if parent is not None:
             self.parent_spine_lookup = parent.spine_lookup
 
@@ -204,7 +204,7 @@ class ZeroNode(Node):
         new_inside = dot_dist[dot_spine_idx]
         return new_inside - old_downside - old_inside
 
-class RNode(Node):
+class FergusRNode(Node):
     def compute_best(self):
         self.best_score = self.distribution.min()
         self.best_index = self.distribution.argmin()

### 2016-10-07 fergus/configs/__init__.py

@@ -1,23 +1,41 @@
 import yaml
 import os
 import sys
-PATH = os.path.dirname(os.path.realpath(__file__))
+from .. import ROOTPATH
 
 
 def get_config(filename):
-    filename = os.path.join(PATH, filename)
+    filename = os.path.join(ROOTPATH, filename)
     config = {}
 
     with open(filename) as fp:
         for key, value in yaml.load(fp.read()).items():
             if "dir_rel" in key:
-                config[key.replace("_rel","")] = os.path.join(PATH, value)
+                config[key.replace("_rel","")] = os.path.join(ROOTPATH, value)
             elif "dir_abs" in key:
                 config[key.replace("_abs","")] = value
             else:
                 config[key] = value
     return config
         
-def global_config(filename='global_config.conf'):
+def global_config(filename='premade_confs/global.conf'):
     return get_config(filename)
+    
+    
+def compose_configs(data=None, model=None, embedding=None, use_premades=True):
+    config = global_config()
+    prefix = ''
+    format_conf = lambda *args: os.path.join(args)+".conf"
+    if use_premades:
+        prefix = 'premade_confs'
+    
+    if data:
+        config.update(get_config(format_conf(prefix, data)))
+    if model:
+        config.update(get_config(format_conf(prefix, model)))
+    if embedding:
+        config['embedding_type'] = embedding
+    
+    return config
+    
     

### 2016-10-07 fergus/models/common/__init__.py

@@ -1,130 +1,4 @@
-from ikelos.layers import DynamicEmbedding, LambdaMask
-import keras.backend as K
-from keras.layers import Convolution2D, Lambda, Flatten, DataLayer, Embedding
-from keras.regularizers import l2
-from keras.engine import merge
-import os
-concat = lambda layers: merge(layers, mode='concat')
+from __future__ import absolute_import
 
-def make_convolutional_embedding(igor):
-    p_emb = igor.p_emb_dropout
-    p_W = igor.p_W_dropout
-    p_U = igor.p_U_dropout
-    w_decay = igor.weight_decay
-    p_mlp = igor.p_mlp_dropout
-    p_type = p_emb * igor.total_num_type / igor.total_num_pos # weight by the ratio to even things out
-    Z_spinepos_layer = DataLayer(igor.X_spinepos, input_dtype='int32', name='spine_pos_lexicon')
-    Z_spinetype_layer = DataLayer(igor.X_spinetype, input_dtype='int32', name='spine_type_lexicon')
-
-    Z_spinepos = Z_spinepos_layer.tensor
-    Z_spinetype = Z_spinetype_layer.tensor
-    igor.preloaded_data = [Z_spinepos, Z_spinetype]
-
-    F_embedpos = Embedding(igor.total_num_pos, igor.pos_embedding_size, mask_zero=True,
-                           W_regularizer=l2(w_decay), dropout=p_emb, name='embed_pos')
-    F_embedtype = Embedding(igor.total_num_type, igor.type_embedding_size, mask_zero=True,
-                           W_regularizer=l2(w_decay), dropout=p_type, name='embed_type')
-
-    F_stripmask = LambdaMask(lambda x, mask: None, name='strip_mask')
-    spine_pos = F_stripmask(F_embedpos(Z_spinepos))
-    spine_type = F_stripmask(F_embedtype(Z_spinetype))
-
-    L = igor.spine_lexicon_size
-    Sp = igor.max_spine_length
-    C = igor.max_context_size
-    poty = igor.pos_embedding_size + igor.type_embedding_size
-
-    spine_cat = concat([spine_pos, spine_type])
-    spine_cat = Lambda(lambda xin: K.permute_dimensions(xin, (0,3,1,2)), output_shape=lambda *args:(L, poty, Sp, C))(spine_cat)
[…]

### 2016-10-07 fergus/models/common/embeddings.py

@@ -1,0 +1,103 @@
+from ikelos.layers import DynamicEmbedding, LambdaMask
+import keras.backend as K
+from keras.layers import Convolution2D, Lambda, Flatten, DataLayer, Embedding
+from keras.regularizers import l2
+from keras.engine import merge
+import os
+concat = lambda layers: merge(layers, mode='concat')
+
+def make_convolutional_embedding(igor):
+    p_emb = igor.p_emb_dropout
+    p_W = igor.p_W_dropout
+    p_U = igor.p_U_dropout
+    w_decay = igor.weight_decay
+    p_mlp = igor.p_mlp_dropout
+    p_type = p_emb * igor.total_num_type / igor.total_num_pos # weight by the ratio to even things out
+    Z_spinepos_layer = DataLayer(igor.X_spinepos, input_dtype='int32', name='spine_pos_lexicon')
+    Z_spinetype_layer = DataLayer(igor.X_spinetype, input_dtype='int32', name='spine_type_lexicon')
+
+    Z_spinepos = Z_spinepos_layer.tensor
+    Z_spinetype = Z_spinetype_layer.tensor
+    igor.preloaded_data = [Z_spinepos, Z_spinetype]
+
+    F_embedpos = Embedding(igor.total_num_pos, igor.pos_embedding_size, mask_zero=True,
+                           W_regularizer=l2(w_decay), dropout=p_emb, name='embed_pos')
+    F_embedtype = Embedding(igor.total_num_type, igor.type_embedding_size, mask_zero=True,
+                           W_regularizer=l2(w_decay), dropout=p_type, name='embed_type')
+
+    F_stripmask = LambdaMask(lambda x, mask: None, name='strip_mask')
+    spine_pos = F_stripmask(F_embedpos(Z_spinepos))
+    spine_type = F_stripmask(F_embedtype(Z_spinetype))
+
+    L = igor.spine_lexicon_size
+    Sp = igor.max_spine_length
+    C = igor.max_context_size
+    poty = igor.pos_embedding_size + igor.type_embedding_size
+
+    spine_cat = concat([spine_pos, spine_type])
+    spine_cat = Lambda(lambda xin: K.permute_dimensions(xin, (0,3,1,2)), output_shape=lambda *args:(L, poty, Sp, C))(spine_cat)
+    spine_conv1 = Convolution2D(igor.spine_convsame_filters, 1, 2, border_mode='valid')(spine_cat) 
[…]

### 2016-10-07 fergus/models/common/loggers.py

@@ -1,0 +1,28 @@
+import logging
+import os
+
+def make_logger(igor, loggername):
+    igor.logfile_level = 'debug'
+    igor.logshell_level = 'info'
+    igor.logfile_location = '.'
+    logger = logging.getLogger(loggername)
+    levels = {"debug": logging.DEBUG, "warning":logging.WARNING,
+              "info": logging.INFO, "error":logging.ERROR,
+              "critical":logging.CRITICAL}
+    logger.setLevel(logging.DEBUG)
+
+    if igor.logfile_level != "off":
+        safe_loc = igor.logfile_location+("/" if igor.logfile_location[-1] != "/" else "")
+        if not os.path.exists(safe_loc):
+            os.makedirs(safe_loc)
+        fh = logging.FileHandler(os.path.join(igor.logfile_location, 
+                                              "{}.debug.log".format(loggername)))
+        fh.setLevel(levels[igor.logfile_level])
+        logger.addHandler(fh)
+
+    if igor.logshell_level != "off":
+        sh = logging.StreamHandler()
+        sh.setLevel(levels[igor.logfile_level])
+        logger.addHandler(sh)
+
+    return logger

### 2016-10-07 fergus/models/fergus_neuralized/model.py

@@ -60,13 +60,26 @@ def traverse_nodes(layer, from_name="top", down=True):
 
 class FergusNModel(object):
     def __init__(self, igor):
-        self.logger = igor.logger = make_logger(igor, 'fergusn_dev')
+        
+        now = datetime.now()
+        self.run_name = "fergusn_{}mo_{}day_{}hr_{}min".format(now.month, now.day, 
+                                                                now.hour, now.minute)
+        log_location = os.path.join(igor.log_dir, self.run_name+".log")
+        self.logger = igor.logger = make_logger(igor, log_location)
         igor.verify_directories()
         self.igor = igor
 
     @classmethod
     def from_yaml(cls, yamlfile, kwargs=None):
         igor = Igor.from_file(yamlfile)
+        igor.prep()
+        model = cls(igor)
+        model.make(kwargs)
+        return model
+
+    @classmethod
+    def from_config(cls, config, kwargs=None):
+        igor = Igor(config)
         igor.prep()
         model = cls(igor)
         model.make(kwargs)
@@ -270,10 +283,12 @@ class FergusNModel(object):
         callbacks = [ProgbarV2(3, 10, replacers=replacers)]
         checkpoint_fp = join(self.igor.model_location,
                              self.igor.saving_prefix,
-                             "cp_weights.h5")
+                             self.igor.checkpoint_weights)
         self.logger.info("+ Model Checkpoint: {}".format(checkpoint_fp))
         callbacks += [ModelCheckpoint(filepath=checkpoint_fp, verbose=1, save_best_only=True)]
-        callbacks += [LearningRateScheduler(lambda epoch: self.igor.LR * 0.99**epoch)]
+        callbacks += [LearningRateScheduler(lambda epoch: self.igor.LR * 0.9)]
+        csv_location = os.path.join(self.igor.log_dir, self.run_name+".csv")
+        callbacks += [CSVLogger(csv_location)]
         self.model.fit_generator(generator=train_data, samples_per_epoch=N, nb_epoch=E,
                                  callbacks=callbacks, verbose=1,
                                  validation_data=dev_data,

### 2016-10-07 fergus/models/fergus_recurrent/__init__.py

@@ -5,30 +5,33 @@ from ...configs import global_config, get_config
 
 PATH = os.path.dirname(os.path.realpath(__file__))
 
-def TrainingModel():
-    print("Retrieving training model")
-    config = global_config()
-    config_file = config['fergus_r_train_config']
-    config_file = os.path.join(config['config_dir'], config_file)
-    return model.FergusRModel.from_yaml(config_file)
-    
-    
-def DevelopmentModel():
-    print("Retrieving development model")
-    config = global_config()
-    config_file = config['fergus_r_dev_config']
-    config_file = os.path.join(config['config_dir'], config_file)
-    return model.FergusRModel.from_yaml(config_file)
+def get_model(init_config):
+    return model.FergusRModel.from_config(init_config)
 
-def TestingModel():
-    print("Retrieving testing model")
-    config = global_config()
-    config_file = config['fergus_r_test_config']
-    config_file = os.path.join(config['config_dir'], config_file)
-    return model.FergusRModel.from_yaml(config_file)
+# def TrainingModel():
+#     print("Retrieving training model")
+#     config = global_config()
+#     config_file = config['fergus_r_train_config']
+#     config_file = os.path.join(config['config_dir'], config_file)
+#     return model.FergusRModel.from_yaml(config_file)
     
-def globally_set_model():
-    config = global_config()
-    return {'dev': DevelopmentModel,
-            'test': TestingModel,
-            'train': TrainingModel}[config['tag_mode']]()
+    
+# def DevelopmentModel():
+#     print("Retrieving development model")
+#     config = global_config()
+#     config_file = config['fergus_r_dev_config']
+#     config_file = os.path.join(config['config_dir'], config_file)
+#     return model.FergusRModel.from_yaml(config_file)
+
+# def TestingModel():
+#     print("Retrieving testing model")
+#     config = global_config()
+#     config_file = config['fergus_r_test_config']
+#     config_file = os.path.join(config['config_dir'], config_file)
+#     return model.FergusRModel.from_yaml(config_file)
+    
+# def globally_set_model():
+#     config = global_config()
+#     return {'dev': DevelopmentModel,
+#             'test': TestingModel,
+#             'train': TrainingModel}[config['tag_mode']]()

### 2016-10-07 fergus/models/fergus_recurrent/model.py

@@ -6,6 +6,7 @@ import time
 import json
 import numpy as np
 import sys
+from datetime import datetime
 
 import theano
 
@@ -27,7 +28,6 @@ from ikelos.data import Vocabulary, VocabManager
 
 from .igor import Igor
 from ..common import make_convolutional_embedding, make_logger, make_token_embedding
-from ...configs import global_config
 
 try:
     import cPickle as pickle
@@ -41,15 +41,26 @@ merge_batch = lambda layers: merge(layers, mode='concat', concat_axis=0)
 
 class FergusRModel(object):
     def __init__(self, igor):
-        config = global_config()
-        self.logger = igor.logger = make_logger(igor, 'fergusr_'+config['tag_mode'])
-        self.igor = igor
-        self.igor.verify_directories()
-        self.model = None
         
+        now = datetime.now()
+        self.run_name = "fergusr_{}mo_{}day_{}hr_{}min".format(now.month, now.day, 
+                                                                now.hour, now.minute)
+        log_location = os.path.join(igor.log_dir, self.run_name+".log")
+        self.logger = igor.logger = make_logger(igor, log_location)
+        igor.verify_directories()
+        self.igor = igor
+
     @classmethod
     def from_yaml(cls, yamlfile, kwargs=None):
         igor = Igor.from_file(yamlfile)
+        igor.prep()
+        model = cls(igor)
+        model.make(kwargs)
+        return model
+
+    @classmethod
+    def from_config(cls, config, kwargs=None):
+        igor = Igor(config)
         igor.prep()
         model = cls(igor)
         model.make(kwargs)
@@ -64,6 +75,7 @@ class FergusRModel(object):
             self.model.load_weights(weight_file, by_name=True)
         else:
             self.logger.warning("- Checkpoint weights do not exist; {}".format(weight_file))
+
 
 
     def plot(self):
@@ -211,7 +223,10 @@ class FergusRModel(object):
         self.logger.info("+ Model Checkpoint: {}".format(checkpoint_fp))
         
         callbacks += [ModelCheckpoint(filepath=checkpoint_fp, verbose=1, save_best_only=True)]
-        callbacks += [LearningRateScheduler(lambda epoch: self.igor.LR * 0.9**(epoch%15))]
+        callbacks += [LearningRateScheduler(lambda epoch: self.igor.LR * 0.9**(epoch))]
+        
+        csv_location = os.path.join(self.igor.log_dir, self.run_name+".csv")
[…]

### 2016-10-07 runners/run_tagger.py

@@ -1,3 +1,22 @@
+"""Supertagging routine for 
+
+Usage:
+    run_tagger.py (fergusr|fergusn) (dev|test) (convolution|token) 
+    run_tagger.py (-h | --help)
+
+Options:
+    fergusr,fergusn             choose the model, fergus recurrent or fergus neuralized
+    dev,test                    choose the dataset to run on
+    convolution,token           chosen the supertag embedding style
+    
+    
+    
+Notes: 
+McMahan and Stone
+"Syntactic realization with data-driven neural tree grammars" 
+published in COLING 2016
+"""
+
 from __future__ import print_function, division, absolute_import
 import pickle
 import sys
@@ -9,43 +28,42 @@ import fergus
 ### additional packages needed
 from tqdm import tqdm
 from sqlitedict import SqliteDict
+from docopt import docopt
 import time
 ### code in this repo
 from fergus.algorithms.tagger.utils import rollout
-from fergus.configs import global_config
+from fergus.configs import global_config, get_config
 from fergus.algorithms.tagger import Tagger
 
-def run(model, config):
-
-    print("loading data")
-    filename_key = "{}_filename".format(config['tag_mode'])
-    filepath = os.path.join(config['data_dir'], config[filename_key])
+def run(model_factory, config, db_name):
     
-    with open(filepath, 'rb') as fp:
+    db_name = os.path.join(config['data_dir'], db_name)
+    data_file = os.path.join(config['data_dir'], config['data_file'])
+
+    with open(data_file, 'rb') as fp:
         data = pickle.load(fp)
-    print("data loaded")
     
     if config['fast_mode']: 
         import numpy as np
         data = [data[i] for i in np.random.random_integers(0, len(data), config['fast_mode'])]
 
     all_stats = []
-    with SqliteDict(config['data_store_db'], tablename='indata') as datadb:
+    with SqliteDict(db_name, tablename='tagged_data') as datadb:
         capacitor = {}
         dbkeys = set(map(int,datadb.keys()))
         if len(dbkeys)>0:
-            print("{} already done; overwrite={}".format(len(dbkeys), config['overwrite_store']))
+            print("{} already done.  will be overwriting={}".format(len(dbkeys), config['overwrite_store']))
         skipped = 0
         for i, datum in tqdm(enumerate(data), total=len(data)):
             if i in dbkeys and not config['overwrite_store']: continue
-
             if len(datum) > 100: 
                 skipped += 1
[…]

### 2016-10-07 runners/run_trainer.py

@@ -12,9 +12,6 @@ Usage:
 Notes: 
 as per Bangalore & Rambow, 2000 
 "Exploiting a Probabilistic Hierarchical Model for Generation"
-
-
-
 """
 import os
 import sys

### 2016-10-07 fergus/configs/__init__.py
moving weights in

@@ -5,14 +5,14 @@ from .. import ROOTPATH
 
 
 def get_config(filename):
-    filename = os.path.join(ROOTPATH, filename)
+    filename = os.path.join(ROOTPATH, "configs", filename)
     config = {}
 
     with open(filename) as fp:
         for key, value in yaml.load(fp.read()).items():
-            if "dir_rel" in key:
+            if "_rel" in key:
                 config[key.replace("_rel","")] = os.path.join(ROOTPATH, value)
-            elif "dir_abs" in key:
+            elif "_abs" in key:
                 config[key.replace("_abs","")] = value
             else:
                 config[key] = value
@@ -25,7 +25,7 @@ def global_config(filename='premade_confs/global.conf'):
 def compose_configs(data=None, model=None, embedding=None, use_premades=True):
     config = global_config()
     prefix = ''
-    format_conf = lambda *args: os.path.join(args)+".conf"
+    format_conf = lambda *args: os.path.join(*args)+".conf"
     if use_premades:
         prefix = 'premade_confs'
     
@@ -35,6 +35,10 @@ def compose_configs(data=None, model=None, embedding=None, use_premades=True):
         config.update(get_config(format_conf(prefix, model)))
     if embedding:
         config['embedding_type'] = embedding
+    
+    if 'saving_prefix' in config and 'embedding_type' in config:
+        config['saving_prefix'] = "{}_{}".format(config['saving_prefix'], 
+                                                 config['embedding_type'])
     
     return config
     

### 2016-10-07 fergus/models/fergus_neuralized/__init__.py

@@ -1,33 +1,32 @@
 from __future__ import absolute_import
 import os
 from . import model
-from ...configs import global_config, get_config
 
-PATH = os.path.dirname(os.path.realpath(__file__))
-
-def TrainingModel():
-    print("Retrieving training model")
-    config = global_config()
-    config_file = config['fergus_n_train_config']
-    config_file = os.path.join(config['config_dir'], config_file)
-    return model.FergusNModel.from_yaml(config_file)
+def get_model(init_config):
+    return model.FergusNModel.from_config(init_config)
+# def TrainingModel():
+#     print("Retrieving training model")
+#     config = global_config()
+#     config_file = config['fergus_n_train_config']
+#     config_file = os.path.join(config['config_dir'], config_file)
+#     return model.FergusNModel.from_yaml(config_file)
     
-def DevelopmentModel():
-    print("Retrieving development model")
-    config = global_config()
-    config_file = config['fergus_n_dev_config']
-    config_file = os.path.join(config['config_dir'], config_file)
-    return model.FergusNModel.from_yaml(config_file)
+# def DevelopmentModel():
+#     print("Retrieving development model")
+#     config = global_config()
+#     config_file = config['fergus_n_dev_config']
+#     config_file = os.path.join(config['config_dir'], config_file)
+#     return model.FergusNModel.from_yaml(config_file)
 
-def TestingModel():
-    print("Retrieving testing model")
-    config = global_config()
-    config_file = config['fergus_n_test_config']
-    config_file = os.path.join(config['config_dir'], config_file)
-    return model.FergusNModel.from_yaml(config_file)
+# def TestingModel():
+#     print("Retrieving testing model")
+#     config = global_config()
+#     config_file = config['fergus_n_test_config']
+#     config_file = os.path.join(config['config_dir'], config_file)
+#     return model.FergusNModel.from_yaml(config_file)
 
-def globally_set_model():
-    config = global_config()
-    return {'dev': DevelopmentModel,
-            'test': TestingModel,
-            'train': TrainingModel}[config['tag_mode']]()
+# def globally_set_model():
+#     config = global_config()
+#     return {'dev': DevelopmentModel,
+#             'test': TestingModel,
+#             'train': TrainingModel}[config['tag_mode']]()

### 2016-10-07 fergus/models/fergus_neuralized/igor.py

@@ -18,7 +18,6 @@ except:
     import pickle
 
 from . import utils
-from ...configs import global_config
 
 class Igor(DataServer):
     """Handle the data for FERGUS 
@@ -320,18 +319,15 @@ class Igor(DataServer):
         if not os.path.exists(save_dir):
             os.makedirs(save_dir)
         
-        config = global_config()
-        self.embedding_type = config['embedding_mode']
-        data_dir = config['data_dir']
-        self.vocman_file = os.path.join(data_dir, self.vocman_file)
+        self.vocman_file = os.path.join(self.data_dir, self.vocman_file)
         self.vocabs = VocabManager.from_file(self.vocman_file)
         self.vocabs.freeze_all()
         self.total_num_type = len(self.vocabs.type)
         self.total_num_pos = len(self.vocabs.pos)
-        self.train_fp = os.path.join(data_dir, self.train_filepath)
-        self.dev_fp = os.path.join(data_dir, self.dev_filepath)
-        self.test_fp = os.path.join(data_dir, self.test_filepath)
-        head2spine_file = os.path.join(data_dir, self.head2spine_file)
+        self.train_fp = os.path.join(self.data_dir, self.train_filepath)
+        self.dev_fp = os.path.join(self.data_dir, self.dev_filepath)
+        self.test_fp = os.path.join(self.data_dir, self.test_filepath)
+        head2spine_file = os.path.join(self.data_dir, self.head2spine_file)
     
         self.debugkey = defaultdict(lambda: False)
 
@@ -353,7 +349,7 @@ class Igor(DataServer):
         utils.compute_parameters(self)
         self.make_spine_tensors()
         if self.embeddings_file and not self.from_checkpoint:
-            embeddings_file = os.path.join(data_dir, self.embeddings_file)
+            embeddings_file = os.path.join(self.data_dir, self.embeddings_file)
             self.saved_embeddings = np.load(embeddings_file)
         else:
             self.saved_embeddings = None

### 2016-10-07 fergus/models/fergus_neuralized/model.py

@@ -6,7 +6,7 @@ import time
 import json
 import numpy as np
 from os.path import join, exists
-
+from datetime import datetime
 
 #### need ikelos library; see braingineer/ikelos
 from ikelos.data import Vocabulary, VocabManager
@@ -64,7 +64,7 @@ class FergusNModel(object):
         now = datetime.now()
         self.run_name = "fergusn_{}mo_{}day_{}hr_{}min".format(now.month, now.day, 
                                                                 now.hour, now.minute)
-        log_location = os.path.join(igor.log_dir, self.run_name+".log")
+        log_location = join(igor.log_dir, self.run_name+".log")
         self.logger = igor.logger = make_logger(igor, log_location)
         igor.verify_directories()
         self.igor = igor
@@ -287,7 +287,7 @@ class FergusNModel(object):
         self.logger.info("+ Model Checkpoint: {}".format(checkpoint_fp))
         callbacks += [ModelCheckpoint(filepath=checkpoint_fp, verbose=1, save_best_only=True)]
         callbacks += [LearningRateScheduler(lambda epoch: self.igor.LR * 0.9)]
-        csv_location = os.path.join(self.igor.log_dir, self.run_name+".csv")
+        csv_location = join(self.igor.log_dir, self.run_name+".csv")
         callbacks += [CSVLogger(csv_location)]
         self.model.fit_generator(generator=train_data, samples_per_epoch=N, nb_epoch=E,
                                  callbacks=callbacks, verbose=1,

### 2016-10-07 fergus/models/fergus_recurrent/model.py

@@ -45,7 +45,7 @@ class FergusRModel(object):
         now = datetime.now()
         self.run_name = "fergusr_{}mo_{}day_{}hr_{}min".format(now.month, now.day, 
                                                                 now.hour, now.minute)
-        log_location = os.path.join(igor.log_dir, self.run_name+".log")
+        log_location = join(igor.log_dir, self.run_name+".log")
         self.logger = igor.logger = make_logger(igor, log_location)
         igor.verify_directories()
         self.igor = igor
@@ -225,7 +225,7 @@ class FergusRModel(object):
         callbacks += [ModelCheckpoint(filepath=checkpoint_fp, verbose=1, save_best_only=True)]
         callbacks += [LearningRateScheduler(lambda epoch: self.igor.LR * 0.9**(epoch))]
         
-        csv_location = os.path.join(self.igor.log_dir, self.run_name+".csv")
+        csv_location = join(self.igor.log_dir, self.run_name+".csv")
         callbacks += [CSVLogger(csv_location)]
         self.model.fit_generator(generator=train_data, samples_per_epoch=N, nb_epoch=E,
                                  callbacks=callbacks, verbose=1,

### 2016-10-07 runners/run_tagger.py

@@ -1,7 +1,7 @@
 """Supertagging routine for 
 
 Usage:
-    run_tagger.py (fergusr|fergusn) (dev|test) (convolution|token) 
+    run_tagger.py (fergusr|fergusn) (dev|test) (convolutional|token) 
     run_tagger.py (-h | --help)
 
 Options:
@@ -32,7 +32,7 @@ from docopt import docopt
 import time
 ### code in this repo
 from fergus.algorithms.tagger.utils import rollout
-from fergus.configs import global_config, get_config
+from fergus.configs import compose_configs
 from fergus.algorithms.tagger import Tagger
 
 def run(model_factory, config, db_name):
@@ -92,7 +92,7 @@ def run(model_factory, config, db_name):
 
 
 if __name__ == '__main__':
-    args = docopt(__doc__, "Supertagger.  publication version. 2016")
+    args = docopt(__doc__, version="Supertagger.  publication version. 2016")
     ### the options in docopt doc were mutually exlusive
     data_type = "dev" if args['dev'] else "test"
     embed_type = "convolutional" if args['convolutional'] else "token"
@@ -104,4 +104,4 @@ if __name__ == '__main__':
     
     config = compose_configs(data=data_type, model=model_type, embedding=embed_type)
     
-    run(model_factory, config, db_name))
+    run(model_factory, config, db_name)

### 2016-10-07 fergus/models/common/embeddings.py
introducing shallowconv

@@ -51,7 +51,7 @@ def make_convolutional_embedding(igor):
                                          name='embed_spine') 
 
 
-def make_convolutional_embedding_v2(igor):
+def make_shallow_convolutional_embedding(igor):
     p_emb = igor.p_emb_dropout
     p_W = igor.p_W_dropout
     p_U = igor.p_U_dropout

### 2016-10-07 fergus/models/fergus_neuralized/model.py

@@ -27,7 +27,8 @@ from keras.utils.visualize_util import plot as kplot
 
 ### project imports
 from .igor import Igor
-from ..common import make_convolutional_embedding, make_logger, make_token_embedding
+from ..common import make_convolutional_embedding, make_logger, make_token_embedding, \
+                     make_shallow_convolutional_embedding
 
 
 try:
@@ -117,6 +118,8 @@ class FergusNModel(object):
             make_convolutional_embedding(self.igor)
         elif self.igor.embedding_type == "token":
             make_token_embedding(self.igor)
+        elif self.igor.embedding_type == "shallowconv":
+            make_shallow_convolutional_embedding(self.igor)
         else:
             raise Exception("Incorrect embedding type")
         

### 2016-10-07 fergus/models/fergus_recurrent/model.py

@@ -27,7 +27,8 @@ from ikelos.data import Vocabulary, VocabManager
 
 
 from .igor import Igor
-from ..common import make_convolutional_embedding, make_logger, make_token_embedding
+from ..common import make_convolutional_embedding, make_logger, make_token_embedding, \
+                     make_shallow_convolutional_embedding
 
 try:
     import cPickle as pickle
@@ -105,6 +106,8 @@ class FergusRModel(object):
             make_convolutional_embedding(self.igor)
         elif self.igor.embedding_type == "token":
             make_token_embedding(self.igor)
+        elif self.igor.embedding_type == "shallowconv":
+            make_shallow_convolutional_embedding(self.igor)
         else:
             raise Exception("Incorrect embedding type")
         

### 2016-10-07 runners/run_trainer.py

@@ -1,7 +1,7 @@
 """An implemenation of FERGUS's supertagger
 
 Usage:
-    run_trainer.py fergusr 
+    run_trainer.py (fergusr|fergusn) (convolutional|token|shallowconv)
     run_trainer.py fergusn 
     run_trainer.py (-h | --help)
 
@@ -36,13 +36,18 @@ if __name__ == "__main__":
     args = docopt(__doc__, version='FERGUS trainer; October 2016')
 
     try:
-        if args['fergusr']:
-            model = fergus_recurrent.TrainingModel()
-        elif args['fergusn']:
-            model = fergus_neuralized.TrainingModel()
+        if args['convolutional']:
+            embed_type = 'convolutional'
+        elif args['token']:
+            embed_type = 'token'
         else:
-            print("No other options currently")
-            sys.exit(0)
+            embed_type = 'shallowconv'
+        model_type = "fergusn" if args['fergusn'] else 'fergusr'
+        model_factory = {'fergusn': fergus_neuralized, 
+                         'fergusr': fergus_recurrent}[model_type]
+        config = compose_configs(data='train', model=model_type, embedding=embed_type)
+    
+        model = model_factory.get_model(config)
         model.train()
 
     except KeyboardInterrupt as e:

### 2016-10-07 fergus/models/fergus_recurrent/igor.py
a fix

@@ -19,7 +19,7 @@ except:
     import pickle
 
 from . import utils
-from ...configs import global_config
+
 
 class Igor(DataServer):
     """Handle the data for FERGUS-R
@@ -346,19 +346,16 @@ class Igor(DataServer):
         if not os.path.exists(save_dir):
             os.makedirs(save_dir)
             
-        config = global_config()
-        self.embedding_type = config['embedding_mode']
-        data_dir = config['data_dir']
-        self.vocman_file = os.path.join(data_dir, self.vocman_file)
+        self.vocman_file = os.path.join(self.data_dir, self.vocman_file)
         self.vocabs = VocabManager.from_file(self.vocman_file)
         self.vocabs.freeze_all()
         self.total_num_type = len(self.vocabs.type)
         self.total_num_pos = len(self.vocabs.pos)
-        self.train_fp = os.path.join(data_dir, self.train_filepath)
-        self.dev_fp = os.path.join(data_dir, self.dev_filepath)
-        self.test_fp = os.path.join(data_dir, self.test_filepath)
-        self.embeddings_file = os.path.join(data_dir, self.embeddings_file)
-        cached_param_file = os.path.join(data_dir, self.cached_parameters_file)
+        self.train_fp = os.path.join(self.data_dir, self.train_filepath)
+        self.dev_fp = os.path.join(self.data_dir, self.dev_filepath)
+        self.test_fp = os.path.join(self.data_dir, self.test_filepath)
+        self.embeddings_file = os.path.join(self.data_dir, self.embeddings_file)
+        cached_param_file = os.path.join(self.data_dir, self.cached_parameters_file)
     
         self.debugkey = defaultdict(lambda: False)
 

### 2016-10-07 runners/run_trainer.py

@@ -21,7 +21,7 @@ import fergus
 
 from ikelos.data import Vocabulary
 from fergus.models import fergus_recurrent, fergus_neuralized
-from fergus.configs import global_config, get_config
+from fergus.configs import compose_configs
 from os.path import join
 from docopt import docopt
 import yaml

### 2016-10-07 fergus/models/fergus_neuralized/model.py
a fix

@@ -21,7 +21,8 @@ from keras.layers import Embedding, LSTM, Input, Lambda, \
                          InputLayer, Convolution2D, MaxPooling2D
 from keras.engine import Model, merge
 from keras.optimizers import Adam
-from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, LearningRateScheduler
+from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, \
+                            LearningRateScheduler, CSVLogger
 from keras.regularizers import l2
 from keras.utils.visualize_util import plot as kplot
 

### 2016-10-07 fergus/models/fergus_recurrent/model.py

@@ -17,7 +17,8 @@ from keras.layers import Embedding, LSTM, Input, Lambda, \
                          Reshape, Flatten, Dense, Dropout, \
                          InputLayer, DataLayer, Convolution2D, MaxPooling2D
 from keras.optimizers import Adam
-from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, LearningRateScheduler
+from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, \
+                            LearningRateScheduler, CSVLogger
 from keras.regularizers import l2
 from keras.utils.visualize_util import plot as kplot
 

### 2016-10-07 fergus/models/common/embeddings.py
a fix

@@ -44,7 +44,7 @@ def make_convolutional_embedding(igor):
 
     condensed_spine = Flatten()(spine_conv5)
 
-    igor.logger.info(" + Making convolutional embeddings")
+    igor.logger.info("+ Making convolutional embeddings")
     igor.F_embedspine = DynamicEmbedding(condensed_spine,
                                          W_regularizer=l2(w_decay), # Output: Spine lexicon x Feature
                                          dropout=p_emb,
@@ -86,14 +86,14 @@ def make_shallow_convolutional_embedding(igor):
 
     condensed_spine = Flatten()(spine_conv2)
 
-    igor.logger.info(" + Making version 2 of convolutional embeddings")
+    igor.logger.info("+ Making shallow convolutional embeddings")
     igor.F_embedspine = DynamicEmbedding(condensed_spine,
                                          W_regularizer=l2(w_decay), # Output: Spine lexicon x Feature
                                          dropout=p_emb,
                                          name='embed_spine') 
 
 def make_token_embedding(igor):
-        igor.logger.info(" + Making token embeddings")
+        igor.logger.info("+ Making token embeddings")
         igor.F_embedspine = Embedding(igor.spine_lexicon_size, 
                                       igor.spine_convsame_filters, 
                                       mask_zero=True, 

### 2016-10-07 fergus/models/fergus_recurrent/igor.py

@@ -382,7 +382,7 @@ class Igor(DataServer):
                 pickle.dump(cached_parameters, fp)
 
         self.make_spine_tensors()
-        if self.embeddings_file and not self.from_checkpoint:
+        if self.embeddings_file:
             self.saved_embeddings = np.load(self.embeddings_file)
         else:
             self.saved_embeddings = None

### 2016-10-07 paper/draft.tex
paper changes

@@
 \centering
-\begin{tabular}{|l|p{2cm}|p{2cm}|p{2.3cm}|p{2cm}|p{2.3cm}|}
+\begin{tabular}{|l|p{2cm}|c|c|}
+\cline{3-4}
+\multicolumn{2}{}{}& \multicolumn{2}{ |c| }{ \textbf{Accuracy}} \\ \hline
+\textbf{Model} & \textbf{Embedding}  & \textbf{Before A*} & \textbf{After A*}  \\ \hline
+Fergus-N &  Convolution &  58.17\% & 57.40\% \\ \cline{2-4}
+         &  Token       &  60.69\% & 55.56\% \\ \cline{2-4}
 \hline
-Model & Embedding & Supertag & Root Category & Root Role & Leaf Category  \\ \hline
-Fergus-N &  Convolution &  57.4\% & 81.8\% & 81.0\% & 86.8\% \\ \cline{2-6}
-         &  Token       &  XX & XX & XX & \\ \cline{2-6}
-\hline
-Fergus-R &  Convolution &  36.4\% & 76.9\% & 78.3\% & 80.8\% \\ \cline{2-6}
-         &  Token       &  XX & XX & XX & \\ \cline{2-6}
+Fergus-R &  Convolution &  67.62\% & 57.04\% \\ \cline{2-4}
+         &  Token       &  82.65\% & 62.73\% \\ \cline{2-4}
 \hline

### 2016-10-07 fergus/algorithms/tagger/tagger.py

@@ -34,7 +34,7 @@ class Tagger(object):
         return tg
 
     @classmethod
-    def fergusr(cls, path):
+    def fergusr(cls, path, init_config):
         tg = cls(path, RecurrentModel(init_config))
         return tg
 

### 2016-10-07 fergus/configs/__init__.py

@@ -33,6 +33,8 @@ def compose_configs(data=None, model=None, embedding=None, use_premades=True):
         config.update(get_config(format_conf(prefix, data)))
     if model:
         config.update(get_config(format_conf(prefix, model)))
+        if model+'_batch_size' in config:
+            config['batch_size'] = config[model+'_batch_size']
     if embedding:
         config['embedding_type'] = embedding
     

### 2016-10-07 runners/run_tagger.py

@@ -54,7 +54,7 @@ def run(model_factory, config, db_name):
         if len(dbkeys)>0:
             print("{} already done.  will be overwriting={}".format(len(dbkeys), config['overwrite_store']))
         skipped = 0
-        for i, datum in tqdm(enumerate(data), total=len(data)):
+        for i, datum in tqdm(enumerate(data), total=len(data), ncols=10):
             if i in dbkeys and not config['overwrite_store']: continue
             if len(datum) > 100: 
                 skipped += 1

### 2016-10-07 runners/runall.sh

@@ -1,0 +1,1 @@
+python run_tagger.py fergusr test token && python run_tagger.py fergusn test token && python run_tagger.py fergusn test convolutional && python run_tagger.py fergusr convolutional

### 2016-10-07 fergus/models/common/loggers.py
adding linearizer

@@ -22,7 +22,7 @@ def make_logger(igor, loggername):
 
     if igor.logshell_level != "off":
         sh = logging.StreamHandler()
-        sh.setLevel(levels[igor.logfile_level])
+        sh.setLevel(levels[igor.logshell_level])
         logger.addHandler(sh)
 
     return logger

### 2016-10-07 fergus/models/language_model/__init__.py

@@ -1,0 +1,6 @@
+from __future__ import absolute_import
+
+from .model import LanguageModel
+
+def from_config(config):
+    return LanguageModel.from_config(config)

### 2016-10-07 fergus/models/language_model/igor.py

@@ -1,0 +1,153 @@
+from __future__ import print_function, division, absolute_import
+import yaml
+from baal.utils import loggers
+from tqdm import tqdm, trange
+import numpy as np
+import itertools
+import time
+import sys
+import matplotlib.pyplot as plt
+from copy import copy, deepcopy
+from keras.utils.np_utils import to_categorical
+import os
+from os.path import join
+from . import utils
+import baal
+import gist
+from ikelos.data import VocabManager, Vocabulary
+import sys
+
+try:
+    input = raw_input
+except:
+    pass
+try:
+    import cPickle as pickle
+except:
+    pass
+
+class Igor(object):
+    def __init__(self, config):
+        self.__dict__.update(config)
+
+    @classmethod
+    def from_file(cls, config_file):
+        with open(config_file) as fp:
+            config = yaml.load(fp)
+        return cls(config)
+
+    @property
+    def num_train_batches(self):
+        return len(self.train_data)//self.batch_size
+
+    @property
+    def num_dev_batches(self):
+        return len(self.dev_data)//self.batch_size
+        
+    @property
+    def num_test_batches(self):
+        return len(self.test_data)//self.batch_size
+        
+    @property
+    def num_train_samples(self):
+        return self.num_train_batches // 3 * self.batch_size
+        
+    @property
+    def num_dev_samples(self):
+        return self.num_dev_batches * self.batch_size
+        
+    @property
+    def num_test_samples(self):
+        return self.num_test_batches * self.batch_size
+
+    def serve_sentence(self, data):
+        for data_i in np.random.choice(len(data), len(data), replace=False):
+            in_X = np.zeros(self.max_sequence_len)
+            out_Y = np.zeros(self.max_sequence_len, dtype=np.int32)
+            bigram_data = zip(data[data_i][0:-1], data[data_i][1:])
+            for datum_j,(datum_in, datum_out) in enumerate(bigram_data):
+                in_X[datum_j] = datum_in
+                out_Y[datum_j] = datum_out
+            yield in_X, out_Y
+
+    def serve_batch(self, data):
+        dataiter = self.serve_sentence(data)
+        V = self.vocab_size
+        S = self.max_sequence_len
+        B = self.batch_size
+
+        while dataiter:
+            in_X = np.zeros((B, S), dtype=np.int32)
+            out_Y = np.zeros((B, S, V), dtype=np.int32)
+            next_batch = list(itertools.islice(dataiter, 0, self.batch_size))
[…]

### 2016-10-07 fergus/models/language_model/model.py

@@ -1,0 +1,255 @@
+
+from __future__ import print_function, division
+
+import sys
+import time
+from copy import copy, deepcopy
+from os.path import join, exists
+from collections import Counter
+from math import log
+import itertools
+from datetime import datetime
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+from ikelos.data import Vocabulary
+
+from keras.layers import LSTM, Dense, Embedding, Distribute, Dropout, Input
+from keras.callbacks import Callback, ProgbarLogger, ModelCheckpoint, ProgbarV2, LearningRateScheduler
+from keras.utils.generic_utils import Progbar
+from keras.regularizers import l2
+from keras.optimizers import Adam, SGD
+from keras.engine import Model
+import keras.backend as K
+
+
+try:
+    input = raw_input
+except:
+    pass    
+try:
+    import cPickle as pickle
+except:
+    pass
+    
+from ..common import make_logger
+from .igor import Igor
+
+
+sys.setrecursionlimit(40000)
+def compose(*layers):
+    def func(x):
+        out = x
+        for layer in layers[::-1]:
+            out = layer(out)
+        return out
+    return func
+
+class LanguageModel(object):
+    def __init__(self, igor):
+        now = datetime.now()
+        self.run_name = "fergusr_{}mo_{}day_{}hr_{}min".format(now.month, now.day, 
+                                                                now.hour, now.minute)
+        log_location = join(igor.log_dir, self.run_name+".log")
+        self.logger = igor.logger = make_logger(igor, log_location)
+        self.igor = igor
+        
+    @classmethod
+    def from_config(cls, config):
+        igor = Igor(config)
+        igor.prep()
+        model = cls(igor)
+        model.make()
+        return model
+
+
+    def make(self):
+
+        B = self.igor.batch_size
+        R = self.igor.rnn_size
+        S = self.igor.max_sequence_len
+        V = self.igor.vocab_size
+        E = self.igor.embedding_size
+        emb_W = self.igor.embeddings.astype(K.floatx())
+        
+        ## dropout parameters
+        p_emb = self.igor.p_emb_dropout
+        p_W = self.igor.p_W_dropout
+        p_U = self.igor.p_U_dropout
+        p_dense = self.igor.p_dense_dropout
+        w_decay = self.igor.weight_decay
+
+        def embedding_parameters():
+            return {"W_regularizer": l2(w_decay),
+                    "weights": [emb_W],
+                    "mask_zero": True,
[…]

### 2016-10-07 fergus/models/language_model/utils.py

@@ -1,0 +1,89 @@
+import os
+from ikelos.data import Vocabulary 
+try:
+    import cPickle as pickle
+except:
+    import pickle
+
+from collections import Counter
+import json
+
+
+def xor(a,b, v=None):
+    return (a is not v and b is v) or (a is v and b is not v)
+
+def load_dataset(filepath, add_period=True, max_length=float('inf')):
+    f = lambda x: [y.lower() for y in x.replace("\n","").split(" ")] + (["."] if "." not in x[-4:] and add_period else [])
+    stats = [0,0]
+    out = []
+    with open(filepath) as fp:
+        if "pkl" in filepath[-4:]:
+            dataset = pickle.load(fp)
+        elif ".txt" in filepath[-4:]:
+            dataset = fp.readlines()
+        else:
+            raise Exception("unknown dataset file")
+        for sent in dataset:
+            datum = f(sent)
+            if len(datum) <= max_length:
+                stats[0] += 1
+                out.append(datum)
+            else:
+                stats[1] += 1
+        print("data file: {}. {} kept; {} discarded.".format(filepath, stats[0], stats[1]))
+        return out
+
+        # too opaque
+        #return filter(lambda x: len(x) < max_length+2, 
+        #               [[['<START>']+f(x)+["<END>"] for x in pickle.load(fp)] if len(s) > ])
+
+
+def vocab(data, frequency_cutoff=None, size_cutoff=None):
+    if not xor(frequency_cutoff, size_cutoff):
+        raise Exception("one or the other cutoffs please")
+
+    counter = Counter(word for sent in data for word in sent)
+
+    if frequency_cutoff is not None:    
+        print("Using a frequency of {} to reduce vocabulary size.".format(frequency_cutoff))
+        words = [word for word,count in counter.most_common() if count > frequency_cutoff]
+        print("Vocabulary size reduced. {} -> {}".format(len(counter), len(words)))
+        
+    elif size_cutoff is not None:
+        print("Using a cutoff of {} to reduce vocabulary size.".format(size_cutoff))
+        words = [word for word,count in counter.most_common(size_cutoff)]
+        print("Vocabulary size reduced. {} -> {}".format(len(counter), len(words)))
+    
+    else:
+        raise Exception("should never happen...")
+    
[…]

### 2016-10-07 runners/run_linearizer.py

@@ -1,0 +1,135 @@
+"""Linearizing routine
+
+Usage:
+    run_linearizer.py (fergusr|fergusn) (dev|test) (convolutional|token) 
+    run_linearizer.py (-h | --help)
+
+Options:
+    fergusr,fergusn             choose the model, fergus recurrent or fergus neuralized
+    dev,test                    choose the dataset to run on
+    convolution,token           chosen the supertag embedding style
+    
+    
+    
+Notes: 
+McMahan and Stone
+"Syntactic realization with data-driven neural tree grammars" 
+published in COLING 2016
+"""
+from __future__ import absolute_import, print_function, division
+
+import os
+import sys
+here = os.path.dirname(os.path.realpath(__file__))
+sys.path.append(os.path.join(here, '..'))
+import fergus
+
+from fergus.algorithms import linearizer
+from fergus.models import language_model
+from fergus.configs import compose_configs
+from sqlitedict import SqliteDict 
+from tqdm import tqdm
+from docopt import docopt
+import numpy as np
+import editdistance
+import os
+
+
+def run(config, db_name):
+    with SqliteDict(db_name, tablename='tagged_data') as db:
+        data = list(db.items())
+    with SqliteDict(db_name, tablename='linearized_data') as db:
+        finished = set(db.keys())
+
+    data = [datum for datum in data if datum[0] not in finished]
+
+    model = language_model.from_config(config)
+    _safety = 2**32
+    
+    beam = linearizer.decode.beam
+    decode_one = linearizer.utils.decode_one
+    gendist = linearizer.utils.gendist
+    PartialTree = linearizer.partialtree.PartialTree
+    editdist = editdistance.eval
+
+    results = {}
+    bar = tqdm(total=len(data), desc='partial to trellis to decoding')
+    
+    for idx, (datum, datum_str, _) in data:
+        partial = PartialTree.from_list(datum)
+        partial.prep()
+        difficulty = partial.measure_difficulty()
+        model.logger.debug(str(idx)+' Difficulty: '+str(difficulty))
+        if difficulty > 2**35:
+            bad_str = "Skipping.  index={}; difficulty={}".format(idx, difficulty)
+            model.logger.debug(bad_str)
+            bar.update(1)
+            continue
+        
+        seqs, memos = linearizer.dp.run(partial)
+        if len(seqs) == 0:
+            bad_str = "Failure.  index={}; difficulty={}".format(idx, difficulty)
[…]

### 2016-10-07 runners/run_tagger.py

@@ -1,4 +1,4 @@
-"""Supertagging routine for 
+"""Supertagging routine
 
 Usage:
     run_tagger.py (fergusr|fergusn) (dev|test) (convolutional|token) 

### 2016-10-08 fergus/models/fergus_neuralized/model.py
small fix

@@ -82,8 +82,8 @@ class FergusNModel(object):
     @classmethod
     def from_config(cls, config, kwargs=None):
         igor = Igor(config)
-        igor.prep()
         model = cls(igor)
+        igor.prep()
         model.make(kwargs)
         return model
 

### 2016-10-09 runners/run_tagger.py
updating tagger

@@ -1,7 +1,7 @@
 """Supertagging routine
 
 Usage:
-    run_tagger.py (fergusr|fergusn) (dev|test) (convolutional|token) 
+    run_tagger.py (fergusr|fergusn) (dev|test) (convolutional|token|minimaltoken) 
     run_tagger.py (-h | --help)
 
 Options:
@@ -95,11 +95,19 @@ if __name__ == '__main__':
     args = docopt(__doc__, version="Supertagger.  publication version. 2016")
     ### the options in docopt doc were mutually exlusive
     data_type = "dev" if args['dev'] else "test"
-    embed_type = "convolutional" if args['convolutional'] else "token"
+    if args['convolutional']:
+        embed_type = 'convolutional'
+    elif args['token']:
+        embed_type = 'token'
+    elif args['minimaltoken']:
+        embed_type = 'minimaltoken'
+    elif args['shallowconv']:
+        embed_type = 'shallowconv'
+    else:
+        raise Exception("bad embedding argument")
     model_type = "fergusn" if args['fergusn'] else 'fergusr'
     base_name = "{}_{}_{}".format(model_type, embed_type, data_type)
     db_name = base_name+".db"
-    conf_name = base_name+".conf"
     model_factory = {'fergusn': Tagger.fergusn, 'fergusr': Tagger.fergusr}[model_type]
     
     config = compose_configs(data=data_type, model=model_type, embedding=embed_type)

### 2016-10-09 runners/run_linearizer.py
minimal token stuff

@@ -1,7 +1,7 @@
 """Linearizing routine
 
 Usage:
-    run_linearizer.py (fergusr|fergusn) (dev|test) (convolutional|token) 
+    run_linearizer.py (fergusr|fergusn) (dev|test) (convolutional|token|minimaltoken) 
     run_linearizer.py (-h | --help)
 
 Options:
@@ -123,7 +123,16 @@ if __name__ == '__main__':
     args = docopt(__doc__, version="Linearizer.  publication version. 2016")
     ### the options in docopt doc were mutually exlusive
     data_type = "dev" if args['dev'] else "test"
-    embed_type = "convolutional" if args['convolutional'] else "token"
+    if args['convolutional']:
+        embed_type = 'convolutional'
+    elif args['token']:
+        embed_type = 'token'
+    elif args['minimaltoken']:
+        embed_type = 'minimaltoken'
+    elif args['shallowconv']:
+        embed_type = 'shallowconv'
+    else:
+        raise Exception("bad embedding argument")
     model_type = "fergusn" if args['fergusn'] else 'fergusr'
     base_name = "{}_{}_{}".format(model_type, embed_type, data_type)
     db_name = base_name+".db"

### 2016-10-09 paper/draft.tex
numbers added

@@
          &  Token       &  60.69\% & 55.56\% \\ \cline{2-4}
+         & Minimal Token & 52.09\% & 54.18\% \\ \cline{2-4}
 \hline
@@
          &  Token       &  82.65\% & 62.73\% \\ \cline{2-4}
+         &  Minimal Token & 10.13\% & 19.66\% \\ \cline{2-4}
 \hline
@@
 \centering
-\begin{tabular}{|l|p{2cm}|p{1.5cm}|r|}
+\begin{tabular}{|l|p{3cm}|p{2.5cm}|r|}
+\cline{3-4}
+\multicolumn{2}{}{} & \multicolumn{2}{|c|}{Accuracy}   \\ \hline
+\textbf{Model} & \textbf{Embedding}  & \textbf{Top Scoring} & \textbf{Best Performance} \\ \hline
+FERGUS & n.a. & \multicolumn{2}{|c|}{74.9  \%} \\
+\hline \hline
+Fergus-N & Convolution & 65.80 \% & 72.58 \% \\ \cline{2-4}
+         & Token       & 65.52 \%  & 71.82 \% \\ \cline{2-4}
+         & Minimal Token & 63.79 \% & 71.09 \% \\ \cline{2-4}
 \hline
-Model & Embedding & Beam Type & Accuracy \\ \hline
-FERGUS & n.a. & n.a. & 74.9 \% \\
-\hline \hline
-Fergus-N & Convolution & $Top_{all}$  & 68.9\% \\ \cline{3-4}
-         &             & $Best_{all}$ & 76.4\% \\ \cline{2-4}
-         & Token       & $Top_{all}$  & XX\% \\ \cline{3-4}
-         &             & $Best_{all}$ & XX\% \\ \cline{3-4}
-\hline
-Fergus-R & Convolution & $Top_{all}$  & 71.9\% \\ \cline{3-4}
-         &             & $Best_{all}$ & 78.9\% \\ \cline{2-4}
-         & Token       & $Top_{all}$  & XX \% \\ \cline{3-4}
-         &             & $Best_{all}$ & XX \% \\ \cline{3-4}
+Fergus-R & Convolution & 68.22\% & 74.70\% \\ \cline{2-4}
+         & Token       &  69.29 \% & 75.56\% \\ \cline{2-4}
+         & Minimal Token &  58.23 \% & 65.04\% \\ \cline{2-4}
 \hline

### 2016-10-10 runners/run_trainer.py
updating; adding readmes

@@ -1,8 +1,7 @@
-"""An implemenation of FERGUS's supertagger
+"""Training routine for the Fergus-R and Fergus-N models
 
 Usage:
     run_trainer.py (fergusr|fergusn) (convolutional|token|shallowconv|minimaltoken)
-    run_trainer.py fergusn 
     run_trainer.py (-h | --help)
 
     fergusr                   train the fergus-r model
@@ -62,28 +61,3 @@ if __name__ == "__main__":
         print("Some more info: {}".format(sys.exc_info()))
         print(traceback.format_exc())
           
-        # elif args['debug']:
-        #     kwargs = {"mode":NanGuardMode(nan_is_error=True, 
-        #                                   inf_is_error=True, 
-        #                                   big_is_error=True)}
-        #     #stage(igor)
-        #     #model = TrainingModel(igor)
-        #     #model.make(kwargs)
-        #     #model.debug()
-        #     #model.train()
-
-        #     model = TrainingModel.from_yaml(args['<config_file>'], kwargs)
-        #     model.train()
-
-        # elif args['profile']:
-        #     kwargs = {"profile": True}
-        #     stage(igor)
-        #     model = TrainingModel(igor)
-        #     model.make(kwargs)
-        #     model.profile(num_iterations=1)
-
-        # elif args['visualize']:
-        #     stage(igor)
-        #     model= TrainingModel(igor)
-        #     model.make()
-        #     model.plot()

### 2016-10-10 runners/runall.sh

@@ -1,1 +1,0 @@
-python run_tagger.py fergusr test token && python run_tagger.py fergusn test token && python run_tagger.py fergusn test convolutional && python run_tagger.py fergusr convolutional

### 2016-10-10 scripts/supertag/tag_fergusn_convolution.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_tagger.py fergusn test convolutional

### 2016-10-10 scripts/supertag/tag_fergusn_minimaltoken.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_tagger.py fergusn test minimaltoken

### 2016-10-10 scripts/supertag/tag_fergusn_token.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_tagger.py fergusn test token

### 2016-10-10 scripts/supertag/tag_fergusr_convolution.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_tagger.py fergusr test convolutional

### 2016-10-10 scripts/supertag/tag_fergusr_minimaltoken.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_tagger.py fergusr test minimaltoken

### 2016-10-10 scripts/supertag/tag_fergusr_token.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_tagger.py fergusr test token

### 2016-10-10 scripts/train/train_fergusn_convolution.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_trainer.py fergusn convolutional

### 2016-10-10 scripts/train/train_fergusn_minimaltoken.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_trainer.py fergusn minimaltoken

### 2016-10-10 scripts/train/train_fergusn_token.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_trainer.py fergusn token

### 2016-10-10 scripts/train/train_fergusr_convolution.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_trainer.py fergusr convolutional

### 2016-10-10 scripts/train/train_fergusr_minimaltoken.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_trainer.py fergusr minimaltoken

### 2016-10-10 scripts/train/train_fergusr_token.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+cd ../../runners
+
+python run_trainer.py fergusr token

### 2016-10-10 paper/draft.tex
first round of wording changes

@@
 
-%\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
-
 %\setlength\titlebox{5cm}
@@
 
-% \author{Brian McMahan \\
-%   Computer Science, Rutgers University \\
-%   {\tt brian.mcmahan@rutgers.edu} \\\And
-%   Matthew Stone \\
-%   Computer Science, Rutgers University \\
-%   {\tt matthew.stone@rutgers.edu} \\}
+\author{Brian McMahan \and Matthew Stone \\
+   Computer Science, Rutgers University \\
+   {\tt brian.mcmahan, matthew.stone@rutgers.edu}}
 
@@
 combinatorial explosions and address the sparsity of training data.
+%
 By contrast, in this paper, we leverage recent advances in deep
@@
 engineers from many of these decisions, but still generalize more
-effectively and match human choices more closely than traditional
-techniques.
+effectively, match human choices more closely, and enable more
+efficient computations than traditional techniques.
 
 We adopt the characterization of syntactic choice from
-\newcite{bangalore2000exploiting} which uses a stochastic tree model and a language model to produce a linearized string from an unordered,
+\newcite{bangalore2000exploiting}, which uses a stochastic tree model
+and a language model to produce a linearized string from an unordered,
 unlabeled dependency graph.
 %
-The first step to producing a linearized string is
-to assign each node an appropriate supertag---a fragment of a parse tree with
-a leaf left open for a lexical item---using a stochastic tree model.
-%A language model is then used to select among the many possible derivation t
-The resulting assignments are consistent with many possible derivation trees
-because of adjunct ambiguitity.
+The first step to producing a linearized string is to assign each item
+an appropriate \emph{supertag}---a fragment of a parse tree with a
+leaf left open for a lexical item---using a stochastic model.
 %
-To finish the linearization, a language model is used to select among them.
+This involves applying a learned model to make predictions for the
+syntax of each item and then searching over the predictions to find a
+consistent assignment for the entire sentence.
+%
+The resulting assignments allow for many possible surface realization
+outputs because they can underdetermine the order and attachment of
+adjuncts.
+%
+To finish the linearization, a language model is used to select the
+most likely surface form from among the alternatives.
 %
@@
 
-In this work, we make two novel contributions to improve upon the stochastic tree model:
-a supertag embedding technique which uses convolutional neural networks and
-a recurrent tree network which can model hierarchical relationships.
+Our work exploits deep learning to improve the model of supertag
+assignment in two ways.
 %
-By embedding supertags and using a recurrent tree network for the relationships between them,
-we enable more information to be used at each node.
+First, we analyze the use of embedding technques to generalize across
+supertags.  
 %
-We evaluate our contributions in two ways: first, by varying the technique used to embed
-supertags, and then by comparing a feed-forward model against our
-recurrent tree model.
+Neural networks offer a number of architectures that can cluster tree
+fragments during training; such models learn to treat related structures
+similarly, and we show that they improve supertag assignments.
+%
+Second, we analyze the use of tree recurrences to track hierarchical
+relationships within the generation process.
+%
+Such networks can track more of the generation context than a simple
+feed-forward model; as a side effect, they can simplify the problem of
+computing consistent supertag assignments for an entire sentence.
+%
+We evaluate our contributions in two ways: first, by varying the
+technique used to embed supertags, and then by comparing a
+feed-forward model against our recurrent tree model.
 
 Our presentation begins in \ref{sec:tree} with an introduction to tree
-grammars and a deterministic metholodogy for inducing the elementary trees of the grammar.
+grammars and a deterministic metholodogy for inducing the elementary
+trees of the grammar.
 %
-Next, \ref{sec:neural} presents the techniques we have developed to represent
-a tree grammar using a neural architecture.
+Next, \ref{sec:neural} presents the techniques we have developed to
+represent a tree grammar using a neural architecture.
 %
-Then, in \ref{sec:models}, we describe the specific model we have
-implemented, and in \ref{sec:algo}, we describe the algorithms used to describe the classification results.
+Then, in \ref{sec:models}, we describe the specific models we have
+implemented and the algorithms used to exploit the models in NLG.
 %
@@
 We conclude with a brief discussion of the future potential for neural
-architectures to predict grammatical choices in NLG.
+architectures to predict NLG choices.
 
@@
 Broadly, tree grammars are a family of tree rewriting formalisms that
-modify structure to produce strings.  The atomic syntactic units are
-called elementary trees; elementary trees combine using tree-rewrite
-rules to form derived phrase structure trees describing complex
-sentences.  Inducing a tree grammar involves fixing a formal inventory
-of structures and operations for elementary trees and then inferring
+produce strings as a side effect of composing primitive
+hierarchical structures.  The basic syntactic units are called
+elementary trees; elementary trees combine using tree-rewrite rules to
+form derived phrase structure trees describing complex sentences.
+Inducing a tree grammar involves fixing a formal inventory of
+structures and operations for elementary trees and then inferring
 instances of those structures to match corpus data.
@@
 %
-The elementary trees of LTAG consist of two disjoint sets with distinct operations
-%%%
-The elementary trees of LTAG are disjoint sets based on the operations
-they can perform: initial trees can perform substitution operations
+The elementary trees of LTAG consist of two disjoint sets with
+distinct operations: initial trees can perform substitution operations
 and auxiliary trees can perform adjunction operations.
@@
 The adjunction operation modifies the internal structure of a target
-tree at by expanding a node identically-labeled with the root and a
+tree by expanding a node identically-labeled with the root and a
 distinguished foot note in the auxiliary tree.
@@
 
-We adopt one particular restricted adjunction, called
+We adopt one particular restriction on adjunction, called
 sister-adjunction or \emph{insertion}, which allows trees to attach to
@@
 and deterministic rules to induce the grammar
-\cite{chiang2000statistical,bangalore2001impact}.
+\cite{bangalore2001impact,chiang2000statistical}.
 %
@@
 trees that combine with the parent node, either to the right or to the
-left of the head; this allows us to segment the tree into units of
+left of the head.  This allows us to segment the tree into units of
 substitution and insertion.%
@@
 %
-First, supertags are represented as vectors through convolutional coding,
-opening up the recent repertoire of vector space techniques.
+First, supertags are represented as vectors through embedding
+techniques that enable to model to generalize naturally over complex, but
+related structures.
 %
-Second, hierarchical dependence between nodes are modeled using a recurrent tree
-network.
+Second, hierarchical dependence between choices is modeled using a recurrent tree
+network that can capture long-distance influences as well as local ones.
 %
@@
 %We embed supertags by embed each of its nodes, combine
-To embed a supertag, each of its nodes are embedded, grouped to form a tensor, and then summarized into a single vector using a series of convolutional neural networks.
+To embed a supertag, we embed each node, group the resulting vectors
+to form a tensor, and then summarize the tensor into a single vector
+using a series of convolutional neural networks.
 %
-Since the weights of a convolutional neural network are optimized during training, the data should determine which parts of the supertag are important.
+Since the weights of a convolutional neural network are optimized
+during training, the data should determine which parts of the supertag
+are important.
 %
@@
 
-The goal of embedding supertags is to summarize the tree structure into a single vector.
+To explain our technique, we begin by observing that each elementary
+tree is a complex structure with nodes labeled by category and
+assigned a role that enables further tree operations.
 %
-Inside the tree structure, there are two separate pieces of information which need to be summarized:
-the syntactic category of each node and the role each node plays in the tree grammar.
+The root node's role represents the overall action associated with
+that elementary tree---either substitution or insertion.
 %
-The root receives a special role---either substitution or insertion to represent
-the supertag's tree operation.
-%g
-The remaining nodes either have the substitution point role or the spine role---they are along the spine from root to the lexical attachment point.
+The remaining nodes either have the substitution point role or the
+spine role---they are along the spine from root to the lexical
+attachment point, and thus provide targets for further insertion.
 
@@
 %
-Each parent-child pair are treated as independent of all other pairs.
+Each parent-child pair is treated as independent of all others.
 %
@@
 %
-The search concludes either when a single complete tree has been found\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}.
+The search concludes either when a single complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}
 
@@
 
+\section*{Acknowledgments}
+
+This research was supported in part by NSF IIS-1526723 and by a
+sabbatical leave from Rutgers to Stone.
+
 \bibliography{coling2016}

### 2016-10-10 paper/draft.tex
more wording and flow

@@
 %
-Our methods also embed elementary tree structures using convolutional
-coding to generalize over tree structures.
+We also induce embeddings to generalize over elementary tree
+structures and exploit a tree recurrence over the input structure to
+model long-distance influences between NLG choices.
 %
 We evaluate the models on the task of linearizing unannotated
-dependency trees, achieving improvements over previous methods.
+dependency trees, documenting the contribution of our modeling
+techniques to improvements in both accuracy and run-time.
 
@@
 %
-This involves applying a learned model to make predictions for the
-syntax of each item and then searching over the predictions to find a
-consistent assignment for the entire sentence.
+This process involves applying a learned model to make predictions for
+the syntax of each item and then searching over the predictions to
+find a consistent assignment for the entire sentence.
 %
@@
 
-\subsection{The Grammar Formalism}
+\subsection{Grammar Formalism}
 
@@
 only substitutions \cite{cohn2009inducing}.
-
+%
 We adopt one particular restriction on adjunction, called
@@
 
-\subsection{Inducing a tree grammar}
+\subsection{Grammar Induction}
 
@@
 
-%There is a tension between vectorized representations and dynamic, nonlinear data structures.
+The grammar induction of \ref{sec:tree} allows us to constrcut an
+inventory of supertags to match a corpus.  For NLG, we also need to
+predict how likely a supertag is to fit a specific lexical item in the
+context of a generation task.  We approach this problem using neural
+networks. 
 %
-%The tension is emphasized by requirement of specialized hardware that both dramatically increases the data throughput and reduces the space of representable models\footnote{Specifically, to run on Graphical Processing Units (GPUs), model structure must be known at compile time or incur heavy computational costs.}.
+In particuaar, this work makes two contributions to improve stochastic
+tree modeling with neural networks.
 %
-This work makes two contributions to improve stochastic tree modeling with neural networks.
-%This work makes two contributions to reduce that tension and progress methodologies for representing nonlinear data structures in linear spaces.
+First, we represent supertags as vectors through embedding techniques
+that enable to model to generalize naturally over complex, but related
+structures.
 %
-First, supertags are represented as vectors through embedding
-techniques that enable to model to generalize naturally over complex, but
-related structures.
+Second, we address the hierarchical dependence between choices using a
+recurrent tree network that can capture long-distance influences as
+well as local ones.
 %
-Second, hierarchical dependence between choices is modeled using a recurrent tree
-network that can capture long-distance influences as well as local ones.
-%
-%recurrent neural network variant that isn't restricted to conditioning hidden states on their temporal predecessors.
-%In brief, this technique is made possible through recent advanced in StackLSTMs \cite{dyer2015transition} and the use of topological indexing of recurrent states.% hidden states of
+We now describe these representations in more detail.
 
@@
 
-%The first step to embedding supertags with a convolutional neural network is to
-%embed each of its nodes and combine them to form a tensor.
-%
-%
-%To embed a supertag, we first embed all of its nodes to form a tensor and then use a
-%To embed a supertag and summarize it into a vector, each of its nodes are
-%first embedded, combined to form a tensor, and then
-%We embed supertags by embed each of its nodes, combine
-To embed a supertag, we embed each node, group the resulting vectors
-to form a tensor, and then summarize the tensor into a single vector
-using a series of convolutional neural networks.
-%
-Since the weights of a convolutional neural network are optimized
-during training, the data should determine which parts of the supertag
-are important.
-%
-%The purpose of convolutional coding is to allow the data to determine which facets of a supertag are relevant and useful.
-%
-The more usual alternative approach is to treat each supertag as the atomic
-unit---essentially a statistical singleton---and map it either to a feature or
-an embedding vector.
-%
-We evaluate our contribution and compare these two approaches to get a sense for whether convolutional coding robustly generalizes across supertags.
+Different supertags for the same word can encode differences in the
+item's own combinatory syntax, differences in argument structure, and
+differences in word order.  Accordingly, words have many related
+supertags, with substantial overlaps in structure, and, presumably,
+corresponding similarities in their patterns of occurrence.  A
+traditional machine learning approach to supertag prediction would
+treat individual supertags as atoms for classification; geenralizing
+across supertags would require linking model parameters to handcrafted
+features or backoff categories.
 
-To explain our technique, we begin by observing that each elementary
-tree is a complex structure with nodes labeled by category and
-assigned a role that enables further tree operations.
+By contrast, neural techniques work by embedding such tokens into a
+vector space. This process learns an abstract representation of tokens
+that clusters similar items together and makes further predictions
+as a function of those items' learned features.  The resulting
+abilities to generalize across sparse data seems to be one of the most
+improtant reasons for the success of deep learning in NLP.
+
+The simplest way to embed supertags is simply to treat each structure
+as distinct token that indexes a corresponding learned vector.  This
+places no constraints on the learned similarity function, but it also
+ignores the hieararchical structure of the elementary trees
+themselves.  Previous work on deep learning with tree structures has
+suggested that convolutional coding can produce embeddings that
+directly exploit similarities in tree structure.  Thus, we developed
+analogous techniques to encode supertags based on their underlying
+tree structure.  In partiuclar, to embed a supertag, we embed each
+node, group the resulting vectors to form a tensor, and then summarize
+the tensor into a single vector using a series of convolutional neural
+networks.
+
+Note that each elementary tree is a complex structure with nodes
+labeled by category and assigned a role that enables further tree
+operations.
 %
@@
 %
-This is visualized in on the right Figure \ref{fig:spineembedding}.
+This is visualized in on the right in Figure \ref{fig:spineembedding}.
 %
@@
 
-To model the relationship between the supertags in the derivation tree, we use
-a recurrent tree network.
+Our models predict supertags as a function of the target word and its
+context.  Neural networks make it possible to generalize over such
+contexts by learning to represent them with a hidden state vector that
+aggregates and clusters information from the relevant history.  Our
+approach is to do this using a recurrent tree network.
 %
@@
 
-In this work, we present two stochastic tree models which represent different modeling decisions.
+To analyze the representations we describe in \ref{sec:tree} and
+\ref{sec:neural}, we developed two alternative architectures for
+predicting supertags in context.
 %
-The first, Fergus-N\footnote{Fergus-N is shorthand for Fergus-Neuralized}, is a
-reimplementation of \newcite{bangalore2000exploiting}'s original FERGUS model but
-implemented in a feed-forward neural network.
+The first is a feed-forward neural network designed to solve a closely
+analogous task to the supertagging step of
+\newcite{bangalore2000exploiting}'s original FERGUS model.
 %
-The second model, Fergus-R\footnote{Fergus-R is shorthand for Fergus-Recurrent}, utilizes a recurrent tree network.
+We call it Fergus-N (for Neuralized).
 %
-Fergus-N serves as a baseline model to which we compare the recurrent tree network.
+The second uses a recurrent tree network to model the generation
+context.
+%
+Because it has this richer context representation, it takes advantage
+of a slightly different chracterization of the supertag prediction
+problem to streamline the problem solving involved in using the model.
+%
+We call this Fergus-R (for Recurrent).
 
@@
 %
-The probability of the parent's supertag is computed using an embedding of the pair's lexical material and an embedding of the child's supertag.
+The probability of the parent's supertag is predicted using an embedding of the pair's lexical material and an embedding of the child's supertag.
 %
-During training, the negative log probability of the observed parent supertags is minimized for each parent-child pair.
+(Our experiments compare the different embedding options surveyed in
+\ref{sec:neural}.) 
 %
-This objective function is formally defined in Eq.
-\ref{eq:fergusn_obj} where $tag_p$ is the parent supertag, $tag_c$ is the child supertag, $lex_p$ is the parent's lexical material, and $lex_c$ is the child's lexical material.
+Training maximizes the likelihood of the training data according to
+the model.
+%
+Formally, our objective is to minimize the negative log probability of
+the observed parent supertags for each parent-child pair, as formally
+defined in Eq. \ref{eq:fergusn_obj}.
+%
+\begin{align}
+&min_{\theta} -[\sum_p\sum_{p\to c} log[P_\theta(tag_{p} | lex_{p}, lex_{c}, tag_{c})] + \sum_c log[P_\theta(tag_{c} |lex_{p}, lex_{c})]] \label{eq:fergusn_obj}
+\end{align}
+%
+Here $tag_p$ is the parent
+supertag, $tag_c$ is the child supertag, $lex_p$ is the parent's
+lexical material, and $lex_c$ is the child's lexical material.
 %
@@
 
-\begin{align}
-&min_{\theta} -[\sum_p\sum_{p\to c} log[P_\theta(tag_{p} | lex_{p}, lex_{c}, tag_{c})] + \sum_c log[P_\theta(tag_{c} |lex_{p}, lex_{c})]] \label{eq:fergusn_obj}
-\end{align}
-
-The model is implemented as a feed-forward neural network that computes a probability distribution over parent supertags conditioned on lexical material and the child's supertag.
+The model is implemented as a feed-forward neural network.
 %
@@
 
-A decoding step uses the probability distributions to compute a high probability assignment for all supertags simueltaneously.
+At generation time, we are given a full dependency tree.  A decoding
+step is necessary to compute a high probability
+assignment for all supertags simultaneously.
 %
@@
 For each node in the input dependency tree, soft attention---a method which learns a vectorized function to weight a group of vectors and sum into a single vector---is used to summarize its children.
-%
-
-The advantage of using a recurrent tree is that nodes are able to be informed of their ancestors.
-%
-Each node is further informed using soft attention over its children ---a method which learns a vectorized function to weight a group of vectors and sum into a single vector.
 %

### 2016-10-10 paper/draft.tex
adding more numbers, appendix

--- 2016-10-09 paper/draft.tex
@@
 
+\begin{figure}
+\centering
+\begin{tabular}{|l|p{2cm}|c|c|c|}
+\cline{3-4}
+\multicolumn{2}{}{}& \multicolumn{2}{ |c| }{ \textbf{Accuracy}}&\multicolumn{1}{}{} \\ \hline
+\textbf{Model} & \textbf{Embedding}  & \textbf{Before A*} 
+& \textbf{After A*} & Running Time \\ \hline
+Fergus-N &  Convolution &  58.17\% & 57.40\%  & 1.97s \\ \cline{2-5}
+         &  Token       &  60.69\% & 55.56\% & 1.81s \\ \cline{2-5}
+         & Minimal Token & 52.09\% & 54.18\% & 2.02s \\ \cline{2-5}
+\hline
+Fergus-R &  Convolution &  67.62\% & 57.04\% & 0.30s \\ \cline{2-5}
+         &  Token       &  82.65\% & 62.73\% & 0.36s\\ \cline{2-5}
+         &  Minimal Token & 10.13\% & 19.66\% & 0.54 \\ \cline{2-5}
+\hline
+\end{tabular}
+\label{table:accresults}
+\caption{For each supertag and embedding pair, the mean accuracy of supertag classification before and after the A* algorithm is shown.
+Also shown is the median running time---which includes model computation and A* search.}
+\end{figure}
+
 Table \ref{table:accresults} shows the results of supertag classification using the stochastic tree models is shown. 
@@
 The Fergus-N model out-performed the Fergus-R model on supertag classification with convolutional supertag embeddings, but the Fergus-R model performed much closer to that of Fergus-N on the individual components.
-
-\begin{figure}
-\centering
-\begin{tabular}{|l|p{2cm}|c|c|}
-\cline{3-4}
-\multicolumn{2}{}{}& \multicolumn{2}{ |c| }{ \textbf{Accuracy}} \\ \hline
-\textbf{Model} & \textbf{Embedding}  & \textbf{Before A*} & \textbf{After A*}  \\ \hline
-Fergus-N &  Convolution &  58.17\% & 57.40\% \\ \cline{2-4}
-         &  Token       &  60.69\% & 55.56\% \\ \cline{2-4}
-         & Minimal Token & 52.09\% & 54.18\% \\ \cline{2-4}
-\hline
-Fergus-R &  Convolution &  67.62\% & 57.04\% \\ \cline{2-4}
-         &  Token       &  82.65\% & 62.73\% \\ \cline{2-4}
-         &  Minimal Token & 10.13\% & 19.66\% \\ \cline{2-4}
-\hline
-\end{tabular}
-\label{table:accresults}
-\caption{Classification accuracy by supertag and its individual properties}
-\end{figure}
 
@@
 
+\newpage
 \appendix
+
+\section{Appendix}
+
+\begin{figure}
+\centering
+\begin{tabular}{|p{6cm}|r|}
+\hline
+Model Parameter & Value \\ \hline
+Convolution filter size & 48 \\ \hline
+Optimization Algorithm & ADAM \\ \hline
+Learning Rate & 0.0001 \\ \hline
+Dense-Layer Dropout Rates & 0.5 \\ \hline
+\end{tabular}
+\label{table:params}
+\caption{.}
+\end{figure}
+
+\subsection{Data}
+
+This work used data processed from the Wall Street Journal portion of the Treebank corpus.
+Specifically, section 02-21 were used for training, section 22 for development, and section 23 for testing.
+The corpus was preprocessed using the Stanford Parser to fix some common issues as well as remove information extra to syntactic category.
+The resulting parse trees were processed.
+They were annotated with head rules that marked the head at each node in the parse tree and which nodes were dependents or adjuncts.
+These annotated trees were split at adjunction and substitution positions to form the grammar.
+For the dependency structures used in the models, we use the derivation tree formed by the grammar for each parse tree.
+The derivation structures were further processed to parent-child pairs and top-down traversals for each of the models.
+The preprocessing and processing code are distributed with this paper.
+
+\subsection{Model parameters}
+
+\end{document}
+
+For the convolutional embeddings, we used
 
@@
     - where the data comes from
-
-\end{document}

### 2016-10-10 fergus/data_pipeline/1_parse/run_stanford_coordination.py

@@ -1,6 +1,6 @@
 import subprocess
 
-command = "~/research/code/stanford-parser/bcm_transformer.sh merged/{0} > sfprocced/{0}"
+command = "~/research/code/stanford-parser/bcm_transformer.sh merged/{0} > preprocessed/{0}"
 
 import glob
 

### 2016-10-10 paper/draft.tex
minor text fixes in main body; fleshing out appendix; converted talbe figures to tables

@@
 %
-An example of a recurrent tree network is shown in \ref{fig:rtn}.
+An example of a recurrent tree network is shown in Figure \ref{fig:rtn}.
 
@@
 \subsection{Results}
+\label{sec:results}
 
-\begin{figure}
+\begin{table}
 \centering
@@
          &  Token       &  82.65\% & 62.73\% & 0.36s\\ \cline{2-5}
-         &  Minimal Token & 10.13\% & 19.66\% & 0.54 \\ \cline{2-5}
+         &  Minimal Token & 10.13\% & 19.66\% & 0.54s \\ \cline{2-5}
 \hline
 \end{tabular}
-\label{table:accresults}
 \caption{For each supertag and embedding pair, the mean accuracy of supertag classification before and after the A* algorithm is shown.
 Also shown is the median running time---which includes model computation and A* search.}
-\end{figure}
+\label{table:accresults}
+\end{table}
 
@@
 
-\begin{figure}
+\begin{table}
 \centering
@@
 \end{tabular}
+\caption{Shown above as accuraccy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
 \label{table:linresults}
-\caption{Shown above as accuraccy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
-\end{figure}
+\end{table}
 
@@
 \section{Appendix}
+\label{sec:appendix}
 
-\begin{figure}
+\begin{table}[t!]
 \centering
-\begin{tabular}{|p{6cm}|r|}
+\begin{tabular}{lr}
+\begin{tabular}[t]{|p{6cm}|r|}
 \hline
-Model Parameter & Value \\ \hline
+\textbf{Model Parameter} & \textbf{Value} \\ \hline
+\hline \multicolumn{2}{|c|}{\textbf{Fergus-N Parameters}} \\ \hline
+Fully connected layer size & 256 \\ \hline
+Batch size & 128 \\ \hline
+\hline  \multicolumn{2}{|c|}{\textbf{Fergus-R Parameters}} \\ \hline
+Fully connected layer size & 256 \\ \hline 
+Hidden state size & 128 \\ \hline
+Batch size & 16 \\ \hline
+\hline \multicolumn{2}{|c|}{\textbf{Embedding Parameters}} \\ \hline
 Convolution filter size & 48 \\ \hline
+Syntactic category embedding size & 32 \\ \hline
+Node role embedding size & 32 \\ \hline
+Word embedding size \newline
+\cite{pennington2014glove} & 300 \\ \hline
+\end{tabular}
+&
+\begin{tabular}[t]{|p{6cm}|r|}
+\hline
+\textbf{Model Parameter} & \textbf{Value} \\ \hline
+\hline \multicolumn{2}{|c|}{\textbf{Language Model Parameters}} \\ \hline
+Hidden state size & 368 \\ \hline
+Batch size & 32 \\ \hline
+\hline \multicolumn{2}{|c|}{\textbf{Optimization Parameters}} \\ \hline
 Optimization Algorithm & ADAM \\ \hline
-Learning Rate & 0.0001 \\ \hline
-Dense-Layer Dropout Rates & 0.5 \\ \hline
+Fergus-R and Fergus-N Learning Rate & 1e-4 \\ \hline
+Language Model Learning Rate & 0.01 \\ \hline
+Fully-Connected Dropout Rate & 0.5 \\ \hline
+Recurrent Weight Dropout Rate & 0.2 \\ \hline
+$L_2$ Weight Decay & 1e-6 \\ \hline
+Max gradient norm & 10.0 \\ \hline
+Gradient clip threshold & 5.0 \\ \hline
 \end{tabular}
-\label{table:params}
-\caption{.}
-\end{figure}
+\end{tabular}
+\caption{The parameters for the Fergus-R, Fergus-N, and language models.  The
+exact specifications in configuration files can be found in the code repository that accompanies this paper.}
+\label{tab:params}
+\end{table}
 
@@
 \subsection{Model parameters}
+
+All of the models were implemented in the Keras \cite{chollet2015keras} and Theano \cite{theano} libraries.
+The specific parameters that were used are shown in Table \ref{tab:params}.
+The parameters were selected by measured performance on the development portion of the dataset.
+In the accompanying code repository, the full experiment parameters---including progammatic parameters controlling the experimental design---are specified in configuration files.
 

### 2016-10-10 paper/draft.tex
results rewritten

@@
 %
-An example of a recurrent tree network is shown in Figure \ref{fig:rtn}.
+An example of a recurrent tree network is shown in eFigure \ref{fig:rtn}.
 
@@
 
-In this work, we make two contributions to stochastic tree modeling and evaluate them on a supertag prediction and linearization task. 
+Using the representations of \ref{sec:neural}, the models of
+\ref{sec:models} can be instantiated in six different ways.  We can
+use a feed-forward Fergus-N architecture or a recurrent Fergus-R
+architecture.  Each architecture can embed supertags \emph{minimally},
+by a simple one-hot vector; \emph{atomically}, by learning an
+embedding vector corresponding to each supertag; or
+\emph{structureally}, by using a convolutional coding over the
+supertag tree structure.
+
+\subsection{Training}
+
+We trained six such models using a common experimental plantform.  We
+started from the Wall Street Journal sections of the Penn Treebank,
+which have been previously used for evaluating statistical tree
+grammars \cite{chiang2000statistical}\footnote{A possible additional
+  data source, the data from the 2011 Shared Task on Surface
+  Realization, was not available}.
 %
-The evaluation uses the Wall Street Journal sections of the Penn Treebank which have been previously used for evaluating statistical tree grammars \cite{chiang2000statistical}\footnote{A possible additional data source, the data from the 2011 Shared Task on Surface Realization, was not available}.
+Our data pipeline breaks each sentence in the treebank into component
+elementary trees and then represents the sentence in terms of a
+derivation tree, specifying the tree-rewriting operations required to
+construct the actual treebank surface tree from the basic supertags.
 %
-Our data pipeline transforms the treebank into a collection of elementary trees and derivation tree---the structured relationships required to compose them into the observed data (see \ref{}).
+Abstracting syntactic information from the derivation tree leads to
+the unlabeled dependency trees our models assume as input.
+
+From this input, we extracted the atomic supertag prediction instances
+and trained a network defined by each of the architectures of
+\ref{sec:models} and each of the supertag representations of
+\ref{sec:neural}.  As always, we used Section XXX for development,
+Sections XXX for training, and Section XXX for testing.  A complete
+description of network organization and training parameters is given
+in the appendix.  Our code is available XXX.
+
+\subsection{Performance Metrics}
+
+We evaluate the performance of the models in several ways.  First, we
+look at the accuracy of the supertag predictions directly output by
+each model.  Second, we look at the accurace of the final supertags
+obtained by decoding the model predictions to the best-ranked
+consistent global assignment.  These metrics directly assess the
+ability of the models to successfully learn the target distributions.
+
+Next, we evaluate the models on the full NLG task, including
+linearization.  
 %
-Abstracting syntactic information from the derivation tree structure leads to the unlabeled dependency trees our models assume as input.
+The linearization tasks licenses more freedom in supertag
+classifications because it may have the same root and leaf syntactic
+categories but differ in minor constructions, such as number of
+arguments.
+%
+The freedom allows for models to be more generalized and less fit to
+the specific supertags, but at the same time, it also mutes the
+distinctions between classification decisions.
+%
+We report XXXMEASURE, following \cite{}.
+%
+They argue that their measure correlates with NLG performance better
+than n-gram counting methods.
+%
+Since linearization uses a beam search, we report statistics both for
+the top-ranked beam and for the empirically based beam among the
+candidates computed during search.
+%
+The difference gives an indication of the effect of the language model
+in guiding the decisions that remain after supertagging.
 
-We evaluate the convolutional coding of the supertags by comparing it against standard embedding techniques.
+Finally, we report statistics about the run time of different
+generation steps.
 %
-Specifically, we construct a matrix that has as many rows as supertags and is initialized with a Glorot uniform distribution.
-%
-Fergus-N and Fergus-R are retrained with the exact same parameters, except that instead of the supertag indices referencing a matrix constructed by a series of convolutions, it references our constructed embedding matrix (which gets optimized during learning).
-%
-The impact of the embedding will be different for each model---Fergus-N uses embedded supertags in constructing the node state and Fergus-R does not---but any differences between the models and their alternate embedding counterparts can be perceived as evidence for either embedding technique.
-
-To evaluate whether a recurrent tree network can capture hierarchical and long-distance relationships, we compare it against a feed-forward model. 
-%
-Specifically, for each model, we measure how often it predicts the correct supertag or if it predicts a different supertag but with the correct root syntactic category, root node role, or lexical leaf node\footnote{The leaf to which a lexical item would attach} syntactic category.
-%
-These four quantities represent different aspects that are important to get right in a stochastic tree model.
-
-%Fergus-R's predictions never observe its neighboring nodes' supertags, so any predictive improvements are based solely on lexical and ancestral context.
-
-We also compare the differences in supertag embeddings and stochastic tree models on the linearization task itself.
-%
-The linearization tasks licenses more freedom in supertag classifications because it may have the same root and leaf syntactic categories but differ in minor constructions, such as number of arguments.
-%
-The freedom allows for models to be more generalized and less fit to the specific supertags, but at the same time, it also mutes the distinctions between classification decisions.
-%
-The metrics used in the linearization task were taken from previous work, who argue that it correlates with generation performance better than n-gram counting methods.
+This allows us to assess the complexity of the different decoding
+steps involved in generation, to reveal any tradeoffs among the models
+between speed and accuracy.
 
@@
 \multicolumn{2}{}{}& \multicolumn{2}{ |c| }{ \textbf{Accuracy}}&\multicolumn{1}{}{} \\ \hline
-\textbf{Model} & \textbf{Embedding}  & \textbf{Before A*} 
-& \textbf{After A*} & Running Time \\ \hline
-Fergus-N &  Convolution &  58.17\% & 57.40\%  & 1.97s \\ \cline{2-5}
-         &  Token       &  60.69\% & 55.56\% & 1.81s \\ \cline{2-5}
-         & Minimal Token & 52.09\% & 54.18\% & 2.02s \\ \cline{2-5}
+\textbf{Model} & \textbf{Embedding}  & \textbf{Raw Model} 
+& \textbf{After Decoding} & \textbf{Running Time} \\ \hline
+Fergus-N &  Structural  &  58.17\% & 57.40\%  & 1.97s \\ \cline{2-5}
+         &  Atomic      &  60.69\% & 55.56\% & 1.81s \\ \cline{2-5}
+         &  Minimal     &  52.09\% & 54.18\% & 2.02s \\ \cline{2-5}
 \hline
-Fergus-R &  Convolution &  67.62\% & 57.04\% & 0.30s \\ \cline{2-5}
-         &  Token       &  82.65\% & 62.73\% & 0.36s\\ \cline{2-5}
-         &  Minimal Token & 10.13\% & 19.66\% & 0.54s \\ \cline{2-5}
+Fergus-R &  Structural &  67.62\% & 57.04\% & 0.30s \\ \cline{2-5}
+         &  Atomic     &  82.65\% & 62.73\% & 0.36s\\ \cline{2-5}
+         &  Minimal    & 10.13\% & 19.66\% & 0.54s \\ \cline{2-5}
 \hline
 \end{tabular}
-\caption{For each supertag and embedding pair, the mean accuracy of supertag classification before and after the A* algorithm is shown.
-Also shown is the median running time---which includes model computation and A* search.}
+\caption{For each supertag and embedding pair, the mean accuracy of
+  supertag classification directly output by the model and in the
+  consistent global assignment output by A* decoding. Also shown is
+  the median running time---which includes model computation and A*
+  search.}
 \label{table:accresults}
@@
 
-Table \ref{table:accresults} shows the results of supertag classification using the stochastic tree models is shown. 
+Table \ref{table:accresults} shows the results of supertag
+prediction.
 %
-The scores are grouped in first by the model and then by the embedding method. 
+All differences between model are significant using a Paired-Sample t-test ($p<10^{-5}$)
 %
-The accuracy of each model is additionally broken down by how often it got the various components correct. 
+The flexible embedding methods perform consistently better,
+suggesting that the clustering capabilities of neural methods is a
+crucial part of their effectiveness.
 %
-The root category and role refer to the syntactic label and grammatical role of a supertag. 
+The convolution encoding seems to work better with Fergus-N.  
 %
-The leaf category refers to the syntactic label of the leaf to which a lexical item would attach---essentially, this is the part of speech of the lexical item.
+This merits further investigation: it might be because Fergus-N
+predicts one supertag as a function of another, and so the
+compositional relationships among the two trees matters more---or
+because Fergus-R's contextualized decisions depend on similarities
+among supertags (involving argument structure or information
+structure) that are difficult for the convolutional coding to
+represent or learn.
+
+The overall best results come from Fergus-R, suggesting that it is
+worthwhile to take additional context into account in this task.
 %
-All differences between model are significant using a Paired-Sample t-test ($p<10^{5}$)
+At the same time, the median time taken to classify and decode a
+sentence with Fergus-R is just one sixth that of Fergus-N.
+%
+We suspect that there is a general lesson in this speedup: because
+neural models can be more flexible about the information they take
+into account in decisions, it's especially important in designing
+neural architectures to break a problem down into decisions that can
+be combined easily.
 
-The Fergus-N model out-performed the Fergus-R model on supertag classification with convolutional supertag embeddings, but the Fergus-R model performed much closer to that of Fergus-N on the individual components.
+Finally, decoding the network generally leads to lower accuracy.
+%
+It seems that our models are not doing a good job of using the
+predictions it makes to triangulate to accurate and consistent
+supertags. 
+%
+This suggests that the models could be improved by taking more or
+better information into account in decoding.
 
@@
 \textbf{Model} & \textbf{Embedding}  & \textbf{Top Scoring} & \textbf{Best Performance} \\ \hline
-FERGUS & n.a. & \multicolumn{2}{|c|}{74.9  \%} \\
-\hline \hline
-Fergus-N & Convolution & 65.80 \% & 72.58 \% \\ \cline{2-4}
-         & Token       & 65.52 \%  & 71.82 \% \\ \cline{2-4}
-         & Minimal Token & 63.79 \% & 71.09 \% \\ \cline{2-4}
+Fergus-N & Convolution & 65.80\% & 72.58\% \\ \cline{2-4}
+         & Token       & 65.52\%  & 71.82\% \\ \cline{2-4}
+         & Minimal Token & 63.79\% & 71.09\% \\ \cline{2-4}
 \hline
 Fergus-R & Convolution & 68.22\% & 74.70\% \\ \cline{2-4}
-         & Token       &  69.29 \% & 75.56\% \\ \cline{2-4}
-         & Minimal Token &  58.23 \% & 65.04\% \\ \cline{2-4}
+         & Token       &  69.29\% & 75.56\% \\ \cline{2-4}
+         & Minimal Token &  58.23\% & 65.04\% \\ \cline{2-4}
 \hline
 \end{tabular}
-\caption{Shown above as accuraccy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
+\caption{Shown above as accuracy is the percentage of tokens in the linearized strings that are in correct positions according to an edit distance measure.}
 \label{table:linresults}
 \end{table}
+
+Figure~\ref{table:linresults} shows the NLG evaluation results for the
+different models.
+%
+The performance confirms our expectation that differences in supertag
+accuracy after decoding correlate with NLG accuracy overall, but that
+differences in NLG performance are attenuated.
+%
+We note by comparison that Bangalore and Rambow report an accuracy of
+74.9\% in their best evaluation of FERGUS---on a data set of just 100
+sentences with an average lenth of XXX.
+%
+Our evaluation, on XXX sentences with an average length of XXX, is much
+more strenuous.
 

### 2016-10-10 paper/draft.tex
commiting to merge

--- 2016-10-10 paper/draft.tex
@@
 %
-In particuaar, this work makes two contributions to improve stochastic
+In particular, this work makes two contributions to improve stochastic
 tree modeling with neural networks.
@@
 traditional machine learning approach to supertag prediction would
-treat individual supertags as atoms for classification; geenralizing
+treat individual supertags as atoms for classification; generalizing
 across supertags would require linking model parameters to handcrafted
@@
 &P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
-\frac{exp(\Psi(\omega_{treelet_i})))}
-{\sum_{j \in |tagset_p|} exp(\Psi(\omega_{treelet_j})))} \nonumber \\
+\frac{exp(\Psi_{predict}(\omega_{treelet_i})))}
+{\sum_{j \in |tagset_p|} exp(\Psi_{predict}(\omega_{treelet_j})))} \nonumber \\
 &P_\theta(tag_{c,i} |lex_{p}, lex_{c}) = 
-\frac{exp(\Psi(\omega_{leaf_i})))}
-{\sum_{j \in |tagset_c|} exp(\Psi(\omega_{leaf_j})))} \nonumber 
+\frac{exp(\Psi_{predict}(\omega_{leaf_i})))}
+{\sum_{j \in |tagset_c|} exp(\Psi_{predict}(\omega_{leaf_j})))} \nonumber 
 \end{align}
@@
 \centering
-\begin{tabular}{|l|p{2cm}|c|c|c|}
+\begin{tabular}{|l|p{3cm}|c|c|c|}
 \cline{3-4}
@@
 
+In this appendix, we describe some design decisions that took place which did not have a place in the paper but should still be reported.
+In particular, this section covers the data processing, parameters used in the final model, and design decisions for the restricting the supertag distributions.
+
 \subsection{Data}
@@
 In the accompanying code repository, the full experiment parameters---including progammatic parameters controlling the experimental design---are specified in configuration files.
+
+\subsection{Model design}
+
+To facilitate efficient model computation, we restricted the computation for supertag distributions and used vectorized functions.
+%
+The set of supertags that are retrieved for probability distribution computation are set of supertags that are seen paired with the lexical item, reducing the output distribution from over 5000 to under 500.
+%
+This was implemented as a set of indices to the embedding matrix passed into the computation with the rest of the data. 
+%
+Future implementations will explore having the supertag subset lookup table be part of the computation.
+%
+While there are several ways the affinity between supertag and lexical state vector could have been computed, we chose to concatenate the vectors, map them to a new space using a fully connected layer, and compute a score with a vectorized function.
+%
+The vectorized function operation is the same mechanism which calculates the probability distribution used in soft attention.
 

### 2016-10-10 paper/draft.tex
conclusion done

--- 2016-10-10 paper/draft.tex
@@
 dependency trees, documenting the contribution of our modeling
-techniques to improvements in both accuracy and run-time.
+techniques to improvements in both accuracy and run time.
 
@@
 We adopt the characterization of syntactic choice from
-\newcite{bangalore2000exploiting}, which uses a stochastic tree model
-and a language model to produce a linearized string from an unordered,
-unlabeled dependency graph.
+\newcite{bangalore2000exploiting}: the problem is to use a stochastic
+tree model and a language model to produce a linearized string from an
+unordered, unlabeled dependency graph.
 %
@@
 an appropriate \emph{supertag}---a fragment of a parse tree with a
-leaf left open for a lexical item---using a stochastic model.
+leaf left open for the lexical item.
 %
@@
 While improving the language model would improve the linearized string,
-this work focuses on more accurately predicting the correct supertags from
+we focus here on more accurately predicting the correct supertags from
 unlabeled dependency trees.
@@
 %
-First, we analyze the use of embedding technques to generalize across
+First, we analyze the use of embedding techniques to generalize across
 supertags.  
@@
 Our presentation begins in \ref{sec:tree} with an introduction to tree
-grammars and a deterministic metholodogy for inducing the elementary
+grammars and a deterministic methodology for inducing the elementary
 trees of the grammar.
@@
 %
-This means inducing a tree grammar from a dataset is a matter of inferring the set of constructions in the data.
+This means inducing a tree grammar from a data set is a matter of inferring the set of constructions in the data.
 
@@
 
-The grammar induction of \ref{sec:tree} allows us to constrcut an
+The grammar induction of \ref{sec:tree} allows us to construct an
 inventory of supertags to match a corpus.  For NLG, we also need to
@@
 Different supertags for the same word can encode differences in the
-item's own combinatory syntax, differences in argument structure, and
+item's own combinatorial syntax, differences in argument structure, and
 differences in word order.  Accordingly, words have many related
@@
 traditional machine learning approach to supertag prediction would
-treat individual supertags as atoms for classification; geenralizing
+treat individual supertags as atoms for classification; generalizing
 across supertags would require linking model parameters to handcrafted
-features or backoff categories.
+features or back-off categories.
 
@@
 abilities to generalize across sparse data seems to be one of the most
-improtant reasons for the success of deep learning in NLP.
+important reasons for the success of deep learning in NLP.
 
@@
 places no constraints on the learned similarity function, but it also
-ignores the hieararchical structure of the elementary trees
+ignores the hierarchical structure of the elementary trees
 themselves.  Previous work on deep learning with tree structures has
@@
 analogous techniques to encode supertags based on their underlying
-tree structure.  In partiuclar, to embed a supertag, we embed each
+tree structure.  In particular, to embed a supertag, we embed each
 node, group the resulting vectors to form a tensor, and then summarize
@@
 The vectors are grouped into a tensor by placing the root node into the
-first cell of the first row and left-aligning the decendents in the subsequent rows.
+first cell of the first row and left-aligning the descendants in the subsequent rows.
 %
@@
 %
-An example of a recurrent tree network is shown in eFigure \ref{fig:rtn}.
+An example of a recurrent tree network is shown in figure \ref{fig:rtn}.
 
@@
 However, using topological indices and a tree state eliminates the need for flow control
-by creating locally linear, homogenous computations.
+by creating locally linear, homogeneous computations.
 
@@
 Because it has this richer context representation, it takes advantage
-of a slightly different chracterization of the supertag prediction
+of a slightly different characterization of the supertag prediction
 problem to streamline the problem solving involved in using the model.
@@
 %
-The search concludes either when a single complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}
+The search succeeds when the first complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}
 
@@
 %
-In particular, our implementations result in a large amount of insertion trees\footnote{Many of the validation examples had more than $2^{40}$ possible linearizations.}.
+In particular, our implementations result in a large amount of insertion trees.\footnote{Many of the validation examples had more than $2^{40}$ possible linearizations.}
 %
@@
 embedding vector corresponding to each supertag; or
-\emph{structureally}, by using a convolutional coding over the
+\emph{structurally}, by using a convolutional coding over the
 supertag tree structure.
@@
 
-We trained six such models using a common experimental plantform.  We
+We trained six such models using a common experimental platform.  We
 started from the Wall Street Journal sections of the Penn Treebank,
 which have been previously used for evaluating statistical tree
-grammars \cite{chiang2000statistical}\footnote{A possible additional
+grammars \cite{chiang2000statistical}.\footnote{A possible additional
   data source, the data from the 2011 Shared Task on Surface
-  Realization, was not available}.
+  Realization, was not available.}
 %
@@
 look at the accuracy of the supertag predictions directly output by
-each model.  Second, we look at the accurace of the final supertags
+each model.  Second, we look at the accuracy of the final supertags
 obtained by decoding the model predictions to the best-ranked
@@
 %
-The linearization tasks licenses more freedom in supertag
-classifications because it may have the same root and leaf syntactic
-categories but differ in minor constructions, such as number of
-arguments.
+The linearization task allows more freedom in supertag
+classifications because supertags may differ in minor ways, such as
+the projections present along the spine, which will not affect
+generation output for a particular target input.
 %
-The freedom allows for models to be more generalized and less fit to
-the specific supertags, but at the same time, it also mutes the
+The freedom means models may not be penalized based on decisions that
+don't matter---thus, at the same time, it also mutes the
 distinctions between classification decisions.
@@
 predicts one supertag as a function of another, and so the
-compositional relationships among the two trees matters more---or
-because Fergus-R's contextualized decisions depend on similarities
-among supertags (involving argument structure or information
-structure) that are difficult for the convolutional coding to
-represent or learn.
+compositional relationships among the two trees are more
+important---or because Fergus-R's contextualized decisions depend on
+similarities among supertags (involving argument structure or
+information structure) that are difficult for the convolutional coding
+to represent or learn.
 
@@
 neural models can be more flexible about the information they take
-into account in decisions, it's especially important in designing
+into account in decisions, it's especially advantageous in designing
 neural architectures to break a problem down into decisions that can
@@
 It seems that our models are not doing a good job of using the
-predictions it makes to triangulate to accurate and consistent
+predictions they make to triangulate to accurate and consistent
 supertags. 
@@
 74.9\% in their best evaluation of FERGUS---on a data set of just 100
-sentences with an average lenth of XXX.
+sentences with an average length of XXX.
 %
@@
 
-There are several lines of related work which explore stochastic tree models, ranging exploration with recurrent neural networks to agglomerative pairwise associations. 
+There are several lines of related work which explore stochastic tree
+models from the standpoint of parsing and understanding, including
+networks that maintain state using recurrences and networks that
+agglomerate information using tree traversals.
 %
+We think the perspective given by NLG is instructive.
+
 Most similar to ours, is the top down tree structure of \newcite{zhang2016top}.
 %
-In this model, first, two long short-term memory (LSTM) networks are used to make branch exploration decisions, and then an additional two LSTM networks are used to continue along these branches. 
+In this model, first, two long short-term memory (LSTM) networks are
+used to make branch exploration decisions, and then an additional two
+LSTM networks are used to continue along these branches.
 %
-The work of \newcite{Tai2015} similarly use an LSTM, but instead work from the leaves upward to merge the hidden states of children inside the recurrent step. 
+The work of \newcite{Tai2015} similarly use an LSTM, but instead work
+from the leaves upward to merge the hidden states of children inside
+the recurrent step.
 %
-The upward merging of children to form representations for parents has also been studied as pairwise agglomerations using recursive neural networks \cite{Socher2010}.
+The upward merging of children to form representations for parents has
+also been studied as pairwise agglomerations using recursive neural
+networks \cite{Socher2010}.
 
-A growing body of work investigates modeling long distance relationships using efficient stacks in modern neural architecture. 
+A growing body of work investigates modeling long distance
+relationships using efficient stacks in modern neural architecture.
 %
-\newcite{dyer2015transition} utilize an indexable LSTM as a neural stack for transition-based parsing.  
+\newcite{dyer2015transition} utilize an indexable LSTM as a neural
+stack for transition-based parsing.
 %
-Similarly, \newcite{bowman2016fast} utilize both a stack and buffer for sentence comprehension by utilizing an indexable structure and performing computations with indices to these structures. 
+Similarly, \newcite{bowman2016fast} utilize both a stack and buffer
+for sentence comprehension by utilizing an indexable structure and
+performing computations with indices to these structures.
 %
-Finally, recent work models long distance relationships using the dynamic composability in ApolloCaffe, a fork of the Caffe library \cite{jia2014caffe}, to build different network structures based on parse trees \cite{Andreas2016LearningTC}.
+Finally, recent work models long distance relationships using the
+dynamic composability in ApolloCaffe, a fork of the Caffe library
+\cite{jia2014caffe}, to build different network structures based on
+parse trees \cite{Andreas2016LearningTC}.
 
@@
 
-While many neural models of grammar have focused on parsing, there has been relatively less attention on generation.
+This paper has explored issues in deep learning of probabilistic tree
+grammars from the standpoint of natural language generation.  
 %
-This is in large part due to the different kinds of decisions that have to be made in order for generation to work.
+For NLG, we need models that predict high-probability structures to
+encode deep linguistic relationships---rather than to infer deep
+relationships from surface cues.
+%
+This problem brings new challenges for learning, as it requires us to
+represent new kinds of linguistic elements and new kinds of structural
+context in order to capture the regularities involved.
+%
+Despite these challenges, however, the problem continues to have the
+mix of data sparsity, rich primitives and combinatorial
+interactions that has made deep learning attractive for use in natural
+language parsing and understanding.
 
-The models outlined in this paper push forward methodologies for modeling tree grammars in neural networks.
+Of the range of models we surveyed here, the best combines a top down
+tree recurrence to cluster contexts with appropriate embedding methods
+to cluster syntactic and lexical elements.
+%
+Our evaluations suggest that the model is more accurate and faster
+than alternative techniques. 
+%
+However, it would still be good to analyze the performance of the
+model more deeply.
+%
+Can we get better results in the key decoding step?  How do human
+readers find the output of the system?
 
-A lot of work has to be done to take advantage of deep learning and translate that to tasks beyond parsing, especially language production tasks.
-
-This is a first step in a larger research program of stochastic generation models with the kind of generalization that deep learning techniques like word embeddings and recurrent neural networks give you.
+Looking forward, we see this research a step towards learned
+models that capture more of the NLG task.
+%
+We plan to explore similar techniques in planning surface text from
+more properly semantic inputs or even from abstract communicative
+goals.
+%
+Further, we plan to integrate learned methods with knowledge-based
+techniques to offer designers more control over system output in
+specific applications.
+%
+Developing methods appropriate to such settings will require
+researchers to revisit the core problems of generalizing
+across linguistic structures and contexts---and, we hope, to build on
+and extend the provisional solutions we have explored here.
 
@@
 The specific parameters that were used are shown in Table \ref{tab:params}.
-The parameters were selected by measured performance on the development portion of the dataset.
-In the accompanying code repository, the full experiment parameters---including progammatic parameters controlling the experimental design---are specified in configuration files.
+The parameters were selected by measured performance on the development portion of the data set.
+In the accompanying code repository, the full experiment parameters---including programmatic parameters controlling the experimental design---are specified in configuration files.
 
@@
 
-The data used in our experiments was taken from the Wall Street Jounral sections of the Treebank corpus.
+The data used in our experiments was taken from the Wall Street Journal sections of the Treebank corpus.
 

### 2016-10-10 paper/draft.tex
for the merge

--- 2016-10-10 paper/draft.tex
@@
 architecture.  Each architecture can embed supertags \emph{minimally},
-by a simple one-hot vector; \emph{atomically}, by learning an
+by learning a scalar corresponding to each supertag; \emph{atomically}, by learning an
 embedding vector corresponding to each supertag; or
-\emph{structureally}, by using a convolutional coding over the
-supertag tree structure.
+\emph{structureally}, by using convolutional coding over each
+supertag's tree structure.
 
@@
 \ref{sec:models} and each of the supertag representations of
-\ref{sec:neural}.  As always, we used Section XXX for development,
-Sections XXX for training, and Section XXX for testing.  A complete
+\ref{sec:neural}.  As always, we used Section 02-21 for development,
+Sections 22 for training, and Section 23 for testing.  A complete
 description of network organization and training parameters is given
-in the appendix.  Our code is available XXX.
+in the appendix.  The code and complete experimental setup are publically available\footnote{https://github.com/braingineer/neural\_tree\_grammar}.
 
@@
 %
-We report XXXMEASURE, following \cite{}.
+We report a modified edit distance measure, Generation String Accuracy, following \cite{bangalore2000evaluation}.
 %

### 2016-10-10 paper/draft.tex

--- 2016-10-10 paper/draft.tex
@@
 architecture.  Each architecture can embed supertags \emph{minimally},
-by a simple one-hot vector; \emph{atomically}, by learning an
+by learning a scalar corresponding to each supertag; \emph{atomically}, by learning an
 embedding vector corresponding to each supertag; or
-\emph{structurally}, by using a convolutional coding over the
-supertag tree structure.
+\emph{structurally}, by using convolutional coding over each
+supertag's tree structure.
 
@@
 In the accompanying code repository, the full experiment parameters---including programmatic parameters controlling the experimental design---are specified in configuration files.
+
+\subsection{Model design}
+
+To facilitate efficient model computation, we restricted the computation for supertag distributions and used vectorized functions.
+%
+The set of supertags that are retrieved for probability distribution computation are set of supertags that are seen paired with the lexical item, reducing the output distribution from over 5000 to under 500.
+%
+This was implemented as a set of indices to the embedding matrix passed into the computation with the rest of the data. 
+%
+Future implementations will explore having the supertag subset lookup table be part of the computation.
+%
+While there are several ways the affinity between supertag and lexical state vector could have been computed, we chose to concatenate the vectors, map them to a new space using a fully connected layer, and compute a score with a vectorized function.
+%
+The vectorized function operation is the same mechanism which calculates the probability distribution used in soft attention.
 

### 2016-10-10 paper/draft.tex
-some changes to related work

@@
 First, we represent supertags as vectors through embedding techniques
-that enable to model to generalize naturally over complex, but related
+that enable to model to generalize over complex, but related
 structures.
@@
 %
-The same language model is to eliminate the confound of
+The same language model is used to eliminate the confound of
 language model performance and measure performance differences in the
@@
 \emph{structurally}, by using convolutional coding over each
-supertag's tree structure.
+supertag's tree structure to form a vector.
+%
+In each case, the vector (a size-one vector in the \emph{mimimal} condition) is concatenated as described in \ref{sec:models}.
 
@@
 74.9\% in their best evaluation of FERGUS---on a data set of just 100
-sentences with an average length of XXX.
-%
-Our evaluation, on XXX sentences with an average length of XXX, is much
+sentences with an average length of 16.7.
+%
+Our evaluation, on 2400 sentences with an average length of 22, is much
 more strenuous.
@@
 There are several lines of related work which explore stochastic tree
-models from the standpoint of parsing and understanding, including
-networks that maintain state using recurrences and networks that
-agglomerate information using tree traversals.
+models from the standpoint of parsing and understanding.
+%, including 
+%networks that maintain state using recurrences and networks that
+%agglomerate information using tree traversals.
 %
 We think the perspective given by NLG is instructive.
-
-Most similar to ours, is the top down tree structure of \newcite{zhang2016top}.
-%
-In this model, first, two long short-term memory (LSTM) networks are
-used to make branch exploration decisions, and then an additional two
-LSTM networks are used to continue along these branches.
-%
-The work of \newcite{Tai2015} similarly use an LSTM, but instead work
-from the leaves upward to merge the hidden states of children inside
-the recurrent step.
-%
-The upward merging of children to form representations for parents has
-also been studied as pairwise agglomerations using recursive neural
-networks \cite{Socher2010}.
-
-A growing body of work investigates modeling long distance
-relationships using efficient stacks in modern neural architecture.
-%
-\newcite{dyer2015transition} utilize an indexable LSTM as a neural
-stack for transition-based parsing.
-%
-Similarly, \newcite{bowman2016fast} utilize both a stack and buffer
-for sentence comprehension by utilizing an indexable structure and
-performing computations with indices to these structures.
-%
-Finally, recent work models long distance relationships using the
-dynamic composability in ApolloCaffe, a fork of the Caffe library
-\cite{jia2014caffe}, to build different network structures based on
-parse trees \cite{Andreas2016LearningTC}.
+%
+Where parsing infers most probable underlying structure, generation infers the most likely way of expressing a semantic structure---with which unlabeled dependency trees have served as our proxy in this work.
+
+The works most similar to ours explicitly model tree structures, but focus on resolving the uncertainty invovled with the latent structure of an observed sentence. 
+%
+For example, the top down tree structure of \newcite{zhang2016top} expresses generation of a dependency structure as the decisions of a set of long short-term memory networks.  
+%
+The competing solutions are different tree structures which generate a single linear form.
+%
+In contrast, the generation problem is concerned with different linear forms that can result from the same tree structure.
+
+Although the methodology presented in this work is closely related to several recent neural networks models for long-distance relationships, it differs distinctly in its treatment of state and search.
+%
+Specifically, forward-planning in a generation task produces a growing horizon of syntactic choices while shrinking the horizon of semantic goals.
+%
+At each step, a generator will choose amongst more possible syntactic operations with a limited number of goals in mind. 
+%
+In contrast, parsing and understanding begins with the surface form and constructs the organized semantic content, either for a downstream decision or for its own sake. 
+%
+The most notable works in this line of research are the recurrent neural network grammars \cite{dyer2016recurrent}, a shift-reduce parser and interpreter \cite{bowman2016fast}, and a dynamic network for composing other neural network modules \cite{Andreas2016LearningTC}.
+%
+Interestingly, there is a common theme of using indexable data structures in neural architecture to make long-distance decisions.
 
@@
     - where the data comes from
+
+In generating the tree structure, it is optimized 
+While this generates a tree structure, it is optimized to discover the most likely dependency structure that matches an input linear form.  
+%
+In this model, first, two long short-term memory (LSTM) networks are
+used to make branch exploration decisions, and then an additional two
+LSTM networks are used to continue along these branches.
+%
+The work of \newcite{Tai2015} similarly use an LSTM, but instead work
+from the leaves upward to merge the hidden states of children inside
+the recurrent step.
+%
+The upward merging of children to form representations for parents has
+also been studied as pairwise agglomerations using recursive neural
+networks \cite{Socher2010}.
+
+differs in its emphasis on
+
+A growing body of work investigates modeling long distance
+relationships using efficient stacks in modern neural architecture.
+%
+\newcite{dyer2015transition} utilize an indexable LSTM as a neural
+stack for transition-based parsing.
+%
+Similarly, \newcite{bowman2016fast} utilize both a stack and buffer
+for sentence comprehension by utilizing an indexable structure and
+performing computations with indices to these structures.
+%
+Finally, recent work models long distance relationships using the
+dynamic composability in ApolloCaffe, a fork of the Caffe library
+\cite{jia2014caffe}, to build different network structures based on
+parse trees.

### 2016-10-11 paper/draft.tex
submitting this version

@@
 %
-They argue that their measure correlates with NLG performance better
-than n-gram counting methods.
+%They argue that their measure correlates with NLG performance better
+%than n-gram counting methods.
 %
@@
   the median running time---which includes model computation and A*
-  search.}
+  search. The structural embeddings are computed with convolutional coding, the atomic embeddings as rows in a matrix, and the minimal embeddings as scalars in a vector.}
 \label{table:accresults}
@@
 %
-The flexible embedding methods perform consistently better,
+The structural and atomic embedding methods consistently perform better,
 suggesting that the clustering capabilities of neural methods is a
@@
 %
-The convolution encoding seems to work better with Fergus-N.  
+With respect to the post-decoding performance, Fergus-N seems able to utilize the structural embeddings more than the atomic embeddings. 
+%Fergus-N seems to utilize the structural embeddings more, as its post-decoding performance is 
+%The convolution encoding seems to work better with Fergus-N.  
 %
@@
 to represent or learn.
+%
+Additionally, the performance of the minimal embeddings suggests that Fergus-N's conditional architecture might provide enough structure to make some predictions greatly simplified. 
+%Fergus-N's conditional supertag predictions might also account for the performance of the minimal encoding method and
 
@@
 better information into account in decoding.
+%
+This is more pronounced in the atomic embeddings than the structural embeddings, which suggests that the lack of structure in the vector representation allows for the model to learn clustering relationships that don't correlate with the structural requirements.
 
@@
 \textbf{Model} & \textbf{Embedding}  & \textbf{Top Scoring} & \textbf{Best Performance} \\ \hline
-Fergus-N & Convolution & 65.80\% & 72.58\% \\ \cline{2-4}
-         & Token       & 65.52\%  & 71.82\% \\ \cline{2-4}
-         & Minimal Token & 63.79\% & 71.09\% \\ \cline{2-4}
+Fergus-N & Structural & 65.80\% & 72.58\% \\ \cline{2-4}
+         & Atomic       & 65.52\%  & 71.82\% \\ \cline{2-4}
+         & Minimal & 63.79\% & 71.09\% \\ \cline{2-4}
 \hline
-Fergus-R & Convolution & 68.22\% & 74.70\% \\ \cline{2-4}
-         & Token       &  69.29\% & 75.56\% \\ \cline{2-4}
-         & Minimal Token &  58.23\% & 65.04\% \\ \cline{2-4}
+Fergus-R & Structural & 68.22\% & 74.70\% \\ \cline{2-4}
+         & Atomic       &  69.29\% & 75.56\% \\ \cline{2-4}
+         & Minimal &  58.23\% & 65.04\% \\ \cline{2-4}
 \hline
@@
 %
+All differences in model are significant using an Independent\footnote{An Indepedent t-test was used instead of a Paired-Sample t-test because of intermittent failures during linearization that resulted in slightly different numbers of observations.} t-test ($p<10^{-5}$).
+%
+For both models, the differences between structural embeddings (using convolutional coding) and atomic embeddings (using standard vector embedding techniques) were not significant, while the differences between the two embeddings and minimal embeddings were significant ($p<10^{-5}$).
+%
 The performance confirms our expectation that differences in supertag
@@
 %
-Our evaluation, on 2400 sentences with an average length of 22, is much
+Our evaluation, on 2400 sentences with an average length of 22, is
 more strenuous.
@@
 models from the standpoint of parsing and understanding.
-%, including 
-%networks that maintain state using recurrences and networks that
-%agglomerate information using tree traversals.
-%
-We think the perspective given by NLG is instructive.
-%
-Where parsing infers most probable underlying structure, generation infers the most likely way of expressing a semantic structure---with which unlabeled dependency trees have served as our proxy in this work.
+%
+While using the same methods, NLG has different goals and we think the perspective is instructive. 
+%
+Where parsing infers most probable underlying structure, generation infers the most likely way of expressing a semantic structure.
+%
+This divergence of goals leads to different concerns, alternatives, and emphasis. 
+%---which unlabeled dependency trees have served as our proxy in this work.
 
@@
 %
-For example, the top down tree structure of \newcite{zhang2016top} expresses generation of a dependency structure as the decisions of a set of long short-term memory networks.  
-%
-The competing solutions are different tree structures which generate a single linear form.
+For example, the top down tree structure of \newcite{zhang2016top} expresses the generation of a dependency structure as the decisions of a set of long short-term memory networks.  
+%
+For each decision, the possible options are different tree structures which can produce the target linear form.
 %
 In contrast, the generation problem is concerned with different linear forms that can result from the same tree structure. 
+%
+In more extensive tasks, the generation problem can include simulated interpretion to inform decisions; using the ease of structural inference from linear form quantifies the understandability of a sentence.
 
@@
 %
-At each step, a generator will choose amongst more possible syntactic operations with a limited number of goals in mind. 
-%
-In contrast, parsing and understanding begins with the surface form and constructs the organized semantic content, either for a downstream decision or for its own sake. 
+At each step, syntactic operations grow the number of available syntactic choices while limiting the number of semantic goals left to express. 
+%
+In contrast, parsing and understanding begin with the surface form and construct the organized semantic content, either for a downstream decision or just for the structure itself. 
 %
@@
 %
-Interestingly, there is a common theme of using indexable data structures in neural architecture to make long-distance decisions.
+Interestingly, there is a common theme of using indexable and dynamic data structures in neural architecture to make long-distance decisions.
 
@@
 In particular, this section covers the data processing, parameters used in the final model, and design decisions for the restricting the supertag distributions.
+As a reminder, the code, tex sources, and experimental setup are hosted on github\footnote{https://github.com/braingineer/neural\_tree\_grammar}.
 
@@
 
-This work used data processed from the Wall Street Journal portion of the Treebank corpus.
-Specifically, section 02-21 were used for training, section 22 for development, and section 23 for testing.
-The corpus was preprocessed using the Stanford Parser to fix some common issues as well as remove information extra to syntactic category.
-The resulting parse trees were processed.
-They were annotated with head rules that marked the head at each node in the parse tree and which nodes were dependents or adjuncts.
-These annotated trees were split at adjunction and substitution positions to form the grammar.
+The corpus was preprocessed using the Stanford Parser to fix common issues and remove extraneous information.
+The resulting parse trees were then annotated by marking the head words, the dependents, and the adjuncts.
+The annotated trees were split at adjunction and substitution positions to form the grammar.
 For the dependency structures used in the models, we use the derivation tree formed by the grammar for each parse tree.
-The derivation structures were further processed to parent-child pairs and top-down traversals for each of the models.
-The preprocessing and processing code are distributed with this paper.
+Derivation trees were processed to parent-child pairs and top-down traversals for Fergus-N and Fergus-R.
+All of these steps are released with the code.
 
@@
 
-To facilitate efficient model computation, we restricted the computation for supertag distributions and used vectorized functions.
-%
-The set of supertags that are retrieved for probability distribution computation are set of supertags that are seen paired with the lexical item, reducing the output distribution from over 5000 to under 500.
-%
-This was implemented as a set of indices to the embedding matrix passed into the computation with the rest of the data. 
-%
-Future implementations will explore having the supertag subset lookup table be part of the computation.
+A subset of all supertags are used in supertag probability distribution calculations
+%
+They are the set of supertags that are seen paired with the lexical item.
+%
+This reduces the output distribution from over 5000 to under 500.  
+%
+During computations, indices to the supertag embedding matrix are passed into the computation with the rest of the data. 
 %

### 2016-10-11 paper/draft.tex
jk. submitting this version. license implanted

@@
 
+\blfootnote{
+    % % final paper: en-us version (to licence, a license)
+    %
+    \hspace{-0.65cm}  % space normally used by the marker
+    This work is licenced under a Creative Commons 
+    Attribution 4.0 International License.
+    License details:
+    \url{http://creativecommons.org/licenses/by/4.0/}
+}
+
 Where natural language understanding systems face problems of
@@
 description of network organization and training parameters is given
-in the appendix.  The code and complete experimental setup are publically available\footnote{https://github.com/braingineer/neural\_tree\_grammar}.
+in the appendix.  The code and complete experimental setup are publically available\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}.
 
@@
 In particular, this section covers the data processing, parameters used in the final model, and design decisions for the restricting the supertag distributions.
-As a reminder, the code, tex sources, and experimental setup are hosted on github\footnote{https://github.com/braingineer/neural\_tree\_grammar}.
+As a reminder, the code, tex sources, and experimental setup are hosted on github\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}.
 

### 2016-10-11 paper/draft.tex
begin the shrinkification

@@
 \maketitle
+
+\vspace*{-0.5cm}
 \begin{abstract}
@@
 \section{Introduction}
+\vspace*{-0.2cm}
 
@@
 \label{sec:tree}
+\vspace*{-0.2cm}
 
@@
 \subsection{Grammar Formalism}
+\vspace*{-0.1cm}
 
@@
 \subsection{Grammar Induction}
+\vspace*{-0.1cm}
 
@@
  \label{fig:spineembedding}
+\vspace*{-0.5cm}
 \end{figure*}
@@
 \label{sec:neural}
+\vspace*{-0.2cm}
 
@@
 \subsection{Embedding Supertags}
+\vspace{-0.1cm}
 
@@
 
-The simplest way to embed supertags is simply to treat each structure
-as distinct token that indexes a corresponding learned vector.  This
+The simplest way to embed supertags is to treat each structure
+as a distinct token that indexes a corresponding learned vector.  This
 places no constraints on the learned similarity function, but it also
 ignores the hierarchical structure of the elementary trees
-themselves.  Previous work on deep learning with tree structures has
-suggested that convolutional coding can produce embeddings that
-directly exploit similarities in tree structure.  Thus, we developed
+themselves.  Previous work on deep learning with graph structures
+suggests convolutional neural networks can exploit similarities in structure.  Thus, we developed
 analogous techniques to encode supertags based on their underlying
@@
 
+\vspace*{-0.5cm}
 \begin{align}
@@
 \end{align}
+\vspace*{-0.5cm}
 
@@
 \label{subsec:rtn}
+\vspace{-0.1cm}
 
@@
  \label{fig:rtn}
+\vspace*{-0.5cm}
 \end{figure*}
@@
 \label{sec:models}
+\vspace*{-0.2cm}
 
@@
 \subsection{Model 1: Fergus-N}
+\vspace{-0.1cm}
 
@@
 defined in Eq. \ref{eq:fergusn_obj}.
-%
+
+\vspace*{-0.5cm}
 \begin{align}
@@
 \end{align}
-%
+\vspace*{-0.49cm}
+
 Here $tag_p$ is the parent
@@
 \subsection{Model 2: Fergus-R}
+\vspace{-0.1cm}
 
@@
 \subsection{Linearization}
+\vspace{-0.1cm}
 
@@
 \label{sec:expt}
+\vspace*{-0.2cm}
 
@@
 \subsection{Training}
+\vspace{-0.1cm}
 
@@
 \subsection{Performance Metrics}
+\vspace{-0.1cm}
 
@@
 \label{sec:results}
+\vspace*{-0.1cm}
 
@@
 \label{table:accresults}
+\vspace*{-0.5cm}
 \end{table}
@@
 \label{table:linresults}
+\vspace*{-0.5cm}
 \end{table}
@@
 \label{sec:relatedwork}
+\vspace*{-0.2cm}
 
@@
 \label{sec:conclusion}
+\vspace*{-0.2cm}
 
@@
 \section*{Acknowledgments}
+\vspace*{-0.2cm}
 
@@
 
-\bibliography{coling2016}
-\bibliographystyle{acl}
-
-\newpage
+%\newpage
 \appendix
@@
 \label{sec:appendix}
+\vspace{-0.2cm}
 
@@
 \subsection{Data}
+\vspace*{-0.1cm}
 
@@
 \subsection{Model parameters}
+\vspace*{-0.1cm}
 
@@
 
-\subsection{Model design}
-
-A subset of all supertags are used in supertag probability distribution calculations
-%
-They are the set of supertags that are seen paired with the lexical item.
-%
-This reduces the output distribution from over 5000 to under 500.  
-%
-During computations, indices to the supertag embedding matrix are passed into the computation with the rest of the data. 
-%
-While there are several ways the affinity between supertag and lexical state vector could have been computed, we chose to concatenate the vectors, map them to a new space using a fully connected layer, and compute a score with a vectorized function.
-%
-The vectorized function operation is the same mechanism which calculates the probability distribution used in soft attention.
+\newpage
+\bibliography{coling2016}
+\bibliographystyle{acl}
 

### 2016-10-11 paper/draft.tex
if it fits, it submits

@@
 \centering
-\includegraphics[width=\textwidth]{rtn.pdf}
+\includegraphics[width=0.8\textwidth]{rtn2.pdf}
 \caption{A recurrent tree network. (A) The dependency structure as a tree.  (B) the dependency structure as a sequence.}
@@
 
-\vspace*{-0.5cm}
+\vspace*{-0.55cm}
 \begin{align}
@@
 \end{align}
-\vspace*{-0.49cm}
+\vspace*{-0.45cm}
 
@@
 \end{align}
-\vspace*{-0.5cm}
+\vspace*{-0.45cm}
 
@@
 %
-With respect to the post-decoding performance, Fergus-N seems able to utilize the structural embeddings more than the atomic embeddings. 
+For post-decoding performance, Fergus-N utilizes the structural embeddings more than the atomic embeddings. 
 %Fergus-N seems to utilize the structural embeddings more, as its post-decoding performance is 
@@
 %
-Additionally, the performance of the minimal embeddings suggests that Fergus-N's conditional architecture might provide enough structure to make some predictions greatly simplified. 
+Additionally, the minimal embeddings suggests that Fergus-N's architecture might provide enough structure to make some predictions greatly simplified. 
 %Fergus-N's conditional supertag predictions might also account for the performance of the minimal encoding method and
@@
 \label{tab:params}
+\vspace*{-0.5cm}
 \end{table}

### 2016-10-11 paper/draft.tex
unshrinkage

@@
 
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \begin{abstract}
@@
 \section{Introduction}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \label{sec:tree}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \subsection{Grammar Formalism}
-\vspace*{-0.1cm}
+%\vspace*{-0.1cm}
 
@@
 \subsection{Grammar Induction}
-\vspace*{-0.1cm}
+%\vspace*{-0.1cm}
 
@@
  \label{fig:spineembedding}
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \end{figure*}
@@
 \label{sec:neural}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \subsection{Embedding Supertags}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
 themselves.  Previous work on deep learning with graph structures
-suggests convolutional neural networks can exploit similarities in structure.  Thus, we developed
+suggests convolutional neural networks can exploit similarities in structure \cite{niepert2016learning,kalchbrenner2014convolutional}.  Thus, we developed
 analogous techniques to encode supertags based on their underlying
@@
 
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \begin{align}
@@
 \end{align}
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 
@@
 \label{subsec:rtn}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
  \label{fig:rtn}
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \end{figure*}
@@
 \label{sec:models}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \subsection{Model 1: Fergus-N}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
 
-\vspace*{-0.55cm}
+%\vspace*{-0.55cm}
 \begin{align}
@@
 \end{align}
-\vspace*{-0.45cm}
+%\vspace*{-0.45cm}
 
@@
 \subsection{Model 2: Fergus-R}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
 \subsection{Linearization}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
 \label{sec:expt}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \subsection{Training}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
 \subsection{Performance Metrics}
-\vspace{-0.1cm}
+%\vspace{-0.1cm}
 
@@
 \label{sec:results}
-\vspace*{-0.1cm}
+%\vspace*{-0.1cm}
 
@@
 \label{table:accresults}
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \end{table}
@@
 \label{table:linresults}
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \end{table}
@@
 \label{sec:relatedwork}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \label{sec:conclusion}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \section*{Acknowledgments}
-\vspace*{-0.2cm}
+%\vspace*{-0.2cm}
 
@@
 \label{sec:appendix}
-\vspace{-0.2cm}
+%\vspace{-0.2cm}
 
@@
 \label{tab:params}
-\vspace*{-0.5cm}
+%\vspace*{-0.5cm}
 \end{table}
@@
 \subsection{Data}
-\vspace*{-0.1cm}
+%\vspace*{-0.1cm}
 
@@
 \subsection{Model parameters}
-\vspace*{-0.1cm}
+%\vspace*{-0.1cm}
 
@@
 
+\subsection{Model design}
+A subset of all supertags are used in supertag probability distribution calculations
+%
+They are the set of supertags that are seen paired with the lexical item.
+%
+This reduces the output distribution from over 5000 to under 500.  
+%
+During computations, indices to the supertag embedding matrix are passed into the computation with the rest of the data. 
+%
+While there are several ways the affinity between supertag and lexical state vector could have been computed, we chose to concatenate the vectors, map them to a new space using a fully connected layer, and compute a score with a vectorized function.
+%
+The vectorized function operation is the same mechanism which calculates the probability distribution used in soft attention.
+
 \newpage

### 2016-10-11 paper/draft.tex
down to length

@@
 as a function of those items' learned features.  The resulting
-abilities to generalize across sparse data seems to be one of the most
+ability to generalize across sparse data seems to be one of the most
 important reasons for the success of deep learning in NLP.
@@
 
-The nodes of a supertag can be independently embedded and combined to form a
-tensor of embeddings.
-%
-Specifically, symbols representing the syntactic category and node roles are treated as
-distinct vocabulary tokens, mapped to integers, and used to retrieve a vector
-representation that is learned during training.
-%
-The vectors are grouped into a tensor by placing the root node into the
-first cell of the first row and left-aligning the descendants in the subsequent rows.
-%
-The two tensors are combined by concatenating along the embedding dimension.
+We first embed each node independently, then combined the vectors to
+form a tensor of embeddings.
+%
+Specifically, symbols representing the syntactic category and node
+roles are treated as distinct vocabulary tokens, mapped to integers,
+and used to retrieve a vector representation that is learned during
+training.
+%
+The vectors are grouped into a tensor by placing the root node into
+the first cell of the first row and left-aligning the descendants in
+the subsequent rows.
+%
+The two tensors are concatenated along the embedding dimension.
 %
@@
 The use of topological indices allows for many recurrent tree networks to be
-run in parallel on a GPU, increasing the efficiency of the implementation.
-%
-The primary concern for running parallel GPU computations is homogeneity because
-the same operation will be applied to the entire data structure.
-%
-Normally, tree operations require flow control, making homogeneity impossible.
-%
-However, using topological indices and a tree state eliminates the need for flow control
-by creating locally linear, homogeneous computations.
+run in parallel on a GPU for efficiency.
+%
+GPU implementations must be formulated homogeneously so that 
+the same operations are always applied to the entire data structure.
+%
+Normally, tree operations require flow control, making homogeneity
+impossible.
+%
+However, using topological indices and a tree state eliminates the
+need for flow control by creating locally linear, homogeneous
+computations.
 

### 2016-10-11 paper/draft.tex
training and development fixed

@@
 inventory of supertags to match a corpus.  For NLG, we also need to
-predict how likely a supertag is to fit a specific lexical item in the
-context of a generation task.  We approach this problem using neural
+predict the most likely supertag for any lexical item given the
+generation context.  We approach this problem using neural
 networks. 
@@
 series of convolutional layers.
-
+%
 %\vspace*{-0.5cm}
@@
 %\vspace*{-0.5cm}
-
+%
 The final product, a vector per supertag, is aggregated with the other vectors
@@
 %
-During training and test time, supertags are input as indices and their feature
-representations retrieved as an embedding.
-%
-Importantly, the convolutional layers are connected to the computational graph during training, so the parameters are optimized with respect to the task.
+During training and test time, supertags are simply input as indices
+and their feature representations retrieved as an embedding.
 
@@
 \includegraphics[width=0.8\textwidth]{rtn2.pdf}
-\caption{A recurrent tree network. (A) The dependency structure as a tree.  (B) the dependency structure as a sequence.}
+\caption{A recurrent tree network. (A) The dependency structure as a tree.  (B) The dependency structure as a sequence.}
  \label{fig:rtn}
@@
 The recurrent state $s_c$ is stored in the tree state, $S_{tree}$, at index $t$.
-
+%
 %\vspace*{-0.5cm}
@@
 %\vspace*{-0.5cm}
-
+%
 The use of topological indices allows for many recurrent tree networks to be
@@
 GPU implementations must be formulated homogeneously so that 
-the same operations are always applied to the entire data structure.
-%
-Normally, tree operations require flow control, making homogeneity
-impossible.
-%
-However, using topological indices and a tree state eliminates the
-need for flow control by creating locally linear, homogeneous
-computations.
+the same operations are applied across the entire data structure.
+%
+Normally, tree operations involve conditional access to parent nodes,
+but using topological indices and a tree state accesses the parent in
+a homogeneous way.
 
@@
 defined in Eq. \ref{eq:fergusn_obj}.
-
+%
 %\vspace*{-0.55cm}
@@
 %\vspace*{-0.45cm}
-
+%
 Here $tag_p$ is the parent
@@
 The final operation induces a probability distribution over the treelet and leaf vectors using a score computed by the vectorized function, $\Psi_{predict}$, as the scalar in a softmax distribution. 
-
+%
 %\vspace*{-0.55cm}
@@
 At generation time, we are given a full dependency tree.  A decoding
-step is necessary to compute a high probability
-assignment for all supertags simultaneously.
-%
-There are two primary difficulties that arise in this computation.
-%
-First, the conditional relationship of parents on children implies that the probability for the root supertag depends on the supertags of the entire tree.
-%
-Second, while it is easy to maintain local consistency---matching the syntactic
-category of the child to an appropriate node on the parent---two children may choose substitution supertags which assume attachment to the same position on the parent.
-
-To efficiently decode the supertag classifications, we implement an A* algorithm to incrementally select consistent supertag assignments.
+step is necessary to compute a high probability assignment for all
+supertags simultaneously.
+%
+In this process, tags for children must be chosen consistently with
+one another, and the resulting probabilistic information must be
+propagated upward to rerank tags higher in the tree.
+%
+We solve this problem with an A* algorithm.
 %
@@
 %
-The search succeeds when the first complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}
+The search succeeds when the first complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed.}
 
@@
 \ref{eq:rtnobj} where $lex_C$ is the children's lexical information, $lex_p$ is the parent's lexical information, $tag_p$ is the supertag for the parent node, and $RTN$ is the recurrent tree network.
-
+%
 %\vspace*{-0.55cm}
@@
 The last step is to compute the probability of each supertag using the vectorized function, $\Psi_{predict}$.
-
+%
 %\vspace*{-0.55cm}
@@
 %
-Abstracting syntactic information from the derivation tree leads to
-the unlabeled dependency trees our models assume as input.
+Removing supertags from the derivation tree leads to the unlabeled
+dependency trees our models assume as input.
 
@@
 \ref{sec:models} and each of the supertag representations of
-\ref{sec:neural}.  As always, we used Section 02-21 for development,
-Sections 22 for training, and Section 23 for testing.  A complete
+\ref{sec:neural}.  As always, we used Section 02-21 for training,
+Sections 22 for development, and Section 23 for testing.  A complete
 description of network organization and training parameters is given
-in the appendix.  The code and complete experimental setup are publically available\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}.
+in the appendix.  The code and complete experimental setup are
+publically
+available.\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}
 
@@
 
-In this appendix, we describe some design decisions that took place which did not have a place in the paper but should still be reported.
-In particular, this section covers the data processing, parameters used in the final model, and design decisions for the restricting the supertag distributions.
-As a reminder, the code, tex sources, and experimental setup are hosted on github\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}.
-
-\subsection{Data}
-%\vspace*{-0.1cm}
-
-The corpus was preprocessed using the Stanford Parser to fix common issues and remove extraneous information.
-The resulting parse trees were then annotated by marking the head words, the dependents, and the adjuncts.
-The annotated trees were split at adjunction and substitution positions to form the grammar.
-For the dependency structures used in the models, we use the derivation tree formed by the grammar for each parse tree.
-Derivation trees were processed to parent-child pairs and top-down traversals for Fergus-N and Fergus-R.
-All of these steps are released with the code.
-
-\subsection{Model parameters}
-%\vspace*{-0.1cm}
-
-All of the models were implemented in the Keras \cite{chollet2015keras} and Theano \cite{theano} libraries.
-The specific parameters that were used are shown in Table \ref{tab:params}.
-The parameters were selected by measured performance on the development portion of the data set.
-In the accompanying code repository, the full experiment parameters---including programmatic parameters controlling the experimental design---are specified in configuration files.
-
-\subsection{Model design}
-A subset of all supertags are used in supertag probability distribution calculations
-%
-They are the set of supertags that are seen paired with the lexical item.
-%
-This reduces the output distribution from over 5000 to under 500.  
-%
-During computations, indices to the supertag embedding matrix are passed into the computation with the rest of the data. 
-%
-While there are several ways the affinity between supertag and lexical state vector could have been computed, we chose to concatenate the vectors, map them to a new space using a fully connected layer, and compute a score with a vectorized function.
-%
-The vectorized function operation is the same mechanism which calculates the probability distribution used in soft attention.
+All of our models were implemented in the Keras
+\cite{chollet2015keras} and Theano \cite{theano} libraries.  The
+specific parameters that were used are shown in Table
+\ref{tab:params}.  The parameters were selected by measured
+performance on the development portion of the data set.  In the
+accompanying code repository, the full experiment
+parameters---including programmatic parameters controlling the
+experimental design---are specified in configuration files.
+
+In our experiments, the corpus was preprocessed using the Stanford
+Parser to fix common issues and remove extraneous information.  The
+resulting parse trees were then analyzed to mark the head words, the
+dependents, and the adjuncts.  The marked-up trees were split at
+adjunction and substitution positions to form the grammar.  Our models
+use an output distribution that's restricted to the set of supertags
+that have occurred with the lexical item, which requires indices to
+the supertag embedding matrix to be passed into the computation with
+the rest of the data.  We implement the affinity matrix between the
+supertag embeddings and lexical state vectors, by concatenating the
+vectors, mapping them to a new space using a fully connected layer,
+and computing a score with a vectorized function.
+%
+(The vectorized function operation is the same mechanism which
+calculates the probability distribution used in soft attention.)
 

### 2016-10-11 paper/draft.tex
figure->Figure
spell check and typos

@@
     \hspace{-0.65cm}  % space normally used by the marker
-    This work is licenced under a Creative Commons 
+    This work is licensed under a Creative Commons 
     Attribution 4.0 International License.
@@
 %
-This means inducing a tree grammar from a data set is a matter of inferring the set of constructions in the data.
+This means that inducing a tree grammar from a data set is a matter of inferring the set of constructions in the data.
 
@@
 themselves.  Previous work on deep learning with graph structures
-suggests convolutional neural networks can exploit similarities in structure \cite{niepert2016learning,kalchbrenner2014convolutional}.  Thus, we developed
+suggests convolutional neural networks can exploit similarities in structure \cite{kalchbrenner2014convolutional,niepert2016learning}.  Thus, we developed
 analogous techniques to encode supertags based on their underlying
@@
 %
-An example of a recurrent tree network is shown in figure \ref{fig:rtn}.
+An example of a recurrent tree network is shown in Figure \ref{fig:rtn}.
 
@@
 %
-The child supertag, $tag_c$, is embedded with $G_\Lambda$ and concatenated the lexical vector, $\omega_{lex}$, forming an intermediate vector representation of the node, $\omega_{node}$.
+The child supertag, $tag_c$, is embedded using the target supertag embedding
+$G_s$ and concatenated with the lexical vector, $\omega_{lex}$, forming an intermediate vector representation of the node, $\omega_{node}$.
 %
@@
 &\omega_{lex} = FC_1([G_w(lex_p); G_w(lex_c)]) \label{eq:fergusn} \\
-&\omega_{node}=concat([G_\Lambda(tag_c);~\omega_{lex}]) \nonumber \\
-&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_\Lambda(tagset_p)]) \nonumber \\
-&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_\Lambda(tagset_c)]) \nonumber \\
+&\omega_{node}=concat([G_s(tag_c);~\omega_{lex}]) \nonumber \\
+&\Omega_{treelet} = concat([repeat(\omega_{node}),~G_s(tagset_p)]) \nonumber \\
+&\Omega_{leaf} = concat([repeat(\omega_{lex}),~G_s(tagset_c)]) \nonumber \\
 &P_\theta(tag_{p,i} | lex_{p}, lex_{c}, tag_{c})=
@@
 one another, and the resulting probabilistic information must be
-propagated upward to rerank tags higher in the tree.
+propagated upward to rerank tags elsewhere in the tree.
 %
@@
 &\omega_{node} = RTN(\omega_{lex},~topology) \nonumber \\
-&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_\Lambda(tagset_p))) \nonumber \\
+&\Omega_{elementary} = FC_2(concat(repeat(\omega_{node}),~G_s(tagset_p))) \nonumber \\
 &P_\theta(tag_{p,i}~|~RTN,~lex_p,~lex_{C}) = 
@@
 %
-There are many possible orderings due to adjunction operation
-What remains to be determined is the order in which adjuncts are attached. 
+There are many possibilities, primarily due to ambiguities in insertion
+order.
 %
@@
 %
-In each case, the vector (a size-one vector in the \emph{mimimal} condition) is concatenated as described in \ref{sec:models}.
+In each case, the vector (a size-one vector in the minimal condition) is concatenated as described in \ref{sec:models}.
 
@@
 \ref{sec:models} and each of the supertag representations of
-\ref{sec:neural}.  As always, we used Section 02-21 for training,
-Sections 22 for development, and Section 23 for testing.  A complete
+\ref{sec:neural}.  As always, we used Sections 02-21 for training,
+Section 22 for development, and Section 23 for testing.  A complete
 description of network organization and training parameters is given
 in the appendix.  The code and complete experimental setup are
-publically
+publicly
 available.\footnote{\url{https://github.com/braingineer/neural_tree_grammar}}
@@
 %
-All differences in model are significant using an Independent\footnote{An Indepedent t-test was used instead of a Paired-Sample t-test because of intermittent failures during linearization that resulted in slightly different numbers of observations.} t-test ($p<10^{-5}$).
+All differences in model are significant using an
+Independent\footnote{An Independent t-test was used instead of a
+  Paired-Sample t-test because of intermittent failures during
+  linearization that resulted in slightly different numbers of
+  observations.} t-test ($p<10^{-5}$).
 %
@@
 %
-Where parsing infers most probable underlying structure, generation infers the most likely way of expressing a semantic structure.
+Where parsing infers the most probable underlying structure,
+generation infers the most likely way of expressing a semantic
+structure.
 %
@@
 
-The works most similar to ours explicitly model tree structures, but focus on resolving the uncertainty invovled with the latent structure of an observed sentence. 
+The works most similar to ours explicitly model tree structures, but focus on resolving the uncertainty involved with the latent structure of an observed sentence. 
 %
@@
 %
-In more extensive tasks, the generation problem can include simulated interpretion to inform decisions; using the ease of structural inference from linear form quantifies the understandability of a sentence.
+In more extensive tasks, the generation problem can include simulated interpretation to inform decisions; using the ease of structural inference from linear form quantifies the understandability of a sentence.
 

### 2016-10-11 paper/draft.tex
merigng

--- 2016-10-11 paper/draft.tex
@@
 %
-Additionally, the minimal embeddings suggests that Fergus-N's architecture might provide enough structure to make some predictions greatly simplified. 
+Additionally, the minimal embeddings suggests that Fergus-N's architecture might provide enough structure to reduce the difficulty of a large number of cases. 
 %Fergus-N's conditional supertag predictions might also account for the performance of the minimal encoding method and
@@
 %
-All differences in model are significant using an Independent\footnote{An Indepedent t-test was used instead of a Paired-Sample t-test because of intermittent failures during linearization that resulted in slightly different numbers of observations.} t-test ($p<10^{-5}$).
+All differences in model are significant using an Independent t-test ($p<10^{-5}$).
+\footnote{An Indepedent t-test was used instead of a Paired-Sample t-test because of intermittent failures during linearization that resulted in slightly different numbers of observations.}
 %
@@
 %
-Our evaluation, on 2400 sentences with an average length of 22, is
+Our evaluation, on 2400 sentences with an average length of 22.1, is
 more strenuous.
@@
 
-\begin{table}[t!]
+\begin{table}[b!]
+\vspace{-0.1cm}
 \centering
@@
 
-In our experiments, the corpus was preprocessed using the Stanford
-Parser to fix common issues and remove extraneous information.  The
+In our experiments, the corpus was preprocessed using Stanford NLP tools \cite{de2006generating} to fix common issues and remove extraneous information.  The
 resulting parse trees were then analyzed to mark the head words, the

### 2016-10-11 paper/draft.tex

--- 2016-10-11 paper/draft.tex
@@
 %
-All differences in model are significant using an
-Independent\footnote{An Independent t-test was used instead of a
-  Paired-Sample t-test because of intermittent failures during
-  linearization that resulted in slightly different numbers of
-  observations.} t-test ($p<10^{-5}$).
+All differences in model are significant using an Independent t-test ($p<10^{-5}$).
+\footnote{An Indepedent t-test was used instead of a Paired-Sample t-test because of intermittent failures during linearization that resulted in slightly different numbers of observations.}
 %
